---
title: "Taxon check"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

# 1 LOAD
```{r clear, echo=F}
#rm(list = ls())
```



```{r, echo=F, warning=F, message=FALSE}
#Load required packages
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)
library(plantlist)
library(vegdata)
library(kewr)
library(RSQLite)
```


# 2 DATA
Here we load the data and chose which dataset we want to use.  
We later remove all rows without a plot ID and check which taxa are present in the dataset
```{r Load data, message= F, warning=FALSE}
fast <- F
if(fast) {
  eva <- read_delim("../EVA Data/onePromilleEva.csv", "\t")
} else {
  eva <- read_delim("../EVA Data/171_NeophyteInvasions20230216_notJUICE_species.csv", "\t")
}

length(unique(eva$PlotObservationID))
# Filter out all rows which don't have a plot ID
eva <- eva[!is.na(eva$PlotObservationID),]

# Which taxon groups are there in EVA?
unique(eva$`Taxon group`)
# there are quite some groups not to be included
# only vascular plants will be studied
```


# 3 EURO+MED
## 3.1 Load database
```{r}
# connect to the database
con <- dbConnect(drv=RSQLite::SQLite(), dbname="C:/Users/u0166342/Documents/Boeren/Impact/eva_neophytes/EIVE Data/EuroSL.sqlite")

# list all tables
tables <- dbListTables(con)
tables <- tables[tables != "sqlite_sequence"]

lDataFrames <- vector("list", length=length(tables))
for (i in seq(along=tables)) {
  lDataFrames[[i]] <- dbGetQuery(conn=con, statement=paste("SELECT * FROM '", tables[[i]], "'", sep=""))
}

# retrieve data with euro med classification
Euro <- lDataFrames[[1]]

rm(lDataFrames,tables,con)
```


## 3.2 Groups
First we will make sure that all taxon groups correspond per species
```{r}
# unique names in database
eva_names <- eva |> select(`Matched concept`, `Taxon group`)
eva_names <- eva_names[!duplicated(eva_names),]

# compare with length unique matched concept
length(unique(eva_names$`Matched concept`))
# there is a deviation between the number of unique species and number of unique species-taxon combinations

# we look at the duplicated species
# retrieve duplicated species
dup <- eva_names[eva_names$`Matched concept` %in% eva_names$`Matched concept`[duplicated(eva_names$`Matched concept`)],]
# get all species with known taxon group
dup_known <- dup[!dup$`Taxon group`=="Unknown",]
# change unkown taxa groups to known
dup$`Taxon group`[dup$`Taxon group`=="Unknown"]<- dup_known$`Taxon group`[match(dup$`Matched concept`[dup$`Taxon group`=="Unknown"],
                                                                                dup_known$`Matched concept`)]

# look at which species are still duplicated
no_dup <- dup[!(dup$`Matched concept` %in% dup$`Matched concept`[duplicated(dup)]),]
no_dup <- no_dup[!duplicated(no_dup),]

# these species are wrong and should be changed
false <- data.frame(Genus = c("Hepatica nobilis", "Lamprothamnium papulosum", "Leptorhaphis epidermidis", "Mycoblastus fucatus", 
                              "Peltaria alliacea", "Peltaria emarginata", "Hepatica species"),
                    TaxonGroup = c("Vascular plant", "Alga", "Lichen", "Lichen", "Vascular plant","Vascular plant", "Vascular plant"))

# change the taxon groups to those are false
dup$`Taxon group`[dup$`Matched concept` %in% false$Genus] <- 
  false$TaxonGroup[match(dup$`Matched concept`[dup$`Matched concept` %in% false$Genus], false$Genus)]

# check whether all species are now better
no_dup <- dup[(dup$`Matched concept` %in% dup$`Matched concept`[duplicated(dup)]),]
no_dup <- no_dup[!duplicated(no_dup),]

# check whether all were changed
any(duplicated(dup$`Taxon group`[dup$`Matched concept` %in% unique(dup$`Taxon group`)]))

# change in eva and eva_names these taxons
eva_names$`Taxon group`[eva_names$`Matched concept` %in% no_dup$`Matched concept`] <- 
  no_dup$`Taxon group`[match(eva_names$`Matched concept`[eva_names$`Matched concept` %in% no_dup$`Matched concept`],
                             no_dup$`Matched concept`)]
eva$`Taxon group`[eva$`Matched concept` %in% no_dup$`Matched concept`] <- 
  no_dup$`Taxon group`[match(eva$`Matched concept`[eva$`Matched concept` %in% no_dup$`Matched concept`],no_dup$`Matched concept`)]
```


We now check whether the taxon groups correspond to TPL
```{r}
# get again unique eva_names
eva_names <- eva_names[!duplicated(eva_names),]
colnames(eva_names)<- c("species","taxon")

# take the TPL groups
plant_list <- TPL(eva_names$species)

# append species and taxon names
plant_list$species <- eva_names$species[match(plant_list$YOUR_SEARCH, eva_names$species)]
plant_list$taxon <- eva_names$taxon[match(plant_list$species, eva_names$species)]

# change names angiosperms, gymnosperms and ferns to Vascular plants
plant_list$GROUP <- gsub("Angiosperms","Vascular plant",plant_list$GROUP)
plant_list$GROUP <- gsub("Gymnosperms","Vascular plant",plant_list$GROUP)
plant_list$GROUP <- gsub("Ferns and lycophytes","Vascular plant",plant_list$GROUP)

# which species are vascular plant delineated by us and not by TPL
vasc <- plant_list[plant_list$taxon=="Vascular plant" & plant_list$GROUP != "Vascular plant" & !is.na(plant_list$GROUP),]
# no problem, all are gnetales which are also plants

# which are vascular plant by TPL and not unknown or vascular plant by EVA
no_vasc <- plant_list[ (plant_list$taxon !="Unknown"& plant_list$taxon !="Vascular plant" ) & 
                         plant_list$GROUP == "Vascular plant" & !is.na(plant_list$GROUP),]
# 23 --> wrong TPL
```


Remove all other groups
```{r}
# check which are in there
unique(eva$`Taxon group`)
# remove
sum(eva$`Taxon group` %in% c("Lichen", "Moss", "Alga", "Stonewort", "Mushroom"))
eva <- eva[!eva$`Taxon group` %in% c("Lichen", "Moss", "Alga", "Stonewort", "Mushroom"),]
# check which are in there
unique(eva$`Taxon group`)
```


## 3.3 Euro+Med
```{r}
# unique names and taxon group combinations (same length as unique species)
eva_names <- as.data.frame(unique(eva$`Matched concept`))
colnames(eva_names)<- c("species")
eva_names$taxon <- eva$`Taxon group`[match(eva_names$species, eva$`Matched concept`)]

# we check length
length(unique(eva$`Matched concept`))

# add other data from euro+med
eva_names$euro <- Euro$TaxonConcept[match(vegdata::taxname.abbr(eva_names$species), vegdata::taxname.abbr(Euro$TaxonName))]

# check how many unknown species are in euro+med
sum(!is.na(eva_names$euro[eva_names$taxon=="Unknown"]))
# how many in total
sum(!is.na(eva_names$euro))
# how many not
sum(is.na(eva_names$euro))

# add to eva
eva$species <-eva_names$euro[match(eva$`Matched concept`, eva_names$species)]
```

## 3.4 Idoia file
```{r}
idoia <- readxl::read_excel("~/Boeren/Impact/eva_neophytes/Core authors/Neophyte paper_taxonomic issues_IB.xlsx")

change <- data.frame(old = c("Acantholimon echinus", "Adenostyles platyphylloides", "Anemone narcissifolia subsp. biarmiensis", "Anthemis montana var. Linneana","Astragalus microcephalus","Atropis distans var. convulata", "Avena bromoides subsp. australis","Bidens tripartita","Bromus tomentellus","Campanula biebersteiniana","Campanula spatulata subsp. sprunerana","Carduus argemone subsp. obtusisquamus","Carex acuta x nigra", "Centaurea pulcherrima","Diplotaxis virgata subsp. Virgata","Festuca oreophila","Nepeta mussinii","Ranunculus auricomus coll.","Taraxacum gr alpinum","Thymus praecox subsp. penyalarensis"),
            new= c("Acantholimon ulicinum", "Caucasalia pontica","Anemonastrum narcissiflorum subsp. crinitum","Anthemis cretica","Astracantha microcephala","Puccinellia festuciformis","Helictochloa cincinnata","Bidens tripartitus","Bromopsis tomentella","Campanula tridentata", "Campanula spatulata","Carduus defloratus","Carex x elytroides","Psephellus pulcherrimus","Diplotaxis virgata subsp. virgata","Festuca valesiaca subsp. hypsophila","Nepeta racemosa subsp. racemosa","Ranunculus auricomus aggr.","Taraxacum alpinum aggr.","Thymus praecox subsp. ligusticus"))

check <- eva_names[eva_names$species %in% change$old,]

# how may observations will be changed
sum(!is.na(match(eva$`Matched concept`[is.na(eva$species)], change$old)))

# change in eva
eva$species[is.na(eva$species)] <- change$new[match(eva$`Matched concept`[is.na(eva$species)], change$old)]

# also change in eva names
eva_names$euro[is.na(eva_names$euro)]<- change$new[match(eva_names$species[is.na(eva_names$euro)], change$old)]
# check how many unique species are extra added to our dataset file
sum(is.na(eva_names$euro))
```

Check species now not yet
```{r}
eva_names <- eva_names[is.na(eva_names$euro),]
```


# 4 POWO
## 4.1 Powo search
Search internet
```{r}
# species that came up later and had to be checked
change <- data.frame(old= c("Carex divulsa  subsp. coriogyne"), new= c("Carex divulsa subsp. coriogyne"))
eva$`Matched concept`[eva$`Matched concept` %in% change$old] <- change$new

# which species are not in data
eva_names <- as.data.frame(unique(eva$`Matched concept`[is.na(eva$species)]))
colnames(eva_names)<- "species"
# add taxon
eva_names$taxon <- eva$`Taxon group`[match( eva_names$species, eva$`Matched concept`)]

if(fast){
  # search powo and store results in list if we find a direct match
  mybiglist <- list()
  for(i in 1: length(eva_names$species)){
    # powo
    r<- search_powo(vegdata::taxname.abbr(eva_names$species[i]))
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      mybiglist[[i]] <- tmp
    } else {
      mybiglist[[i]] <- NA
    }
  } 
  #saveRDS(mybiglist, "species_powo_new.RData")
} else {
    # read all data
    mybiglist<- readRDS("species_powo_new.RData")
}
```


## 4.2 Accepted names
```{r}
# empty dataframe
accepted <- data.frame(new_name = character(), old_name = character(), number= numeric(), taxon= character(), acc = numeric())

test <- mybiglist[!is.na(mybiglist[])]

# give the accepted name to the old species name
for(i in 1: length(eva_names$species)){
  # first check if there is an accepted name
  if(!all(is.na(mybiglist[[i]]))){
    # check whether there are accepted names
    if(sum(mybiglist[[i]]$accepted==T)){
      if(sum(mybiglist[[i]]$accepted==T)<2){
          accepted <- add_row(accepted, new_name= mybiglist[[i]]$name[mybiglist[[i]]$accepted==T], old_name =eva_names$species[i],  
                              number= i, taxon= eva_names$taxon[i], acc= 1)
      } else {
          accepted <- add_row(accepted, new_name= mybiglist[[i]]$name[mybiglist[[i]]$accepted==T], old_name =eva_names$species[i],  
                              number= i, taxon= eva_names$taxon[i], acc= 2) 
          }
    } else {
      accepted <- add_row(accepted, new_name= NA, old_name =eva_names$species[i] ,  number= i, taxon=eva_names$taxon[i], acc= 0)
    }
  } else {
      accepted <- add_row(accepted, new_name= NA,old_name = eva_names$species[i],  number= i, taxon= eva_names$taxon[i], acc=0 )
  } 
}


# some species have multiple accepted names
accepted_dup <- accepted[accepted$old_name %in% accepted$old_name[duplicated(accepted$old_name)],]
length(unique(accepted_dup$old_name))

# check the unique names
extra<- as.data.frame(unique(accepted_dup$old_name))
# use TPL to make work easier
plant_list<- TPL(extra$`unique(accepted_dup$old_name)`)

# these are definetely wrong so remove
remove <- c("unknown* species","Grass species","? species", "Mos ?", "Unknown species", "Moss species","Fungus species", 
            "unknown species", "Unknown species 3", "Unknown species 5", "moss* species", "Unknown* species", "Other mosses", 
            "Gray lichen", "Small grass", "Spec species", "Tuberous plant","Rubiales species","Orchidaceae* species","Asteraceae* species","Cruciferae* species","Boraginaceae* species","Morio species", "Decumbens species","Fusarium species","Pteridophyta species","Nostoc species","Genus species","Liverworts* species","Unknown sp", "bryophytes* species","Algae species","Mos species","Orchidacea sp","Gras species","Sponge species","Algae* species", "Algas species", "Liverwort species","Palustris species", "Chamaecytisus species")


# remove these species from the duplicates and eva
accepted_dup <- accepted_dup[!accepted_dup$old_name %in% remove,]
eva <- eva[!eva$`Matched concept` %in% remove,]

# look at accepted_dup which species remain and assign ourselves
unique(accepted_dup$old_name)

# some species were not found
accepted_na <- accepted[is.na(accepted$new_name),]


# only retain unique new names
accepted<- accepted[!(accepted$old_name %in% accepted$old_name[duplicated(accepted$old_name)] | is.na(accepted$new_name)),]

# check which species are no vascular plants and are assigned
accepted_no_vasc <- accepted[!accepted$taxon=="Vascular plant",]

# not even necessary step to check if they are all vascular plants --> powo only has plants
plant_list<- TPL(accepted_no_vasc$new_name)
any(plant_list$GROUP=="Bryophytes")

# assign taxon 'vascular plant'
accepted_no_vasc$taxon <- rep("Vascular plant", times=length(accepted_no_vasc$taxon))

# assign taxon to all accepted species
accepted$taxon[accepted$old_name %in% accepted_no_vasc$old_name]<- 
  accepted_no_vasc$taxon[match(accepted$old_name[accepted$old_name %in% accepted_no_vasc$old_name], accepted_no_vasc$old_name)]


eva_names <- as.data.frame(unique(eva$`Matched concept`[is.na(eva$species)]))
colnames(eva_names)<- "species"
# add taxon
eva_names$taxon <- eva$`Taxon group`[match( eva_names$species, eva$`Matched concept`)]

# add to eva
eva$species[eva$`Matched concept` %in% accepted$old_name] <- 
  accepted$new_name[match(eva$`Matched concept`[eva$`Matched concept` %in% accepted$old_name], accepted$old_name)]
```


# 5 NAs
## 5.1 Mosses Euro Med
```{r}
# which species are present and unknown taxon group --> mosses from euro+med
unknown <- eva[!is.na(eva$species)& eva$`Taxon group`=="Unknown",]
plant_list <- TPL(unknown$species)
remove <- plant_list$YOUR_SEARCH[plant_list$GROUP=="Bryophytes" & !is.na(plant_list$GROUP)]
length(unique(remove))

# remove these from database
eva <- eva[!eva$species %in% remove,]
```



## 5.2 Irena file
```{r}
# load file 
taxonNameCorrections <- read_csv("../Neophyte-Impact/Neophyte Assignments/UniqueTaxaEurope-2023-04-23.csv", show_col_types = FALSE)


# add column
eva$irena <- taxonNameCorrections$species[match(tolower(eva$`Matched concept`),                                                    tolower(taxonNameCorrections$Matched.concept))]

# create list species irena not we
irena <- as.data.frame(unique(eva$irena[is.na(eva$species)]))
colnames(irena)<-"species"
irena$old_names <- eva$`Matched concept`[match(irena$species, eva$irena)]
irena$taxon <- eva$`Taxon group`[match(irena$species, eva$irena)]

change <- data.frame(old = c("Acantholimon echinus", "Adenostyles platyphylloides", "Anemone narcissifolia subsp. biarmiensis", "Anthemis montana var. Linneana","Astragalus microcephalus","Atropis distans var. convulata", "Avena bromoides subsp. australis","Bidens tripartita","Bromus tomentellus","Campanula biebersteiniana","Campanula spatulata subsp. sprunerana","Carduus argemone subsp. obtusisquamus","Carex acuta x nigra", "Centaurea pulcherrima","Diplotaxis virgata subsp. Virgata","Festuca oreophila","Nepeta mussinii","Ranunculus auricomus coll.","Taraxacum gr alpinum","Thymus praecox subsp. penyalarensis"),
            new= c("Acantholimon ulicinum", "Caucasalia pontica","Anemonastrum narcissiflorum subsp. crinitum","Anthemis cretica","Astracantha microcephala","Puccinellia festuciformis","Helictochloa cincinnata","Bidens tripartitus","Bromopsis tomentella","Campanula tridentata", "Campanula spatulata","Carduus defloratus","Carex x elytroides","Psephellus pulcherrimus","Diplotaxis virgata subsp. virgata","Festuca valesiaca subsp. hypsophila","Nepeta racemosa subsp. racemosa","Ranunculus auricomus aggr.","Taraxacum alpinum aggr.","Thymus praecox subsp. ligusticus"))

irena$idoia <- change$new[match(irena$species, change$old)]

# use tpl to delineate species which are certainly not relevant for us
plant_list <- TPL(vegdata::taxname.abbr(irena$old_names))

# append species and taxon names
plant_list$species <- irena$species[match(plant_list$YOUR_SEARCH, irena$old_names)]
plant_list$taxon <- irena$taxon[match(plant_list$YOUR_SEARCH, irena$old_names)]

# check file for irena species
check <- plant_list[is.na(plant_list$GROUP),]

remove <- c("Hakvrucht cultuurgewas","Graan cultuurgewas","Gras species",
            "Monocotyledonae* species","Liana indet.","Planta indeterminate","Dicot sp.2",
            "Hepaticae (overig)","Keimling species 1","Dycot spec","Kale bodem",
            "Dicot sp.1","Keimling species 2","Plant onbekend","Dicotyledone species",
            "Spermatophyta species","Kiemplant species","Gefaesspflanze species",
            "Dicotyledones species","Microspermae species","Pteridophyta species",
            "Crucifer species","Rubiales species","Hepaticae*","Hepaticopsida",
            "Hepaticae* species","Hepaticopsida species")

# check again
check <- check[!check$YOUR_SEARCH %in% remove,]

# remove indefinetely
eva<- eva[!eva$irena %in% remove,]

# which will we add
length(eva$irena[is.na(eva$species)])
unique(eva$irena[is.na(eva$species)])
# give na species irena names
eva$species[is.na(eva$species)] <- eva$irena[is.na(eva$species)]
```


```{r}
# some species are incorrectly assigned by powo previously and have to be removed, this can be checked by comparing old name and assigned name
remove <- c("bryophytes* species","Hepaticopsida species","Nostoc species", 
            "Bryophyta sp","Nostocaceae species","Algae species",
            "Unknown species 4","Unknown sp","Algae* species","Poaceae",
            "leguminosae...","crustaceous lichens","Biota species",
            "Parmeliaceae species", "Carex sp.1")
eva <- eva[!eva$`Matched concept`%in% remove,]
```

```{r}
eva_names <- unique(eva[, c(6,12:13, 3)])
```


# 6 TPL
## 6.1 Unknowns
```{r}
# look at unknowns without species from POWO or Euro+Med
unknown <- eva[is.na(eva$species)& eva$`Taxon group`=="Unknown",]
unknown <- unknown[!duplicated(unknown$`Matched concept`),]

# Check against TPL
plant_list <- TPL(vegdata::taxname.abbr(unknown$`Matched concept`))
# append species and taxon names
plant_list$species <- unknown$`Matched concept`[match(plant_list$YOUR_SEARCH, vegdata::taxname.abbr(unknown$`Matched concept`))]

# check file for wrong species
check <- plant_list[is.na(plant_list$GROUP),]

# read file delineation taxa Christian to check whether some of our species not recognised in TPL have to be changed
taxonGroupAssignments <- read_csv("../Intermediate Data/taxa_taxon_group_assignments.csv", show_col_types = FALSE)
# combine
check$christian <- taxonGroupAssignments[match(taxonGroupAssignments$Matched.concept, check$species)]

# remove these from eva --> not reasonable checking all of them
eva <- eva[!eva$`Matched concept`%in% check$species,]

##################################################################
# work

# here are some species that should be changed
own_class <- data.frame(species= c("Centaurea bimorpha",
                                   "Aconitum anthora subsp. anthora",
                                   "Ostericum maximowiczii",
                                   "Saccharum ravennae",
                                   "Monanthes brachycaulos",
                                   "Sinapidendron gymnocalyx",
                                   "Betula pendula subsp. pendula",
                                   "Bunium bulbocastanum",
                                   "Asphodelus cerasiferus",
                                   "Rosmarinus eriocalyx",
                                   "Muehlenbeckia platyclada",
                                   "Gypsophila tomentosa",
                                   "Pseudognaphalium luteoalbum",
                                   "Myosotis decumbens subsp. rifana",
                                   "Lonicera kabylica"),
                        old_name=c("Centaurea dimorpha", 
                                  "Aconitum anthoroidetum",
                                  "Angelica maximoviczii",
                                  "Saccharum ravenae",
                                  "Monanthes brachycaulon",
                                  "Sinapidendron gymnocalix",
                                  "Betula xpubescens subsp. xverrucosa",
                                  "Bunium bulbostanum",
                                  "Asphodelus cerasifera",
                                  "Rosmarinus eriocalix",
                                  "Muehlenbeckia platyclados",
                                  "Gypsophila tomentosa L.",
                                  "Pseudognaphalium luteo-album",
                                  "Myosotis alpestre var. rifana",
                                  "Lonicera aborea var. kabylica"))


# change these names in eva
eva$species[eva$`Matched concept` %in% own_class$old_name]<- 
  own_class$species[match(eva$`Matched concept`[eva$`Matched concept` %in% own_class$old_name], own_class$old_name)]
```



## 6.2 Mosses unknown
```{r}
# all species that are still unknown
unknown <- eva[is.na(eva$species) & eva$`Taxon group`=="Unknown",]
unknown <- unknown[!duplicated(unknown$`Matched concept`),]

# check whether we can apply them or not
plant_list <- TPL(unknown$`Matched concept`)
plant_list$species <- unknown$`Matched concept`[match(plant_list$YOUR_SEARCH, unknown$`Matched concept`)]
plant_list$taxon <- unknown$`Taxon group`[match(plant_list$YOUR_SEARCH, unknown$`Matched concept`)]

# remove na and 
remove <- plant_list$species[plant_list$GROUP=="Bryophytes" | is.na(plant_list$GROUP)]
remove <- as.data.frame(remove)

# we remove these species from eva
eva <- eva[!eva$`Matched concept` %in% remove$remove,]
```


## 6.3 File Christian
```{r}
# read file delineation taxa Christian to check whether some of our unknown species have to be changed
taxonGroupAssignments <- read_csv("../Intermediate Data/taxa_taxon_group_assignments.csv", show_col_types = FALSE)

# only names here for assessment to make it quicker
eva_names <- unique(eva[, c(6,12:13, 3)])

# check assessment
check <- eva_names[eva_names$`Matched concept` %in% taxonGroupAssignments$Matched.concept[taxonGroupAssignments$TaxonGroup=="Vascular plant"] &
                     eva_names$`Taxon group`=="Unknown",]
check <- taxonGroupAssignments[taxonGroupAssignments$Matched.concept %in% eva_names$`Matched concept`,]

# assign taxa group
eva$`Taxon group`[eva_names$`Matched concept` %in% taxonGroupAssignments$Matched.concept[taxonGroupAssignments$TaxonGroup=="Vascular plant"]] <-
  "Vascular plant"
```



## 6.4 Rest
```{r}
# how many species are now still present with no species designation
vasc <- eva[!duplicated(eva$`Matched concept`) & is.na(eva$species),]
# how many of these are vascular species
sum(!vasc$`Taxon group`=="Unknown")

# how many observations will we remove
length(eva$species[eva$`Taxon group`=="Unknown" & is.na(eva$species)])
check <- as.data.frame(eva$`Matched concept`[eva$`Taxon group`=="Unknown" & is.na(eva$species)])
colnames(check)<- "species"
check <- check |> group_by(species) |> summarise(n=n())
# remove all unknown species with no name
eva <- eva[!eva$`Matched concept` %in% vasc$`Matched concept`[vasc$`Taxon group`=="Unknown"], ]

# check unique names we will assign (172)
names <- vasc$`Matched concept`[!vasc$`Taxon group`=="Unknown"]

# Assign all species 
eva$species[is.na(eva$species)]<- eva$`Matched concept`[is.na(eva$species)]

# check all species left whether there are still mosses etc
eva_names <- eva[!duplicated(eva$species) & eva$`Taxon group`=="Unknown",]
plant_list <- TPL(eva_names$species)

# assign plant_list
plant_list$species <- eva_names$species[match(plant_list$YOUR_SEARCH, eva_names$species)]

# check whether there are species
remove_mosses <- plant_list$species[plant_list$GROUP=="Bryophytes" & !is.na(plant_list$GROUP)]

# remove
eva <- eva[!eva$`Matched concept` %in% remove_mosses,]
```



# 7 CHECK
```{r}
# check whether there still are na species 
any(is.na(eva$species))

# check how much species we have now
length(unique(eva$species))
length(unique(eva$`Matched concept`))

# check whether we have the duplicated species from the accepted names
check <- eva[eva$`Matched concept` %in% accepted_dup$old_name,]
check <- check[!duplicated(check$`Matched concept`),]
# 200 from the 241 species with multiple entries were given their matched concept name or the name from Irena
# 41 species were removed
```

# 8 Aggregate
## 8.1 Subspecies
```{r}
# unique names
eva_names <- as.data.frame(unique(eva$species))
colnames(eva_names)<- c("species")
eva_names$taxon <- eva$`Taxon group`[match(eva_names$species, eva$`Matched concept`)]

read.delim("../Impact/eva_neophytes/EIVE Data/Names/Austria_ExpertSystem.txt")

# reduce size
x<- vegdata::parse.taxa(eva_names$species)
# look at data
x <- x[,1:5]
# make the epi1 named sect. NA
x$epi1[x$epi1=="sect."] <- NA

eva_names$name <- eva_names |> unite

# combine names
eva_names <- cbind(eva_names, paste(x$genus, coalesce(x$epi1, ""),coalesce(x$rank.suff, ""), coalesce(x$epi2, "")))


# add name
colnames(eva_names)[3]<- "name"
# remove all names for which we put an NA previously
eva_names$name[is.na(x$epi1)]<- NA


# add to eva
eva$name <- eva_names$name[match(eva$species, eva_names$species)]

eva_species <- unique(eva_names$name)
```




# 9 SAVE
```{r}
#write_csv(eva, "eva_euro_own.csv")
#write_csv(eva, "eva_euro_own2.csv")
```

