---
title: "Generate taxon groups"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

```{r clear, echo=F}
rm(list = ls())
```



```{r, echo=F, warning=F, message=FALSE}
#Load required packages
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)
```

# Data
Here we load the data and chose which dataset we want to use.  

We later remove all rows without a plot ID and check which taxa are present in the dataset
```{r Load data, message= F, warning=FALSE}
fast <- T
if(fast) {
  eva <- read_delim("../EVA Data/onePromilleEva.csv", "\t")
} else {
  eva <- read_delim("../EVA Data/171_NeophyteInvasions20230216_notJUICE_species.csv", "\t")
}

# Filter out all rows which don't have a plot ID
eva <- eva[!is.na(eva$PlotObservationID),]

# Which taxon groups are there in EVA?
unique(eva$`Taxon group`)
```

Make a new column with only the genus' name.  
```{r Matched concept,  message= F, warning=FALSE}
# Fast method to extract the first string from the 'Matched concept' column e.g. Genus
# \\1 says to use the first part (ie word) as input for the new column
# []+* is used to take all subsequent letters (with capitals and small letters)
# gsub is the function to replace the text (designated in the first part) with the second part (unless it is like here, then we just take that part) from the column designated last
eva$Genus <- gsub("([A-Za-z]+).*", "\\1", eva$`Matched concept`)

# This worked, all plant observations have a genus
#sum(is.na(eva$Genus))
```

# Wrong taxa
We first create three functions required further in the computation. They do the following:  

1.    Get the ID for each taxon group
2.    Get the name of the taxon group, if it is unique it renders the correct one, if not it displays NA. We make sure that the NA's and Unknowns are not given as unique classes
3.    Give the number of options for the taxon
```{r Functions, message= F, warning=FALSE}
taxonGroupToId <- function(taxonGroup) {
  eva$`Taxon group ID`[match(taxonGroup, eva$`Taxon group`)]
}

computeTaxonGroup <- function(possibleTaxonGroups) {
  # get all possible groups for that specific genus
  uniqueTaxonGroups <- unique(possibleTaxonGroups[!(is.na(possibleTaxonGroups) | possibleTaxonGroups == "Unknown")])
  # if more than 1 return NA because one must be incorrect
  if(length(uniqueTaxonGroups) == 1) {
    uniqueTaxonGroups[1]
  } else {
    NA
  }
}

numberOfDifferentTaxonGroups <- function(possibleTaxonGroups) {
  # count number of possible groups
  length(unique(possibleTaxonGroups[!(is.na(possibleTaxonGroups) | possibleTaxonGroups == "Unknown")]))
}
```

We create a matrix with the taxon group for each genus.  

We later show all geni with multiple taxa, i.e. there is a wrong taxon in Eva. We combine all erroneous inputs in a dataframe per genus and taxon group.
```{r, message= F, warning=FALSE }
# For every genus: if all taxa with that genus have the same taxon group, give it that taxon group. Else give it NA. 
# Furthermore: for every genus count the number of different taxon groups found in EVA for that genus.
taxonGroupForGeni <- eva |> 
  # group eva per genus
  group_by(Genus) |> 
  # summarize by computing for each genus the group, if there are multiple return NA and give number
  summarise(
    TaxonGroup = computeTaxonGroup(`Taxon group`), # Unique =name, else= NA
    NumberOfDifferentTaxonGroups = numberOfDifferentTaxonGroups(`Taxon group`)) # Number of options per genus

# Show the geni for which we have multiple taxon groups. E.g. there must be erroneous taxon groups in EVA.
taxonGroupForGeni[taxonGroupForGeni$NumberOfDifferentTaxonGroups > 1,]

# All geni with more than one taxon group assignment
# Checked for correctness
genusTaxonGroupCorrections <- data.frame(Genus = c("Hepatica", "Lamprothamnium", "Leptorhaphis", "Mycoblastus", "Peltaria"),
                                       TaxonGroup = c("Vascular plant", "Alga", "Lichen", "Lichen", "Vascular plant"))
```
We combine all genus that need a new taxon group.  
First we make a function to obtain a vector with TRUE if taxon starts with a specified genus.  
We then make a new dataframe and store in this one the matched concept and old and new taxon.
```{r, message= F, warning=FALSE}
# Function to obtain a vector with TRUE if taxon starts with a specified genus
hasGenus <- function(taxa, genus) {
 startsWith(taxa, paste(genusTaxonGroupCorrections$Genus[i], " ", sep = ""))
  # check if the taxa (given) stars with the same words as the name of the genus in the correction file
  # paste is used to ensure that the genus name is one word
}

# Compute all taxa which need to be reassigned to a different taxon group and store the result to file
# fist an empty file
taxaTaxonGroupCorrections <- data.frame(`Matched concept` = character(0), NewTaxonGroup = character(0), OldTaxonGroup = character(0))

for(i in seq(1,length(genusTaxonGroupCorrections$Genus))) {
  #make vector with T when the genus in eva corresponds to a genus that must be corrected and has the wrong taxon group
  selection <- eva$Genus == genusTaxonGroupCorrections$Genus[i] & eva$`Taxon group` != genusTaxonGroupCorrections$TaxonGroup[i] 
  # count number of erroneous observations per genus and taxon group
  # give the number of species per group of the same matched concept and taxon group
  erroneousTaxa <- eva[selection,] |> group_by(`Matched concept`, `Taxon group`) |> summarise(n = n())
  # Add all taxa with correction to the final result
  # This is a file that we can use, containing the name, the taxon group we just assigned and the old one
  taxaTaxonGroupCorrections <- add_row(taxaTaxonGroupCorrections, #Dataframe
          Matched.concept = erroneousTaxa$`Matched concept`,  #genus
          NewTaxonGroup = rep(genusTaxonGroupCorrections$TaxonGroup[i],times=length(erroneousTaxa$`Matched concept`)),
          OldTaxonGroup = erroneousTaxa$`Taxon group`) 
}
```

We store the results in a new file
```{r}
#write_csv(taxaTaxonGroupCorrections, "taxa_taxon_group_corrections.csv")
```


# No taxon
We first load the data and check uniqueness of taxa
```{r, message= F, warning=FALSE}
taxonGroupForGeni <- read_csv("../Intermediate Data/taxon_group_for_unassigned_geni.csv")
# check what we will actually change
x<-eva[which(eva$Genus%in%taxonGroupForGeni$Genus),c(1,3,6,12)]
# these are all species with apparently no taxon
# check for sure
y<-eva[eva$`Taxon group`=="Unknown", c(1,3,6,12)]
# check difference and similarity two datasets just created
setdiff(y,x)
# which taxa are in our dataset
unique(taxonGroupForGeni$TaxonGroup)
```

Check all taxa that need to be assigned and save to file.
First create empty dataset.
Then we compute the names of all species for which the genus did not have a taxon and later assign it based on our dataset
```{r, message= F, warning=FALSE}
# Compute all taxa which need to be assigned to a taxon group and store the result to file
taxaTaxonGroupAssignments <- data.frame(`Matched concept` = character(0), TaxonGroup = character(0))

for(i in seq(1,length(taxonGroupForGeni$Genus))) {
  # Select all geni equal to the ones in the dataset
  selection <- eva$Genus == taxonGroupForGeni$Genus[i]
  # Take names of all species of a specific genus that did not have the taxon yet
  taxaToBeAssigned <- unique(eva$`Matched concept`[selection])
  # Add all taxa which need to be assigned to the final result
  taxaTaxonGroupAssignments <- add_row(taxaTaxonGroupAssignments,
                                       Matched.concept = taxaToBeAssigned,
                                       TaxonGroup = taxonGroupForGeni$TaxonGroup[i])
}
```

We store the results in a new file
```{r}
#write_csv(taxaTaxonGroupAssignments, "taxa_taxon_group_assignments.csv")
```



# 4 MONOSPECIFIC 
```{r}
y<- unique(eva$species)

split_after_second_word <- function(text) {
  words <- str_split(text, "\\s+")[[1]]  # Split the text into words
  if (length(words) > 2) {
    new_text <- paste(words[1:2], collapse = " ")  # Combine the first two words
    remaining_text <- paste(words[3:length(words)], collapse = " ")  # Combine the remaining words
    return(c(text,new_text, words[2], remaining_text, words[4]))  # Return both parts
  } else {
    return(c(text,text,words[2],NA, NA))  # If less than 3 words, return the original text and an empty string
  }
}

split_text <- t(sapply(y, split_after_second_word))

# Convert the result to a dataframe
new_df <- as.data.frame(split_text)

# get duplicates of the species name without the subspecies
new_df<- new_df |> get_dupes(V2)

new_df[is.na(new_df)] <- "NA"

y<- unique(new_df$V2)
i<- y[1]

test<- new_df[-c(1:length(new_df$V2)),]
for(i in y){
  tmp<- new_df[new_df$V2==i,]
  x<- tmp[tmp$V4=="NA",]
  for(j in 1:length(tmp$V2)){
    tmp$V1[tmp$V4=="NA"]<- ifelse(x$V3==tmp$V5[j], tmp$V1[j], tmp$V2[j])
    tmp$V4[tmp$V4=="NA"]<- ifelse(x$V3==tmp$V5[j], tmp$V4[j], "NA")
  }
  test<- rbind(test, tmp)
}


```


```

