---
title: "Unique plant per region"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

```{r, echo=FALSE}
rm(list=ls())
```

```{r, message=F}
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
```

# Load data

## Eva

```{r, message=F, warning=F}
# Read in EVA
fast <- F
if(fast) {
  eva <- read_delim("../EVA Data/onePromilleEva.csv", "\t")
} else {
  eva <- read_delim("../EVA Data/171_NeophyteInvasions20230216_notJUICE_species.csv", "\t")
}

# Filter out all rows which don't have a plot ID
eva <- eva[!is.na(eva$PlotObservationID),]
```

## Taxon corrections

```{r, message=F, warning=F}
# Taxa which had incorrect taxon assigned
taxonGroupCorrections <- read_csv("../Intermediate Data/taxa_taxon_group_corrections.csv")
# Places in eva where the matched concept is the same as from the correction file
index <- eva$`Matched concept` %in% taxonGroupCorrections$Matched.concept
# For all places in eva that have to be changed, give the new taxon group based on the matched concept
eva$`Taxon group`[index] <- taxonGroupCorrections$NewTaxonGroup[match(eva$`Matched concept`[index], taxonGroupCorrections$Matched.concept)]

# Geni that did not have a taxon
taxonGroupAssignments <- read_csv("../Intermediate Data/taxa_taxon_group_assignments.csv")
index <- eva$`Matched concept` %in% taxonGroupAssignments$Matched.concept
eva$`Taxon group`[index] <- taxonGroupAssignments$TaxonGroup[match(eva$`Matched concept`[index], taxonGroupAssignments$Matched.concept)]
```

## Filter Eva

```{r}
# Remove all non-vascular plants from eva
eva <- eva[eva$`Taxon group` == "Vascular plant",]
```

# Regions

Read the data where each plot is assigned to a region based on the distance or geometry (from file [Region check](3_Region_Check.rmd))

```{r, message=F}
# Read in the plot to region assignments
plotRegionAssignments <- read_csv("../Intermediate Data/plot_to_region_assignment.csv")
```

Assign regions to EVA. This file might be very usefull when used together with glonaf. I think the order would be something like this:

-   read in glonaf
-   generate for every country a list of species that are native
-   assign for every country the native status
-   look at other species status within Europe, if native elsewhere give intracontinental neophyte
-   otherwise give extra european status

I think this can be done with a for loop? for (i in 1: length(unique(eva$regions))) { country_nat<- if species %in% glonaf$ native status\<- "native" elseif species %in% glonaf\$europe status\<- EU neo else status\<- non-EU neo }

TOO BE CONTINUED

```{r, message=F}
# Remove plots in EVA which are not in the scope of the study 
# Outside of scope of study means Region not assigned, ergo NA
# function returns all plots in eva that are also assigned
eva <- eva[!is.na(plotRegionAssignments$Region[match(eva$PlotObservationID, plotRegionAssignments$PlotObservationID)]),]

# Add a column to EVA which contains in which region each observation lies
eva$Region <- plotRegionAssignments$Region[match(eva$PlotObservationID, plotRegionAssignments$PlotObservationID)]

# Check regions
allRegions <- unique(plotRegionAssignments$Region)
allRegions <- allRegions[!is.na(allRegions)]
allRegions
```


# Species per region
```{r}
# make empty dataframe
allResults <- data.frame(Region = character(0), Matched.concept = character(0), NumberOfPlotsWithTaxa = integer(), PercentageOfPlotsWithTaxa = double())

# For each region: compute all the unique plants in the region and in how many of the plots they appear. 
for(i in 1:length(allRegions)) {
  # Create temperate file in every run with the region we want to assess
  tmp <- eva[eva$Region == allRegions[i],]
  # Make a data frame with the matched concept (immediately also counts the number of times present)
  result <- as.data.frame(table(tmp$`Matched concept`))
  colnames(result)[1] <- "Matched.concept"
  colnames(result)[2] <- "NumberOfPlotsWithTaxa"

  # count number of plots per country
  numberOfPlots <- length(unique(tmp$PlotObservationID))
  # add the percentage of plots
  result$PercentageOfPlotsWithTaxa <- result$NumberOfPlotsWithTaxa / numberOfPlots * 100
  # give the region his name
  result$Region <- allRegions[i]
  
  # add all data together
  allResults <- add_row(allResults, Region = result$Region, 
                        Matched.concept = result$Matched.concept, 
                        NumberOfPlotsWithTaxa = result$NumberOfPlotsWithTaxa, 
                        PercentageOfPlotsWithTaxa = result$PercentageOfPlotsWithTaxa)
}

```

Save data
```{r}
write_csv(allResults, "UniqueTaxaPerRegion.csv")
```

