---
title: "Species diversity and eveness"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

# 1 DATA
## 1.1 Preparation
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(hillR)
library(arsenal)
library(doParallel)
library(FSA)
library(vegan)
library(stringr)
library(janitor)
```

## 1.2 Data loaded
Then, we load data of EVA
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva_cover_all_layer.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
fullPlotData <- read_csv("fullPlotData_cover_all_layer.csv")

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.90,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=F
if(native_intra_analysis){
  species_country_status <- read_csv("eva2_country_status_new.csv")
} else{
species_country_status<- read_csv("species_country_status_new.csv")
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
```


Remove species that are not defined or to be excluded
```{r, warnings=F}
remove<- read_csv("not_defined.csv", show_col_types = FALSE)
remove<- as.vector(unlist(remove))
eva<- eva[!(eva$species %in% remove),]
```

Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status, by= c("Region"= "Region", "species"= "species"))
eva<- eva_country_neophyte
# Look at how much from every type are present
tt<- as.data.frame(table(eva_country_neophyte$Neophyte))
tt

rm(fullPlot2, eva_country_neophyte, remove)
```


# 2 ENS
## 2.1 Prepare
```{r}
# reduce size eva
eva<- eva[,c(1:3)]
# change name
colnames(eva)[3]<- "cover"

# check data layer
# eva$Layer <- as.numeric(eva$Layer)
# We now have some trees, which can possibly affect our calculations (I will have to check the effect for the data )
# eva<- filter(eva, Layer == 0 | (Layer >=4 ) )
# eva<- eva[,-4]

```


## 2.2 Hill numbers
```{r}
# make cluster
n.cores<- parallel::detectCores()-6
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK" )
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# get all unique plot IDs required to split our database
plotID<- unique(eva$PlotObservationID)
# run for all and time
begin<-Sys.time()
x<- foreach(i= 1:100, .combine = "rbind", .packages = c("dplyr","mgcv","tidyr","vegan")) %dopar% {
  # create the length of the to be calculated data chunk
  j <- (length(plotID)/100)
  # create begin and end values for the plot observation ID
  begin<- round((i-1)*j+1)
  end<- round(i*j)
  # make the vegetation dataframe for this chunk
  matrix<- pivot_wider(eva[eva$PlotObservationID %in% plotID[begin:end],], names_from= c("species"),
            values_from =("cover" ), values_fill = 0, values_fn = sum)
  # get the plot values
  plots<- matrix[,1]
  # remove the plot values (required to calculate the hill values)
  matrix<-matrix[,-1]
  # calculate hill values for chunk
  hill<- tsallis(matrix, hill=TRUE, scales=c(0,1,2))
  # bind plot ID
  hill<- cbind(plots, hill)
  # show the total hill dataframe, which will be binded together
  hill
}
# calculate time required
end<-Sys.time()
round(end-begin)

colnames(x)<- c("PlotObservationID","ENS0","ENS1","ENS2")
# stop cluster
stopCluster(my.cluster)
```


## 2.3 Analysis
```{r}
# check the values --> some difference --> related to summing different layers 
all.equal(x[,2], fullPlotData$numberOfVascularPlantSpecies)
(x[x$ENS0 != fullPlotData$numberOfVascularPlantSpecies,])
fullPlotData[fullPlotData$numberOfVascularPlantSpecies != x$ENS0,]

# plot
apply(x[, 2:4], MARGIN=2, FUN = hist)
```


## 2.4 Save
```{r}
save=T
if(save){
  fullPlotData<- left_join(fullPlotData, x)
  write_csv(fullPlotData, "fullPlotData_cover_all_layer.csv")
}
```


# 3 EVA
As a result of this script, I got to know that there are some species duplications in our dataset that will probably
