---
title: "Species diversity and eveness"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

# 1 DATA
## 1.1 Preparation
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(hillR)
library(arsenal)
library(doParallel)
library(FSA)
library(vegan)
library(stringr)
library(janitor)
```

## 1.2 Data loaded
First, we load data of EVA
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
fullPlotData <- read_csv("fullPlotData.csv")

fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=T
if(native_intra_analysis){
  species_country_status <- read_csv("eva2_country_status_new.csv")
} else{
species_country_status<- read_csv("species_country_status_new.csv")
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
```


Remove species that are not defined or to be excluded
```{r, warnings=F}
remove<- read_csv("not_defined.csv", show_col_types = FALSE)
remove<- as.vector(unlist(remove))
eva<- eva[!(eva$species %in% remove),]
```

Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status, by= c("Region"= "Region", "species"= "species"))
eva<- eva_country_neophyte
# Look at how much from every type are present
x<- table(eva_country_neophyte$Neophyte)
```


```{r}
# reduce size eva
eva<- eva[,c(1, 10, 8, 7)]
# change name
colnames(eva)[3]<- "cover"

# check data layer
eva$Layer <- as.numeric(eva$Layer)
#eva<- filter(eva, Layer == 0 | (Layer >=4 ) )
eva<- eva[,-4]

# make vegetation dataframe
matrix<- pivot_wider(eva, names_from= c("species"),
            values_from =("cover" ), values_fill = 0, values_fn = sum)
# remove the plot observation ID
x<- matrix[, -1]
# create Hill numbers
y<- tsallis(x, hill=TRUE, scales=c(0,1,2))
# bind met plot ID
x<- cbind(matrix$PlotObservationID, y)

# varies --> some replicates in original dataset so added as extra species which is not true
all.equal(x[,2], fullPlotData$numberOfVascularPlantSpecies)
```


Additional check species names
```{r}
y<- unique(eva$species)

split_after_second_word <- function(text) {
  words <- str_split(text, "\\s+")[[1]]  # Split the text into words
  if (length(words) > 2) {
    new_text <- paste(words[1:2], collapse = " ")  # Combine the first two words
    remaining_text <- paste(words[3:length(words)], collapse = " ")  # Combine the remaining words
    return(c(text,new_text, words[2], remaining_text, words[4]))  # Return both parts
  } else {
    return(c(text,text,words[2],NA, NA))  # If less than 3 words, return the original text and an empty string
  }
}

split_text <- t(sapply(y, split_after_second_word))

# Convert the result to a dataframe
new_df <- as.data.frame(split_text)

# get duplicates of the species name without the subspecies
new_df<- new_df |> get_dupes(V2)

new_df[is.na(new_df)] <- "NA"

y<- unique(new_df$V2)
i<- y[1]

test<- new_df[-c(1:length(new_df$V2)),]
for(i in y){
  tmp<- new_df[new_df$V2==i,]
  x<- tmp[tmp$V4=="NA",]
  for(j in 1:length(tmp$V2)){
    tmp$V1[tmp$V4=="NA"]<- ifelse(x$V3==tmp$V5[j], tmp$V1[j], tmp$V2[j])
    tmp$V4[tmp$V4=="NA"]<- ifelse(x$V3==tmp$V5[j], tmp$V4[j], "NA")
  }
  test<- rbind(test, tmp)
}

```

