---
title: "Data preparation"
subtitle: "based on the work of Christian Ledergerber"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---
# Start
We first clean the environment
```{r, echo=T}
rm(list=ls())
```

We load the required libraries
```{r, warning=F, message=F}
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(AER)
library(mgcv)
library(modEvA)
library(doParallel)
library(arsenal)
```


# Eva Data
We load the data
```{r , message=FALSE, warning=FALSE}
fast <- F
if(fast) {
  eva <- read_delim("../EVA Data/onePromilleEva.csv", "\t")
} else {
  eva <- read_delim("../EVA Data/171_NeophyteInvasions20230216_notJUICE_species.csv", "\t")
}
```

We remove the rows without observations.
```{r}
# Count the number of rows with NA values
numberOfRowsWithoutPlotID <- sum(is.na(eva$PlotObservationID))
# Check rows with NA
eva[which(is.na(eva$PlotObservationID)),]
# Filter out all rows which don't have a plot ID
eva <- eva[!is.na(eva$PlotObservationID),]
# Count number of plots in eva before data removal
originalPlotNumberEva<- length(unique(eva$PlotObservationID))
```

We update the taxa data
```{r, warning=F, message=FALSE}
# Update taxon group for some species where it was specified incorrectly
taxonGroupCorrections <- read_csv("../Intermediate Data/taxa_taxon_group_corrections.csv")
# What is the location of the species in eva that are specified incorrectly
index <- eva$`Matched concept` %in% taxonGroupCorrections$Matched.concept
# Change to correct for each species
eva$`Taxon group`[index] <- taxonGroupCorrections$NewTaxonGroup[match(eva$`Matched concept`[index], taxonGroupCorrections$Matched.concept)]

# Update taxon group for species that did not have one in EVA
taxonGroupAssignments <- read_csv("../Intermediate Data/taxa_taxon_group_assignments.csv")
index <- eva$`Matched concept` %in% taxonGroupAssignments$Matched.concept
eva$`Taxon group`[index] <- taxonGroupAssignments$TaxonGroup[match(eva$`Matched concept`[index], taxonGroupAssignments$Matched.concept)]

# Some taxon names seem incorrect
taxonNameCorrections <- read_csv("../Neophyte Assignments/UniqueTaxaEurope-2023-04-23.csv")
# give the species name from taxon name corrections to eva in the species column
# to lower makes sure all data is written without capitals
# match gives vector with position of argument eva in taxonNameCorrections, and does this for the total length of eva
eva$species <- taxonNameCorrections$species[match(tolower(eva$`Matched concept`), tolower(taxonNameCorrections$Matched.concept))]
# some data of species is.na, we assign for these the values of the matched concept
eva$species[is.na(eva$species)] <- eva$`Matched concept`[is.na(eva$species)] # fix problem with plant not present in neophyte file from irena (only a handful of plant observations in EVA)
```

We only keep vascular plants and remove the columns in which this data is designated.
```{r}
eva <- eva[eva$`Taxon group` == "Vascular plant",!names(eva) %in% c("Taxon group", "Taxon group ID")]
```

# Header Data
Read in the header information. Look at and remove plots for which:

-   Area is NA or smaller than 0
-   Area is larger than 1000 m²
-   The longitude or latitude data is not known
-   The plot is not assigned to a region
-   The plot only contains presence/absence data

```{r, message=F, warning=F}
header <- read_delim("../EVA Data/171_NeophyteInvasions20230216_notJUICE_header.csv", "\t")
totalNumberOfPlots <- length(header$PlotObservationID)

# Plot sizes too small/ inexistent
plotsWithoutPlotSize <- is.na(header$`Relevé area (m²)`) | header$`Relevé area (m²)` <= 0
numberOfPlotsWithoutPlotSize <- sum(plotsWithoutPlotSize)

# Plot sizes too large
tooLargePlots <- header$`Relevé area (m²)` > 1000
numberOfTooLargePlots<- sum(tooLargePlots, na.rm=T)

# Plots without coordinates
plotsWithoutLocation <- is.na(header$Latitude) | is.na(header$Longitude)
numberOfPlotsWithoutLocation <- sum(plotsWithoutLocation)

# Check whether the amount of observation plots correspond
# This is the data made in 3_Region_check (here we also removed all data without coordinates)
plotsRegionAssignments <- read_csv("../Intermediate Data/plot_to_region_assignment.csv")
# Error message if the two are not the same
if( sum(header$PlotObservationID[!plotsWithoutLocation] != plotsRegionAssignments$PlotObservationID) > 0) {
  message("Error: plotsRegionAssignments expected to contain the same plots as contained in header data.")
}

# Count number of plots unassigned
# create logical vector of header length
plotsWithoutRegionAssignment <- vector(mode = "logical", length = length(header$PlotObservationID))
# make true when the plot has no region, taking into consideration that some plots have no coordinates
plotsWithoutRegionAssignment[!plotsWithoutLocation] <- is.na(plotsRegionAssignments$Region)
numberOfPlotsWithoutRegionAssignment = sum(plotsWithoutRegionAssignment)

# Remove only Presence/ Absence data and NA data
namesForPresenceAbsence <- c("Presence/Absence", "Presentie/Absentie", NA)
# logical vector with length header 
plotsWithPresenceAbsenceOnly <- header$`Cover abundance scale` %in% namesForPresenceAbsence
numberOfPlotsWithoutCoverInformation <- sum(plotsWithPresenceAbsenceOnly)

# sum all plots to be removed
allPlotsToBeRemoved <- plotsWithoutPlotSize | tooLargePlots | plotsWithoutLocation | plotsWithPresenceAbsenceOnly | plotsWithoutRegionAssignment
totalNumberRemoved <- sum(allPlotsToBeRemoved)
```

`r numberOfPlotsWithoutPlotSize` of the plots do not contain a size, `r numberOfTooLargePlots` were larger than 1000m² and therefore most likely no plots, `r numberOfPlotsWithoutLocation` of the plots do not have a location, `r numberOfPlotsWithoutRegionAssignment` cannot be assigned to one of the neophyte regions and `r numberOfPlotsWithoutCoverInformation` of the plots do not contain cover data. All these plots (`r totalNumberRemoved`) are removed:
```{r}
header <- header[!allPlotsToBeRemoved,]
```

We also add the region to the data for analysis and change the region and dataset to a factor.
```{r}
header <- left_join(header, plotsRegionAssignments, by = "PlotObservationID")
header$Region <- as.factor(header$Region)
header$Dataset <- as.factor(header$Dataset)
```

Finally, we will keep only the plots that are still available in both Eva and Header
```{r}
plotsInEvaAndHeader <- intersect(eva$PlotObservationID, header$PlotObservationID)
eva <- eva[eva$PlotObservationID %in% plotsInEvaAndHeader,]
header <- header[header$PlotObservationID %in% plotsInEvaAndHeader,]
```
After filtering `r length(header$PlotObservationID)` plots remain.

# EIVE
Reading in the EIVE data and explore data by boxplotting the moisture (M), nitrogen (N), pH (R), light (L) and temperature (T) variables.
```{r, message=F}
# 08 contains the indicator values for position and width
eive <- read_delim("../EIVE Data/EIVE_Paper_1.0_SM_08.csv", ",")
# data contains n --> number of sources, nw--> niche width and ''--> value

par(mfrow = c(1,1))
boxplot(eive$`EIVEres-M`, xlab="M")
boxplot(eive$`EIVEres-N`, xlab="N")
boxplot(eive$`EIVEres-R`, xlab="R")
boxplot(eive$`EIVEres-L`, xlab="L")
boxplot(eive$`EIVEres-T`, xlab="T")
par(mfrow=c(1,1))
```

The EIVE values are OK balanced with some outliers. No extreme skewness to be expected for the average indicator values at the plot level.

Extend eva with columns from EIVE.
```{r}
# match gives position of species in eva in eive
# positions
eva$EIVEresM <- eive$`EIVEres-M`[match(eva$species, eive$TaxonConcept)]
eva$EIVEresN <- eive$`EIVEres-N`[match(eva$species, eive$TaxonConcept)]
eva$EIVEresR <- eive$`EIVEres-R`[match(eva$species, eive$TaxonConcept)]
eva$EIVEresL <- eive$`EIVEres-L`[match(eva$species, eive$TaxonConcept)]
eva$EIVEresT <- eive$`EIVEres-T`[match(eva$species, eive$TaxonConcept)]

# niche widths
eva$EIVEnwM <- eive$`EIVEres-M.nw3`[match(eva$species, eive$TaxonConcept)]
eva$EIVEnwN <- eive$`EIVEres-N.nw3`[match(eva$species, eive$TaxonConcept)]
eva$EIVEnwR <- eive$`EIVEres-R.nw3`[match(eva$species, eive$TaxonConcept)]
eva$EIVEnwL <- eive$`EIVEres-L.nw3`[match(eva$species, eive$TaxonConcept)]
eva$EIVEnwT <- eive$`EIVEres-T.nw3`[match(eva$species, eive$TaxonConcept)]

numberOfPlantsInEvaFoundInEive <- sum(!(is.na(eva$EIVEresM) | 
                                              is.na(eva$EIVEresN) | 
                                              is.na(eva$EIVEresR) | 
                                              is.na(eva$EIVEresL) | 
                                              is.na(eva$EIVEresT)))
percentEvaInEive<- numberOfPlantsInEvaFoundInEive/length(eva$PlotObservationID)
```


`r numberOfPlantsInEvaFoundInEive` of the `r length(eva$PlotObservationID)` plant observations in Eva have indicator values in EIVE.


TODO: how to handle niche width and number of original EIVs in EIVE to improve the estimation of the EIVE value in EVA?

# DIV
Reading in the disturbance indicator values en plotting severity and frequency.
```{r, messagge=F, warning=F}
div <- read_delim("../Disturbance Data/disturbance_indicator_values.csv", ",")

par(mfrow = c(1,1))
# Relatively balanced
boxplot(div$Disturbance.Severity, xlab="Severity")
boxplot(div$Disturbance.Frequency, xlab="Frequency")
# more skewed (check with hist)
boxplot(div$Mowing.Frequency, xlab="Mowing fr")
hist(div$Mowing.Frequency, xlab="Mowing fr")
boxplot(div$Grazing.Pressure, xlab="Grazing Pr")
boxplot(div$Soil.Disturbance, xlab="Soil dist")
# oke balanced still
hist(div$Grazing.Pressure, xlab="Grazing Pr")
hist(div$Soil.Disturbance, xlab="Soil dist")
```


Extend eva with columns from disturbance indicator values:
```{r}
eva$Disturbance.Severity <- div$Disturbance.Severity[match(eva$species, div$species)]
eva$Disturbance.Severity.herblayer <- div$Disturbance.Severity.herblayer[match(eva$species, div$species)]
eva$Disturbance.Frequency <- div$Disturbance.Frequency[match(eva$species, div$species)]
eva$Disturbance.Frequency.herblayer <- div$Disturbance.Frequency.herblayer[match(eva$species, div$species)]
eva$Grazing.Pressure <- div$Grazing.Pressure[match(eva$species, div$species)]
eva$Mowing.Frequency <- div$Mowing.Frequency[match(eva$species, div$species)]
eva$Soil.Disturbance <- div$Soil.Disturbance[match(eva$species, div$species)]

numberOfPlantsInEvaFoundInDiv <- sum(!is.na(eva$Disturbance.Severity))
percentEvaInDiv<- numberOfPlantsInEvaFoundInDiv/length(unique(eva$species))
length(!is.na(unique((eva$Disturbance.Frequency))))/length(!is.na(unique(eva$species)))
```

`r numberOfPlantsInEvaFoundInDiv` of the `r length(eva$PlotObservationID)` plant observations in Eva have disturbance indicator value.

# Plot level EVA
Aggregating EVA (#plant per plot and average indicator values) and calculating all kinds of weighting methods for the Eive values.
```{r}
# Function to compute the square root of coverage weighted indicator value 
# Calculate the weight for every species in the plot, then take the indicator value and finally compute the weighted mean
computeSquareRootWeightedMean <- function(indicatorValue, coverage) {
  weights <- sqrt(coverage[!is.na(indicatorValue)])
  indicators <- indicatorValue[!is.na(indicatorValue)]
  weighted.mean(indicators, weights)
}

# Function to compute 1/niche-width weighted mean
NWwmean <- function(indicatorValue, nicheWidth) {
  weights <- 1 / nicheWidth[!is.na(indicatorValue)]
  indicators <- indicatorValue[!is.na(indicatorValue)]
  weighted.mean(indicators, weights)
}

# Function to compute sqrt(cover)/niche-width weighted mean
cNWwmean <- function(indicatorValue, cover, nicheWidth) {
  weights <- sqrt(cover[!is.na(indicatorValue)]) / nicheWidth[!is.na(indicatorValue)]
  indicators <- indicatorValue[!is.na(indicatorValue)]
  weighted.mean(indicators, weights)
}

aggregatedEva <- eva |> 
  group_by(PlotObservationID) |> 
  summarise(
    numberOfVascularPlantSpecies = n(), 
    
    EIVEresM.sqrt = computeSquareRootWeightedMean(EIVEresM, `Cover %`), 
    EIVEresN.sqrt = computeSquareRootWeightedMean(EIVEresN, `Cover %`), 
    EIVEresR.sqrt = computeSquareRootWeightedMean(EIVEresR, `Cover %`), 
    EIVEresL.sqrt = computeSquareRootWeightedMean(EIVEresL, `Cover %`),
    EIVEresT.sqrt = computeSquareRootWeightedMean(EIVEresT, `Cover %`),
    
    EIVEresM.nw = NWwmean(EIVEresM, EIVEnwM), 
    EIVEresN.nw = NWwmean(EIVEresN, EIVEnwN), 
    EIVEresR.nw = NWwmean(EIVEresR, EIVEnwR), 
    EIVEresL.nw = NWwmean(EIVEresL, EIVEnwL),
    EIVEresT.nw = NWwmean(EIVEresT, EIVEnwT), 

    EIVEresM.cnw = cNWwmean(EIVEresM, `Cover %`, EIVEnwM), 
    EIVEresN.cnw = cNWwmean(EIVEresN, `Cover %`, EIVEnwN), 
    EIVEresR.cnw = cNWwmean(EIVEresR, `Cover %`, EIVEnwR), 
    EIVEresL.cnw = cNWwmean(EIVEresL, `Cover %`, EIVEnwL),
    EIVEresT.cnw = cNWwmean(EIVEresT, `Cover %`, EIVEnwT), 

    EIVEresM = mean(EIVEresM, na.rm = T), 
    EIVEresN = mean(EIVEresN, na.rm = T), 
    EIVEresR = mean(EIVEresR, na.rm = T), 
    EIVEresL = mean(EIVEresL, na.rm = T), 
    EIVEresT = mean(EIVEresT, na.rm = T), 

    DistSeverity.sqrt = computeSquareRootWeightedMean(Disturbance.Severity, `Cover %`), 
    DistSeverityHerblayer.sqrt = computeSquareRootWeightedMean(Disturbance.Severity.herblayer, `Cover %`),
    DistFrequency.sqrt = computeSquareRootWeightedMean(Disturbance.Frequency, `Cover %`),
    DistFrequencyHerblayer.sqrt = computeSquareRootWeightedMean(Disturbance.Frequency.herblayer, `Cover %`),
    Grazing.Pressure.sqrt = computeSquareRootWeightedMean(Grazing.Pressure, `Cover %`),
    Mowing.Frequency.sqrt = computeSquareRootWeightedMean(Mowing.Frequency, `Cover %`),
    Soil.Disturbance.sqrt = computeSquareRootWeightedMean(Soil.Disturbance, `Cover %`),
    
    DistSeverity = mean(Disturbance.Severity, na.rm = T),
    DistFrequency = mean(Disturbance.Frequency, na.rm = T),
    Grazing.Pressure=mean(Grazing.Pressure, na.rm=T),
    Mowing.Frequency= mean(Mowing.Frequency, na.rm=T),
    Soil.Disturbance=mean(Soil.Disturbance, na.rm=T),
    
    # I would not go for this, it is just replicating 1 fr the length of the cover values
    DistSeverity.cnw = cNWwmean(Disturbance.Severity, `Cover %`, rep(1,length(`Cover %`))),
    DistFrequency.cnw = cNWwmean(Disturbance.Frequency, `Cover %`, rep(1,length(`Cover %`))),
    #Grazing.Pressure.cnw = cNWwmean(Grazing.Pressure, `Cover %`, rep(1,length(`Cover %`))),
    #Mowing.Frequency.cnw = cNWwmean(Mowing.Frequency, `Cover %`, rep(1,length(`Cover %`))),
    #Soil.Disturbance.c,w = cNWwmean(Soil.Disturbance, `Cover %`, rep(1,length(`Cover %`))),
    
    logDistSeverity = mean(log(Disturbance.Severity), na.rm=T),
    logDistFrequency = mean(log(Disturbance.Frequency), na.rm = T),
    logGrazing.Pressure=mean(log(Grazing.Pressure), na.rm=T),
    logMowingFrequency=mean(log(Mowing.Frequency), na.rm=T),
    logSoil.Disturbance=mean(log(Soil.Disturbance), na.rm=T)
    #logDistSeverity = mean(log(Disturbance.Severity + 1/2*min(div$Disturbance.Severity[div$Disturbance.Severity>0])), na.rm = T),
    #logDistFrequency = mean(log(Disturbance.Frequency + 1/2*min(div$Disturbance.Frequency[div$Disturbance.Frequency>0])), na.rm = T)
    )
```


Join the result with the header data:
```{r}
fullPlotData <- full_join(header, aggregatedEva, by = "PlotObservationID")
fullPlotData <- rename(fullPlotData, Area = `Relevé area (m²)`, Date = `Date of recording`)
```

# Filtering
## High SR
Check which plots are extremely species rich:
```{r}
# Which plots SR>100
highDiversityPlots <- fullPlotData[fullPlotData$numberOfVascularPlantSpecies > 100,]

# Jurgen plots suspiciously high
highDiversityPlotsToBeRemoved <- c(842839, 843192, 842921) # plots published by Juergen which are for sure incorrect

# Look at datasets containing the plots with SR> 100
highDiversityDatasets <- unique(highDiversityPlots$Dataset)
# Look when SR is high for more than two plots in a dataset and obtain their observation ID
for(dataset in highDiversityDatasets) {
  # take all plots for a specific dataset
  plotsInDataset <- highDiversityPlots[highDiversityPlots$Dataset == dataset,]
  # take per amount of SR the frequency this specific SR is present
  multiplicityOfRichnessOccurrence <- as.data.frame(table(plotsInDataset$numberOfVascularPlantSpecies))
  # take all times the same SR is present multiple times in a plot
  multiplicityOfRichnessOccurrence <- multiplicityOfRichnessOccurrence[multiplicityOfRichnessOccurrence$Freq >= 2,] # we want to remove only the plots which > 100 species and suspiciously frequent in the database
  for(richness in multiplicityOfRichnessOccurrence$Var1) {
    # take the index of all plots in the dataset with this species richness
    plotsToBeRemoved <- plotsInDataset[plotsInDataset$numberOfVascularPlantSpecies == richness,]
    highDiversityPlotsToBeRemoved <- append(highDiversityPlotsToBeRemoved, plotsToBeRemoved$PlotObservationID)
  }
}


numberOfHighDiversityPlotsToBeRemoved <- length(highDiversityPlotsToBeRemoved)
# remove the too high diversiy plots from the dataset
fullPlotData <- fullPlotData[!(fullPlotData$PlotObservationID %in% highDiversityPlotsToBeRemoved),]

# make a new highdiversityplots dataset for all true diverse plots (according to the method)
highDiversityPlots <- fullPlotData[fullPlotData$numberOfVascularPlantSpecies > 100,]
# plot to have an idea how the SR is distributed
hist(fullPlotData$numberOfVascularPlantSpecies)
#write_csv(highDiversityPlots, "highDiversityPlots.csv")
```

`r numberOfHighDiversityPlotsToBeRemoved` of the plots were removed because they have suspiciously high, non unique richness in the same database or are known to be incorrect.

## Not all IVs calculatable
Filtering out plots for which some of the indicator values could not be estimated:
```{r}
# Cells with NA in fullPlot
fullPlotDataNas <- is.na(fullPlotData)
# Get the plot observations
plotsForWhichNotAllIndicatorValuesWereEstimated <- fullPlotDataNas[,1]
# 35 is number of vascular plants
for(i in 35:dim(fullPlotDataNas)[2]) {
  plotsForWhichNotAllIndicatorValuesWereEstimated <- plotsForWhichNotAllIndicatorValuesWereEstimated | fullPlotDataNas[,i]
}
numberOfPlotsWithMissingIndicatorValues <- sum(plotsForWhichNotAllIndicatorValuesWereEstimated)
# remove the plots without an indivator value
fullPlotData <- fullPlotData[!plotsForWhichNotAllIndicatorValuesWereEstimated,]

# remove the removed plots also from eva, header and aggregated eva
aggregatedEva <- aggregatedEva[aggregatedEva$PlotObservationID %in% fullPlotData$PlotObservationID,]
header <- header[header$PlotObservationID %in% fullPlotData$PlotObservationID,]
eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
```

`r numberOfPlotsWithMissingIndicatorValues` of the plots were removed because not all indicator values could be estimated.

# Plot

```{r}
par(mfrow=c(1,1))
boxplot(fullPlotData$numberOfVascularPlantSpecies)
```

As expected the count data on species richness looks quite skewed. Hence, we will need to use a poisson glm.

```{r}
# Square root
boxplot(fullPlotData$DistSeverity.sqrt)
boxplot(fullPlotData$DistFrequency.sqrt)
boxplot(fullPlotData$EIVEresM.sqrt)
boxplot(fullPlotData$EIVEresN.sqrt)
boxplot(fullPlotData$EIVEresR.sqrt)
boxplot(fullPlotData$EIVEresL.sqrt)
boxplot(fullPlotData$EIVEresT.sqrt)

# nw
#boxplot(fullPlotData$DistSeverity.nw)
#boxplot(fullPlotData$DistFrequency.nw)
boxplot(fullPlotData$EIVEresM.nw)
boxplot(fullPlotData$EIVEresN.nw)
boxplot(fullPlotData$EIVEresR.nw)
boxplot(fullPlotData$EIVEresL.nw)
boxplot(fullPlotData$EIVEresT.nw)

# cnw
boxplot(fullPlotData$DistSeverity.cnw)
boxplot(fullPlotData$DistFrequency.cnw)
boxplot(fullPlotData$EIVEresM.cnw)
boxplot(fullPlotData$EIVEresN.cnw)
boxplot(fullPlotData$EIVEresR.cnw)
boxplot(fullPlotData$EIVEresL.cnw)
boxplot(fullPlotData$EIVEresT.cnw)

# mean
boxplot(fullPlotData$DistSeverity)
boxplot(fullPlotData$DistFrequency)
boxplot(fullPlotData$EIVEresM)
boxplot(fullPlotData$EIVEresN)
boxplot(fullPlotData$EIVEresR)
boxplot(fullPlotData$EIVEresL)
boxplot(fullPlotData$EIVEresT)
```

The distribution of the EIVE predictors look symmetrical. However, the disturbance indicators are heavily skewed. Hence, they need to be transformed. Because there are zeros in the data we choose a root transoformation. Trying different powers we found that the 4th root worked best:

As histogram, is easier for me to see distribution
```{r}
# Square root
hist(fullPlotData$DistSeverity.sqrt)
hist(fullPlotData$DistFrequency.sqrt)
hist(fullPlotData$EIVEresM.sqrt)
hist(fullPlotData$EIVEresN.sqrt)
hist(fullPlotData$EIVEresR.sqrt)
hist(fullPlotData$EIVEresL.sqrt)
hist(fullPlotData$EIVEresT.sqrt)

# nw
#hist(fullPlotData$DistSeverity.nw)
#hist(fullPlotData$DistFrequency.nw)
hist(fullPlotData$EIVEresM.nw)
hist(fullPlotData$EIVEresN.nw)
hist(fullPlotData$EIVEresR.nw)
hist(fullPlotData$EIVEresL.nw)
hist(fullPlotData$EIVEresT.nw)

# cnw
hist(fullPlotData$DistSeverity.cnw)
hist(fullPlotData$DistFrequency.cnw)
hist(fullPlotData$EIVEresM.cnw)
hist(fullPlotData$EIVEresN.cnw)
hist(fullPlotData$EIVEresR.cnw)
hist(fullPlotData$EIVEresL.cnw)
hist(fullPlotData$EIVEresT.cnw)

# mean
hist(fullPlotData$DistSeverity)
hist(fullPlotData$DistFrequency)
hist(fullPlotData$EIVEresM)
hist(fullPlotData$EIVEresN)
hist(fullPlotData$EIVEresR)
hist(fullPlotData$EIVEresL)
hist(fullPlotData$EIVEresT)
```



```{r}
fullPlotData$transformedDisturbanceSeverity.sqrt <- fullPlotData$DistSeverity.sqrt^(1/4)
fullPlotData$transformedDisturbanceFrequency.sqrt <- fullPlotData$DistFrequency.sqrt^(1/4)
fullPlotData$transformedDisturbanceSeverity <- fullPlotData$DistSeverity^(1/4)
fullPlotData$transformedDisturbanceFrequency <- fullPlotData$DistFrequency^(1/4)
fullPlotData$transformedDisturbanceSeverity.cnw <- fullPlotData$DistSeverity.cnw^(1/4)
fullPlotData$transformedDisturbanceFrequency.cnw <- fullPlotData$DistFrequency.cnw^(1/4)

par(mfrow = c(2,2))
boxplot(fullPlotData$transformedDisturbanceSeverity.sqrt)
boxplot(fullPlotData$transformedDisturbanceFrequency.sqrt)
boxplot(fullPlotData$DistSeverity.sqrt) # imo even better distributed...
boxplot(fullPlotData$DistFrequency.sqrt)
```

After transformation with log(x + eps) the disturbance indicator values are much better centered. However, the transformation function is a bit arbitrary and there are still outliers which cause difficulties with analysis as can be seen by fitting a straight forward model:

# Base model test
Note that this is not the final Base Model, rather we will use this to check the assumptions made.
```{r}
glm0 <- glm(numberOfVascularPlantSpecies ~ 
                    log(Area) + 
                    EIVEresM.sqrt + I(EIVEresM.sqrt^2) + 
                    EIVEresN.sqrt + I(EIVEresN.sqrt^2) + 
                    EIVEresR.sqrt + I(EIVEresR.sqrt^2) + 
                    EIVEresL.sqrt + I(EIVEresL.sqrt^2) + 
                    EIVEresT.sqrt + I(EIVEresT.sqrt^2) +
                    transformedDisturbanceSeverity.sqrt + I(transformedDisturbanceSeverity.sqrt^2) +
                    transformedDisturbanceFrequency.sqrt + I(transformedDisturbanceFrequency.sqrt^2)
                    , family=poisson, fullPlotData)

summary(glm0)
par(mfrow = c(2,2))
plot(glm0)
```
We can see that the residuals of this model are in a reasonable range and hence the model fits reasonably well to the data.

## Check individual country data
```{r}
dataFromGermany <- fullPlotData[fullPlotData$Region == "Netherlands",]
dataFromGermany$Dataset <- as.factor(as.character(dataFromGermany$Dataset))
par(mfrow = c(1,1))
boxplot(numberOfVascularPlantSpecies ~ Dataset, data = dataFromGermany)
boxplot(log(fullPlotData$Area))
```

Looking at different datasets from a similar region exhibits quite large differences in terms of species richness. It will most likely make sense to include Dataset as a random factor. The log transformed size of the plots is very symmetrical. This looks good.

# Write CSV
```{r}
# Make sure all plot data is in eva
fullPlotData <- fullPlotData[fullPlotData$PlotObservationID %in% eva$PlotObservationID,]
fullPlotData$Dataset <- as.factor(as.character(fullPlotData$Dataset))
#write_csv(fullPlotData, "fullPlotData.csv")

# Make sure all eva data has a plot 
eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
#write_csv(eva, "fullPlotEva.csv")
```

# Species occurence
Compute which species occurs how often in eva:
```{r}
# For speed optimization we extract only the necessary information from EVA
reducedEva <- eva |> select(PlotObservationID, species, `Cover %`)

uniqueSpecies <- unique(reducedEva$species)

# Empty dataset
plotsPerSpecies <- data.frame(names = character(), numberOfPlots = integer())

parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()


# Count in how many plots a species occurs

begin<- Sys.time()
plotsPerSpecies<-foreach(i= 1:length(uniqueSpecies), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  tmp <- reducedEva$PlotObservationID[reducedEva$species == uniqueSpecies[i]]
  names=uniqueSpecies[i]
  numberOfPlots= length(unique(tmp))
  plotsPerSpecies[i,]<- c(names, numberOfPlots)
}
plotsPerSpecies<- as.data.frame(plotsPerSpecies)
plotsPerSpecies<-setNames(plotsPerSpecies, c("names", "numberOfPlots"))
row.names(plotsPerSpecies)<- c(1: length(uniqueSpecies))
end<- Sys.time()
round(end-begin)

parallel::stopCluster(cl = my.cluster)

#write_csv(plotsPerSpecies, "plotsPerSpeciesInEurope.csv")
```

Determine which plant species have more than 10 plots with cover > 50% and cover < 1% for analysis:
```{r, message=F}
# Load data previously made with the number of plots per species across the dataset
plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope.csv")

parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# Get the names of the species which have at least 20 plots (since we want 10 smaller and 10 bigger)
candidatesForCoverClassAnalysis <- plotsPerSpecies$names[plotsPerSpecies$numberOfPlots >= 20]

# create empty vector
taxaForCoverClassAnalysis <- c()
# run for all candidates
begin<- Sys.time()
taxaForCoverAnalysis<- foreach(i= 1: length(candidatesForCoverClassAnalysis), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar%  {
  # if the sum of all species with a cover of 50% for a species and less than one each are more than 10 
  # && makes sure to only look at the second if it is already T
  taxon<- candidatesForCoverClassAnalysis[i]
  if(sum(eva$species == taxon & eva$`Cover %` >= 50) > 10 && sum(eva$species == taxon & eva$`Cover %` < 1) > 10) {
    # write the species name to the list
    taxaForCoverClassAnalysis[i] <- c(taxon)
  }
}
end<- Sys.time()
round(end-begin)

parallel::stopCluster(cl = my.cluster)
#write_csv(data.frame(taxa = taxaForCoverClassAnalysis), "taxaForCoverClassAnalysis.csv")
```

# To Do
TODO: how to handle niche width and number of original EIVs in EIVE to improve the estimation of the EIVE value in EVA?
TODO also calculate richness without neophytes?
