---
title: 'Abundance analysis based on new classification'
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

This script analyses whether neophytes are more dominant and frequent than native plants or not:

# 1 DATA
## 1.1 Preparation
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(ggplot2)
library(arsenal)
library(doParallel)
library(FSA)
```

## 1.2 Data loaded
This script analyses whether neophytes are more dominant and frequent than native plants or not.  
First, we load data of EVA
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
fullPlotData <- read_csv("fullPlotData.csv")

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=T
if(native_intra_analysis){
  species_country_status <- read_csv("eva2_country_status_new.csv")
} else{
species_country_status<- read_csv("species_country_status_new.csv")
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
```


Remove species that are not defined or to be excluded
```{r, warnings=F}
remove<- read_csv("not_defined.csv", show_col_types = FALSE)
remove<- as.vector(unlist(remove))
eva<- eva[!(eva$species %in% remove),]
```

Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status, by= c("Region"= "Region", "species"= "species"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
```


# 2 COVER
## 2.1 Synthesis
```{r}
if(fast){
# For speed optimization we extract only the necessary information from EVA
reducedEva <- eva_country_neophyte |> select(PlotObservationID, Region, species, Neophyte,`Cover %`)

uniqueSpecies <- unique(reducedEva[, c(3,4)])

# Make empty datafraùe
speciesDominance <- data.frame(names = character(), coverMedian = numeric(), coverMean = numeric(), var = numeric(), numberOfPlots = integer(), neophyte = logical())

# Prepare running in parallel
parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# Take time beginning
begin<- Sys.time()
i=1
# Do calculation of average and median cover+ variance in dataset
speciesDominance <-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  # Get dataset for each unique species
  tmp <- reducedEva[reducedEva$species == uniqueSpecies$species[i]& reducedEva$Neophyte == uniqueSpecies$Neophyte[i],]
  # Make ID a factor
  tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
  # Group by plot and calculate the mean species cover for that species in all plots where it is present
  tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`Cover %`))
  # We are only interested in the dominance of species occurring in at least 50 (?) plots 
  if(length(tmp$PlotObservationID) >= 50) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present, also incorporating the variance and number of Plots
    sum <- summary(tmp$cover)
    x <- c(names = uniqueSpecies$species[i], coverMedian = sum[[3]], coverMean = sum[[4]], var = var(tmp$cover), 
        numberOfPlots =length(tmp$PlotObservationID), neophyte = uniqueSpecies$Neophyte[i])
  }
}
# Ending and calculate time it took to perform the calculations
end<- Sys.time()
round(end-begin)

# Make the matrix a dataframe and change the row names
speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))
# Make the Median, Mean, Variance and Number numeric
speciesDominance[,c(2:5)]<- apply(speciesDominance[, c(2:5)], MARGIN= 2, FUN= as.numeric)
# Stop the cluster
parallel::stopCluster(cl = my.cluster)
} else{
  if(native_intra_analysis){
  speciesDominance<- read.csv("speciesDominance_native_intra.csv")
  }else{
  speciesDominance<- read.csv("speciesDominance_new_categorization_50.csv")
  }
}
```


We also analyse which species are significantly negative / positive when compared to the distribution of the native species. We do this for all plants.
```{r}
# Get the quantile values of all native plant species.Cutoff values at 0.01 and 0.05 chances (both upper and lower limit)
quantiles <- quantile(speciesDominance$coverMean[speciesDominance$neophyte=="native"], probs = c(0, 0.01, 0.05, 0.95, 0.99, 1))

# Assign each species to a class
speciesDominance$significantCoverMean <- as.factor(cut(speciesDominance$coverMean, breaks = quantiles, labels = c("p=0.01 negative", "p=0.05 negative", "", "p=0.05 positive", "p=0.01 positive"), right = TRUE))
```


## 2.5 Analysis
We perform a small analysis on the data by looking at the mean values for native, intra and extra and analyse whether they are significantly different from each other
```{r}
speciesDominance<- speciesDominance[speciesDominance$coverMean>0,]

hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"]))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"]))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"]))
#hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native_intra"]))


model<-(aov(log10(coverMean) ~ neophyte, data= speciesDominance))
summary(model)
TukeyHSD(model)
plot(TukeyHSD(model, conf.level=.95), las = 2)
```


```{r}
# Take a look at the most dominant species
topdominant<- speciesDominance[order(speciesDominance$coverMean, decreasing=T),]
topdominant_intra<- topdominant[topdominant$neophyte=="intra",]
topdominant_extra <- topdominant[topdominant$neophyte=="extra",]
```


## 2.6 Plot
```{r}
# Violin plot
native_intra_analysis=T
if(!native_intra_analysis){
p<-ggplot(speciesDominance, mapping = aes(x= (neophyte), y = log10(coverMean), colour = neophyte)) + 
  # create violin plot with x axis no title and y 
  geom_violin()+ylab("Mean cover when present (%)") +
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European neophyte", "extra European neophyte")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 2.2))+
  scale_x_discrete(labels=  c("extra-European \n neophytes",  "intra-European \n neophytes", "native species"))+ 
  theme(axis.text.x = element_text(size=10), axis.title.y=element_text(size=12), axis.title.x = element_blank(), axis.text.y = element_text(size=10))+
  annotate("text", x = 1, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"])), label = "c", size = 3.5, vjust = -2, hjust = 0.5) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"])), label = "a", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x=1, y=2.1, label= paste("n=",length(extraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.1, label= paste("n=",length(intraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.1, label= paste("n=",length(nativeCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)
p
ggsave("Dominance New 50.jpeg", p,width = 6, height = 4)
}


if(native_intra_analysis){
  p<-ggplot(speciesDominance, mapping = aes(x= (neophyte), y = log10(coverMean), colour = neophyte)) + 
  # create violin plot with x axis no title and y 
  geom_violin()+ylab("Mean cover when present (%)") +
  scale_colour_manual(values=c("#1E88E5","#B71C1C", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native","native_intra", "intra","extra"),
                    labels=c("native species","native species alien elsewhere", "intra European neophyte", "extra European neophyte")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 2.2))+
  scale_x_discrete(labels=  c("extra-European \n neophytes",  "intra-European \n neophytes","native species", "native species \nalien elsewhere"))+
   theme(axis.text.x = element_text(size=10), axis.title.y=element_text(size=12), axis.title.x = element_blank(), axis.text.y = element_text(size=10))+
  annotate("text", x = 1, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"])), label = "d", size = 3.5, vjust = -2, hjust = 0.5) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"])), label = "c", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"])), label = "a", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 4, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x=1, y=2.1, label= paste("n=",sum(speciesDominance$neophyte=="extra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.1, label= paste("n=",sum(speciesDominance$neophyte=="intra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.1, label= paste("n=",sum(speciesDominance$neophyte=="native")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=4, y=2.1, label= paste("n=",sum(speciesDominance$neophyte=="native_intra")), vjust = -0.5, hjust = 0.5, size=3.5)
p
ggsave("Dominance New 50 native_inta.jpeg", p,width = 6, height = 4)
  
  
}
```


# 3 FREQUENCY
## 3.1 Analysis
The data is present in speciesDominance, which facilitates the calculations
```{r}
hist((log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])))
hist(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"]))
hist(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"]))

model<-(aov(log10(numberOfPlots) ~ neophyte, data= speciesDominance))
summary(model)
TukeyHSD(model)
plot(TukeyHSD(model, conf.level=.95), las = 2)
```

## 3.2 Plot
```{r}
# Violin plot
# plot frequency
if(native_intra_analysis){
p<- ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte)) + 
   geom_violin() +
  ylab("log10 (number of occurences)") + xlab(NULL)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European neophyte", "extra European neophyte")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_x_discrete(labels= c("extra-European \n neophytes",  "intra-European \n neophytes", "native species")) +
  theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )  +
  theme(plot.subtitle=element_text(size=12)) + 
  annotate("text", x = 1, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"])), label = "a", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x=1, y=5.2, label= paste("n=",length(extraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=5.2, label= paste("n=",length(intraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=5.2, label= paste("n=",length(nativeCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)
p
#ggsave("Frequency New 50.jpeg", p,width = 6, height = 4)
} else {
p<- ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte)) + 
   geom_violin() +
  ylab("log10 (number of occurences)") + xlab(NULL)+
  scale_colour_manual(values=c("#1E88E5","#B71C1C", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native","native_intra", "intra","extra"),
                    labels=c("native species","native species alien elsewhere", "intra European neophyte", "extra European neophyte")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_x_discrete(labels=  c("extra-European \n neophytes",  "intra-European \n neophytes","native species", "native species \nalien elsewhere"))+
   theme(axis.text.x = element_text(size=10), axis.title.y=element_text(size=12), axis.title.x = element_blank(), axis.text.y = element_text(size=10))+
  annotate("text", x = 1, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"])), label = "a", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"])), label = "c", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 4, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"])), label = "d", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x=1, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="extra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="intra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="native")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=4, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="native_intra")), vjust = -0.5, hjust = 0.5, size=3.5)
p
#ggsave("Frequency New 50 native_intra.jpeg", p,width = 6, height = 4)
  
}
```

# 4 Export
```{r}
#write.csv(speciesDominance, "speciesDominance_new_categorization_50.csv")
```

