---
title: 'Abundance analysis based on new classification'
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

This script analyses whether neophytes are more dominant and frequent than native plants or not:

# 1 DATA
## 1.1 Preparation
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(ggplot2)
library(arsenal)
library(doParallel)
library(FSA)
```

## 1.2 Data loaded
This script analyses whether neophytes are more dominant and frequent than native plants or not.  
First, we load data of EVA
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
fullPlotData <- read_csv("fullPlotData.csv")

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
species_country_status<- read_csv("species_country_status_new.csv")

species_country_status$species[(duplicated(unique(species_country_status$species[species_country_status$Neophyte=="intra"]),unique(species_country_status$species[species_country_status$Neophyte=="extra"])))]

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
```


Remove species that are not defined or to be excluded
```{r, warnings=F}
remove<- read_csv("not_defined.csv", show_col_types = FALSE)
remove<- as.vector(unlist(remove))
eva<- eva[!(eva$species %in% remove),]
```

Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status, by= c("Region"= "Region", "species"= "species"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
```


# 2 COVER
## 2.1 Native
```{r}
# For speed optimization we extract only the necessary information from EVA
reducedEva <- eva_country_neophyte |> select(PlotObservationID, Region, species, Neophyte,`Cover %`)

# Previously it was possible to assign the class later, now we have the definition already in the beginning (as some species in dataset can be native to one region and neophyte in another) this is not possible anymore. We will do this analysis per class and later combine the results
nativeEva <- reducedEva[reducedEva$Neophyte=="native",]
uniqueNative <- unique(nativeEva$species)

# Make empty datafraùe
speciesDominance <- data.frame(names = character(), coverMedian = numeric(), coverMean = numeric(), var = numeric(), numberOfPlots = integer(), neophyte = logical())

# Prepare running in parallel
parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# Take time beginning
begin<- Sys.time()

# Do calculation of average and median cover+ variance in dataset
speciesDominance<-foreach(i= 1:length(uniqueNative), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  # Get dataset for each unique species
  tmp <- nativeEva[nativeEva$species == uniqueNative[i],]
  # Make ID a factor
  tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
  # Group by plot and calculate the mean species cover for that species in all plots where it is present
  tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`Cover %`))
  # We are only interested in the dominance of species occurring in at least 50 (?) plots --> I now ran for all
  if(length(tmp$PlotObservationID) >= 50) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present, also incorporating the variance and number of Plots
    sum <- summary(tmp$cover)
    speciesDominance <- c(names = uniqueNative[i], coverMedian = sum[[3]], coverMean = sum[[4]], var = var(tmp$cover), 
        numberOfPlots =length(tmp$PlotObservationID), neophyte = "native")
  }
}
# Ending and calculate time it took to perform the calculations
end<- Sys.time()
round(end-begin)

# Make the matrix a dataframe and change the row names
speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))
# Make the Median, Mean, Variance and Number numeric
speciesDominance[,2]<- as.numeric(speciesDominance[,2])
speciesDominance[,3]<- as.numeric(speciesDominance[,3])
speciesDominance[,4]<- as.numeric(speciesDominance[,4])
speciesDominance[,5]<- as.numeric(speciesDominance[,5])
# Stop the cluster
parallel::stopCluster(cl = my.cluster)

# Save the data
nativeCover<- speciesDominance
```

## 2.2 Intra-EU
```{r}
# Now per intra-EU
intraEva <- reducedEva[reducedEva$Neophyte=="intra",]
uniqueintra <- unique(intraEva$species)

# Make empty dataframe
speciesDominance <- data.frame(names = character(), coverMedian = numeric(), coverMean = numeric(), var = numeric(), numberOfPlots = integer(), neophyte = logical())

# Prepare parallel
parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# Begin timing
begin<- Sys.time()

# Perform in parallel to get required values
speciesDominance<-foreach(i= 1:length(uniqueintra), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  # Get dataset for each unique species
  tmp <- intraEva[intraEva$species == uniqueintra[i],]
  # Make ID a factor
  tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
  # Group by plot and calculate the mean species cover for that species in all plots where it is present
  tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`Cover %`))
  # We are only interested in the dominance of species occuring in at least 50 plots
  if(length(tmp$PlotObservationID) >= 50) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present
    sum <- summary(tmp$cover)
    speciesDominance <- c(names = uniqueintra[i], coverMedian = sum[[3]], coverMean = sum[[4]], var = var(tmp$cover), 
        numberOfPlots =length(tmp$PlotObservationID), neophyte = "intra")
  }
}
# Time it took
end<- Sys.time()
round(end-begin)

# Dataframe and rownames
speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))
# Make numeric
speciesDominance[,2]<- as.numeric(speciesDominance[,2])
speciesDominance[,3]<- as.numeric(speciesDominance[,3])
speciesDominance[,4]<- as.numeric(speciesDominance[,4])
speciesDominance[,5]<- as.numeric(speciesDominance[,5])

# Stop cluster
parallel::stopCluster(cl = my.cluster)

# Save as intra cover
intraCover<- speciesDominance
```

## 2.3 Extra
```{r}
# Take extra EU species
extraEva <- reducedEva[reducedEva$Neophyte=="extra",]
uniqueextra <- unique(extraEva$species)

# Make empty dataframe
speciesDominance <- data.frame(names = character(), coverMedian = numeric(), coverMean = numeric(), var = numeric(), numberOfPlots = integer(), neophyte = logical())

# Prepare parallel
parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# Time
begin<- Sys.time()

# Perform analysis
speciesDominance<-foreach(i= 1:length(uniqueextra), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  # Get dataset for each unique species
  tmp <- extraEva[extraEva$species == uniqueextra[i],]
  # Make ID a factor
  tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
  # Group by plot and calculate the mean species cover for that species in all plots where it is present
  tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`Cover %`))
  # We are only interested in the dominance of species occuring in at least 50 plots
  if(length(tmp$PlotObservationID) >= 50) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present
    sum <- summary(tmp$cover)
    speciesDominance <- c(names = uniqueextra[i], coverMedian = sum[[3]], coverMean = sum[[4]], var = var(tmp$cover), 
        numberOfPlots =length(tmp$PlotObservationID), neophyte = "extra")
  }
}
# End timing and look at total time passed
end<- Sys.time()
round(end-begin)

# Dataframe and name
speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))
# Numeric columns
speciesDominance[,2]<- as.numeric(speciesDominance[,2])
speciesDominance[,3]<- as.numeric(speciesDominance[,3])
speciesDominance[,4]<- as.numeric(speciesDominance[,4])
speciesDominance[,5]<- as.numeric(speciesDominance[,5])

# Stop cluster
parallel::stopCluster(cl = my.cluster)

# Take dataset
extraCover<- speciesDominance
```

## 2.4 Combine
```{r}
# Combine all in one dataframe
speciesDominance<- rbind(nativeCover, intraCover, extraCover)

# Some species are duplicated
length(speciesDominance$names[(duplicated(speciesDominance$names))])
# Are they all intra and native?
intra_names<-  speciesDominance$names[speciesDominance$neophyte=="intra"]
native_names<-  speciesDominance$names[speciesDominance$neophyte=="native"]
length(intersect(intra_names, native_names))
# Take names
# intersect(intra_names, native_names)
```

We also analyse which species are significantly negative / positive when compared to the distribution of the native species. We do this for all plants.
```{r}
# Get the quantile values of all native plant species.Cutoff values at 0.01 and 0.05 chances (both upper and lower limit)
quantiles <- quantile(speciesDominance$coverMean[speciesDominance$neophyte=="native"], probs = c(0, 0.01, 0.05, 0.95, 0.99, 1))

# Assign each species to a class
speciesDominance$significantCoverMean <- as.factor(cut(speciesDominance$coverMean, breaks = quantiles, labels = c("p=0.01 negative", "p=0.05 negative", "", "p=0.05 positive", "p=0.01 positive"), right = TRUE))
```


## 2.5 Analysis
We perform a small analysis on the data by looking at the mean values for native, intra and extra and analyse whether they are significantly different from each other
```{r}
summary(nativeCover$coverMean)
summary(intraCover$coverMean)
summary(extraCover$coverMean)

speciesDominance<- speciesDominance[speciesDominance$coverMean>0,]

hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"]))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"]))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"]))

min(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"]))

model<-(aov(log10(coverMean) ~ neophyte, data= speciesDominance))
summary(model)
TukeyHSD(model)
plot(TukeyHSD(model, conf.level=.95), las = 2)
```


```{r}
# Take a look at the most dominant species
topdominant<- speciesDominance[order(speciesDominance$coverMean, decreasing=T),]
topdominant_intra<- topdominant[topdominant$neophyte=="intra",]
topdominant_extra <- topdominant[topdominant$neophyte=="extra",]
```


## 2.6 Plot
```{r}
# Violin plot
p<-ggplot(speciesDominance, mapping = aes(x= (neophyte), y = log10(coverMean), colour = neophyte)) + 
  # create violin plot with x axis no title and y 
  geom_violin()+ylab("Mean cover when present (%)") +
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European neophyte", "extra European neophyte")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 2.2))+
  scale_x_discrete(labels=  c("extra-European \n neophytes",  "intra-European \n neophytes", "native species"))+ 
  theme(axis.text.x = element_text(size=10), axis.title.y=element_text(size=12), axis.title.x = element_blank(), axis.text.y = element_text(size=10))+
  annotate("text", x = 1, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"])), label = "c", size = 3.5, vjust = -2, hjust = 0.5) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"])), label = "a", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x=1, y=2.1, label= paste("n=",length(extraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.1, label= paste("n=",length(intraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.1, label= paste("n=",length(nativeCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)
p
ggsave("Dominance New 50.jpeg", p,width = 6, height = 4)
```


# 3 FREQUENCY
## 3.1 Analysis
The data is present in speciesDominance, which facilitates the calculations
```{r}
summary(nativeCover$numberOfPlots)
summary(intraCover$numberOfPlots)
summary(extraCover$numberOfPlots)

hist((log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])))
hist(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"]))
hist(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"]))

model<-(aov(log10(numberOfPlots) ~ neophyte, data= speciesDominance))
summary(model)
TukeyHSD(model)
plot(TukeyHSD(model, conf.level=.95), las = 2)

model<-kruskal.test(log10(numberOfPlots) ~ neophyte, data= speciesDominance)
summary(model)
(dunnTest(log10(numberOfPlots) ~ neophyte, data= speciesDominance, method="bonferroni"))
```

## 3.2 Plot
```{r}
# Violin plot
# plot frequency
p<- ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte)) + 
   geom_violin() +
  ylab("log10 (number of occurences)") + xlab(NULL)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European neophyte", "extra European neophyte")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_x_discrete(labels= c("extra-European \n neophytes",  "intra-European \n neophytes", "native species")) +
  theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )  +
  theme(plot.subtitle=element_text(size=12)) + 
  annotate("text", x = 1, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"])), label = "a", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"])), label = "b", size = 3.5, vjust = -2, hjust = 0.5)+
  annotate("text", x=1, y=5.2, label= paste("n=",length(extraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=5.2, label= paste("n=",length(intraCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=5.2, label= paste("n=",length(nativeCover$names)), vjust = -0.5, hjust = 0.5, size=3.5)
p
ggsave("Frequency New 50.jpeg", p,width = 6, height = 4)
```

# 4 Export
```{r}
#write.csv(speciesDominance, "speciesDominance_new_categorization_50.csv")
```

