---
title: 'Abundance analysis based on new classification'
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

This script analyses whether neophytes are more dominant and frequent than native plants or not:

# 1 DATA
## 1.1 Preparation
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(ggplot2)
library(arsenal)
library(doParallel)
library(car)
library(FSA)
library(ggpubr)
library(flextable)
library(biostat)
```

## 1.2 Data loaded
This script analyses whether neophytes are more dominant and frequent than native plants or not.  
First, we load data of EVA
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva_ESy.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
normal = T
if(normal){
  fullPlotData <- read_csv("fullPlotData_ESy.csv")
} else {
  fullPlotData_EUNIS <- read_csv("fullPlotData_EUNIS.csv")
  habitats <- unique(fullPlotData$habitat)
  # adapt here the type you want to assess, e.g. 'grasslands'=1
  fullPlotData <- fullPlotData[fullPlotData$habitat==habitats[3] & !is.na(fullPlotData$habitat),]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}


fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=F
if(native_intra_analysis){
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
} else{
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
  species_country_status$Neophyte[species_country_status$Neophyte=="native_intra"] <- "native"
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
native <-  unique(species_country_status$species[species_country_status$Neophyte=="native"])
```


Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status[, -c(2:4,6:8)], by= c("Region"= "Region", "name"= "name"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
```


# 2 COVER
## 2.1 Synthesis
```{r}
fast=F
if(fast){
  # For speed optimization we extract only the necessary information from EVA
  reducedEva <- eva_country_neophyte |> select(PlotObservationID, Region,name, Neophyte,
                                               Layer,`Cover %`)
  
  # Add covers per layer
  reducedEva <- reducedEva |> group_by(PlotObservationID, Region,name, Neophyte, Layer) |>
    summarise(cover= sum(`Cover %`))
  
  # Fisher for multiple layers
  reducedEva <- reducedEva|> group_by(PlotObservationID, Region, name, Neophyte) |> 
    summarise(cover = (1-prod((1-ifelse(cover/100>1, 1, cover/100 ))))*100)
  
  # list unique species
  uniqueSpecies <- unique(reducedEva[, c(3,4)])
  
  # add total cover to fullPlotData
  fullPlotData <- reducedEva %>% group_by(PlotObservationID) %>% 
    summarise(totalCover = sum(cover)) %>% 
    left_join(fullPlotData, by = "PlotObservationID")
  
  # Copy the full dataset without aberrant totalCover
  fullPlotData <- fullPlotData %>% filter(totalCover <= 900)
  reducedEva <- reducedEva[reducedEva$PlotObservationID %in%
                             fullPlotData$PlotObservationID,]

  
  # Make empty dataframe
  speciesDominance <- data.frame(names = character(), coverMedian = numeric(), 
                                 coverMean = numeric(), var = numeric(), 
                                 numberOfPlots = integer(), neophyte = logical())
  
  # Prepare running in parallel
  parallel::detectCores()
  n.cores <- parallel::detectCores() - 2
  my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK")
  print(my.cluster)
  doParallel::registerDoParallel(cl = my.cluster)
  foreach::getDoParRegistered()
  foreach::getDoParWorkers()
  
  i=1
  # Take time beginning
  begin<- Sys.time()
  # Do calculation of average and median cover+ variance in dataset
 speciesDominance <-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind',
                            .packages=c("dplyr","mgcv")) %dopar% {
    # Get dataset for each unique species
      tmp <- reducedEva[reducedEva$name == uniqueSpecies$name[i]& 
                          reducedEva$Neophyte == uniqueSpecies$Neophyte[i],]
    # Make ID a factor
    tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
    # Group by plot and calculate the mean species cover for that species in all plots where
    # it is present (this step is not necessary anymore)
    tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`cover`))
    
    # We are only interested in the dominance of species occurring in at least 30 plots 
    if(length(tmp$PlotObservationID) >= 30) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present, 
    # also incorporating the variance and number of Plots
      sum <- summary(tmp$cover)
      x <- c(names = uniqueSpecies$name[i], coverMedian = sum[[3]], coverMean = sum[[4]], 
             var = var(tmp$cover),numberOfPlots =length(tmp$PlotObservationID), 
             neophyte = uniqueSpecies$Neophyte[i])
  }
}
# Ending and calculate time it took to perform the calculations
end<- Sys.time()
round(end-begin)

# Make the matrix a dataframe and change the row names
speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))

# Make the Median, Mean, Variance and Number numeric
speciesDominance[,c(2:5)]<- apply(speciesDominance[, c(2:5)], MARGIN= 2, FUN= as.numeric)

# Stop the cluster
parallel::stopCluster(cl = my.cluster)
#write_csv(speciesDominance, "speciesDominance_native_intra.csv")
} else{
  if(native_intra_analysis){
  speciesDominance <- read.csv("speciesDominance_native_intra.csv")
  }else{
  speciesDominance<- read.csv("speciesDominance.csv")
  }
}
```




We also analyse which species are significantly negative / positive when compared to the distribution of the native species. We do this for all plants.
```{r}
# Get the quantile values of all native plant species.Cutoff values at 0.01 and 0.05 chances (both upper and lower limit)
quantiles <- quantile(speciesDominance$coverMean[speciesDominance$neophyte=="native"], probs = c(0, 0.01, 0.05, 0.95, 0.99, 1))

# Assign each species to a class
speciesDominance$significantCoverMean <- as.factor(cut(speciesDominance$coverMean, breaks = quantiles, labels = c("p=0.01 negative", "p=0.05 negative", "", "p=0.05 positive", "p=0.01 positive"), right = TRUE))
```



## 2.2 Analysis
We perform a small analysis on the data by looking at the mean values for native, intra and extra and analyse whether they are significantly different from each other
```{r}
names <- speciesDominance$names

check <-vegdata::parse.taxa(names)

genus <- check[is.na(check$epi1),]

speciesDominance <- speciesDominance[!speciesDominance$names %in% genus$original,]
genus_country <- species_country_status[species_country_status$name %in% genus$original,]
#write_csv(genus_country, "genus_country.csv")
```




```{r}
speciesDominance<- speciesDominance[speciesDominance$coverMean>0,]
speciesDominance <- speciesDominance[!speciesDominance$names %in% genus$original,]

# test the distributions
hist(log10(speciesDominance$coverMean))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"]))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"]))
hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"]))
#hist(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native_intra"]))

# test the variance per group 
leveneTest(log10(speciesDominance$coverMean), speciesDominance$neophyte, location="mean")
# not normal residuals
shapiro.test(resid(aov(log10(speciesDominance$coverMean) ~ speciesDominance$neophyte))[1:5000])
hist(resid(aov(log10(speciesDominance$coverMean) ~ speciesDominance$neophyte)))
# test homoscedasticity
bartlett.test(log10(speciesDominance$coverMean), speciesDominance$neophyte)


# non parametric

# ranked test
model<-(aov(rank(log10(coverMean)) ~ neophyte, data= speciesDominance))
summary(model)
TukeyHSD(model)
plot(TukeyHSD(model, conf.level=.95), las = 2)

# Welch Anova
model<-(oneway.test((log10(coverMean)) ~ neophyte, data= speciesDominance, var.equal = F))
model
tt1 <- posthoc_anova(log10(coverMean) ~ neophyte, data= speciesDominance, method="Tukey")
#comp<- agricolae::kruskal(log10(speciesDominance$coverMean),speciesDominance$neophyte, p.adj="BH")
tt1$groups

# kruskal wallis test
KW<-kruskal.test(log10(coverMean) ~ neophyte, data = speciesDominance)
KW$p.value
pairwise.wilcox.test(log10(speciesDominance$coverMean), speciesDominance$neophyte,
                 p.adjust.method = "BH")
dunnTest(log10(coverMean) ~ neophyte, data = speciesDominance,method = "bh")
```


```{r}
# Take a look at the most dominant species
topdominant<- speciesDominance[order(speciesDominance$coverMean, decreasing=T),]
topdominant<- topdominant[1:100,]
topdominant_intra<- topdominant[topdominant$neophyte=="intra",]
topdominant_extra <- topdominant[topdominant$neophyte=="extra",]
```


## 2.3 Plot
```{r}
# Violin plot

if(!native_intra_analysis){
  my_comparisons <- list( c("extra", "intra"), c("intra", "native"), c("extra", "native") )

p<-ggplot(speciesDominance, mapping = aes(x= (neophyte), y = log10(coverMean), group= neophyte, colour = neophyte, fill= neophyte)) + 
  # create violin plot with x axis no title and y 
  geom_violin(alpha= 0.5, scale= "width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white") +
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", label.y = c(2.1, 2.3, 2.5))+
  ylab("Mean cover when present (%)") +
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European aliens", "extra European aliens")) +
  guides(color="none")+
  theme(legend.position = "none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 2.7))+
  scale_x_discrete(labels=  c("extra-European \n aliens",  "intra-European \n aliens", "native species")) +
  theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12), axis.title.x = element_blank(), 
        axis.text.y = element_text(size=10)) +
  annotate("text", x = 1, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"])), label = "a", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"])), label = "b", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"])), label = "c", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="extra")), vjust = 0, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="intra")), vjust = 0, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="native")), vjust = 0, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.0001"))))
Dominance <- p

#ggsave("Dominance New 30.jpeg", p,width = 5, height = 5)
}


if(native_intra_analysis){
  my_comparisons <- list( c("extra", "intra"), c("intra", "native"), c("extra", "native"), c("native", "native_intra"),
                          c("native_intra", "intra"), c("native_intra", "extra"))

p<-ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(coverMean), group= neophyte,  fill= neophyte, colour=neophyte)) + 
  # create violin plot with x axis no title and y 
  geom_violin(alpha= 0.3, scale= "width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white") +
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", label.y = c(2.1, 2.3, 2.5, 2.7,2.9,3.1), size= 3)+
  ylab("Mean cover when present (%)") +
  scale_colour_manual(values=c( "#004D40","#FFC107", "#1E88E5", "darkgreen"), 
                    breaks=c("extra","intra", "native",  "native_intra")) +
  scale_fill_manual(values=c( "#004D40","#FFC107", "#1E88E5", "darkgreen"), 
                    breaks=c("extra","intra", "native",  "native_intra")) +
  guides(color="none")+
  theme(legend.position = "none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 3.3))+
 scale_x_discrete(labels=  c("extra-European \n neophytes",  "intra-European \n neophytes","native species", 
                             "native species \nalien elsewhere")) +
  theme(axis.text.x = element_text(size=10), axis.title.y=element_text(size=12), axis.title.x = element_blank(), 
        axis.text.y = element_text(size=10)) +
  annotate("text", x = 1, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="extra"])), label = "a", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="intra"])), label = "b", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native"])), label = "d", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 4, y = mean(log10(speciesDominance$coverMean[speciesDominance$neophyte=="native_intra"])), label = "c", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="extra")), vjust = 0, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="intra")), vjust = 0, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="native")), vjust = 0, hjust = 0.5, size=3.5)+
  annotate("text", x=4, y=2.0, label= paste("n=",sum(speciesDominance$neophyte=="native_intra")), vjust = 0, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.0001"))))
p
Dominance <- p
}
p
```

# 3 FREQUENCY
## 3.1 Analysis
The data is present in speciesDominance, which facilitates the calculations
```{r}
# distribution
hist((log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])))
hist(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"]))
hist(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"]))
# test the variance per group 
leveneTest(log10(speciesDominance$numberOfPlots), speciesDominance$neophyte, location="mean")
# not normal residuals
shapiro.test(resid(aov(log10(speciesDominance$numberOfPlots) ~ speciesDominance$neophyte))[1:5000])
hist(resid(aov(log10(speciesDominance$numberOfPlots) ~ speciesDominance$neophyte)))
# test homoscedasticity
bartlett.test(log10(speciesDominance$numberOfPlots), speciesDominance$neophyte)


# ranked test
model<-(aov(log10(numberOfPlots) ~ neophyte, data= speciesDominance))
summary(model)
TukeyHSD(model)
plot(TukeyHSD(model, conf.level=.95), las = 2)


# kruskal wallis test
kruskal.test(log10(numberOfPlots) ~ neophyte, data = speciesDominance)
pairwise.wilcox.test(log10(speciesDominance$numberOfPlots), speciesDominance$neophyte,
                 p.adjust.method = "BH")
dunnTest(log10(coverMean) ~ neophyte, data = speciesDominance,method = "bh")
```

```{r}

# Take a look at the most dominant species
topfrequent<- speciesDominance[order(speciesDominance$numberOfPlots, decreasing=T),]
topfrequent<- topfrequent[1:100,]
topdominant_intra<- topdominant[topdominant$neophyte=="intra",]
topdominant_extra <- topdominant[topdominant$neophyte=="extra",]
```



## 3.2 Plot
```{r}
# Violin plot
# plot frequency
if(!native_intra_analysis){
p<- ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte, group=neophyte, fill= neophyte)) + 
  geom_violin(alpha= 0.5, scale= "width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white") +
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", label.y = c(5.5, 5.8, 6.1), size=3)+
  ylab(expression("log"[10]* " number of occurences")) + xlab(NULL)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European aliens", "extra European aliens")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  guides(color="none")+
  theme(legend.position = "none")+
  scale_x_discrete(labels= c("extra-European \n aliens",  "intra-European \n aliens", "native species")) +
  theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )  +
  theme(plot.subtitle=element_text(size=12)) + 
  annotate("text", x = 1, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])), label = "b", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"])), label = "c", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"])), label = "a", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="extra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="intra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="native")), vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.0001"))))
Frequency <- p
p
#ggsave("Frequency New 10.jpeg", p,width = 6, height = 4)
} else {
p<- ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte, group=neophyte, fill= neophyte)) + 
  geom_violin(alpha= 0.3, scale= "width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white") +
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", label.y = c(5.5, 5.8, 6.1, 6.4,6.7,7), size=3)+
  ylab(expression("log"[10]* " number of occurences")) + xlab(NULL)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40", "darkgreen"), 
                    name="Legend",
                    breaks=c("native", "intra","extra", "native_intra"),
                    labels=c("native species", "intra European aliens", "extra European aliens", "native species alien elsewhere")) +
  scale_fill_manual(values=c( "#004D40","#FFC107", "#1E88E5", "darkgreen"), 
                    breaks=c("extra","intra", "native",  "native_intra")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  guides(color="none")+
  theme(legend.position = "none")+
  scale_x_discrete(labels= c("extra-European \n aliens",  "intra-European \n aliens", "native species",
                             "native species \nalien elsewhere")) +
  theme(axis.text.x = element_text(size=10), axis.title.y=element_text(size=12) )  +
  theme(plot.subtitle=element_text(size=12)) + 
  annotate("text", x = 1, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="extra"])), label = "c", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="intra"])), label = "d", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native"])), label = "b", size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 4, y = mean(log10(speciesDominance$numberOfPlots[speciesDominance$neophyte=="native_intra"])), label = "a",size = 4, 
           vjust = -0.7, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="extra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="intra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="native")), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=4, y=5.2, label= paste("n=",sum(speciesDominance$neophyte=="native_intra")), vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.0001"))))
Frequency <- p
p  

}


plot <- ggarrange(Dominance,Frequency, nrow=1, ncol=2, labels= c("a","b"),font.label = list(size = 12, color))
plot

#ggsave("Dominance_frequency.jpg", plot= plot, width = 10, height = 5)

```





# 4 Analysis
## 4.1 Correlation
```{r}
cor(speciesDominance$coverMean, speciesDominance$numberOfPlots)
cor(topdominant$coverMean, topdominant$numberOfPlots)
cor(topfrequent$coverMean, topfrequent$numberOfPlots)
```


## 4.2 Species
```{r}
species <- speciesDominance[speciesDominance$significantCoverMean== "p=0.01 positive",]
```



