---
title: "Analysis of abundance and frequency"
subtitle: "based on the work of Christian Ledergerber"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---
This script analyses whether neophytes are more dominant and frequent than native plants or not:

# Start
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(ggplot2)
library(arsenal)
library(doParallel)
```

This script analyses whether neophytes are more dominant and frequent than native plants or not.
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
fullPlotData <- read_csv("fullPlotData.csv")

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
neophyteDefinitions <- read_csv("../Neophyte Assignments/UniqueTaxaEurope-2023-04-23.csv")
# Assign these names to the eva list
neophyteNames <- neophyteDefinitions$species[neophyteDefinitions$statusEurope == "neo"]
neophyteNames <- neophyteNames[neophyteNames %in% eva$species]
```


# Cover distribution
Compute which species has which cover distribution in eva:
```{r}
# For speed optimization we extract only the necessary information from EVA
reducedEva <- eva |> select(PlotObservationID, species, `Cover %`)

# Get the species from eva
uniqueSpecies <- unique(reducedEva$species)

# Make new dataframe to store in the obtained data. 
speciesDominance <- data.frame(names = character(), coverMedian = numeric(), coverMean = numeric(), var = numeric(), numberOfPlots = integer(), neophyte = logical())

parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

begin<- Sys.time()
speciesDominance<-foreach(i= 1:length(uniqueSpecies), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  # Get dataset for each unique species
  tmp <- reducedEva[reducedEva$species == uniqueSpecies[i],]
  # Make ID a factor
  tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
  # Group by plot and calculate the mean species cover for that species in all plots where it is present
  tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`Cover %`))
  # We are only interested in the dominance of species occuring in at least 50 plots
  if(length(tmp$PlotObservationID) >= 50) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present
    sum <- summary(tmp$cover)
    speciesDominance <- 
      c(names = uniqueSpecies[i], coverMedian = sum[[3]], coverMean = sum[[4]], var = var(tmp$cover), numberOfPlots = length(tmp$PlotObservationID), neophyte = (uniqueSpecies[i] %in% neophyteNames))
  }
}
end<- Sys.time()
round(end-begin)


speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))
parallel::stopCluster(cl = my.cluster)
speciesDominance$neophyte<- as.logical(speciesDominance$neophyte)
speciesDominance[,2]<- as.numeric(speciesDominance[,2])
speciesDominance[,3]<- as.numeric(speciesDominance[,3])
speciesDominance[,4]<- as.numeric(speciesDominance[,4])
speciesDominance[,5]<- as.numeric(speciesDominance[,5])
```

We also analyse which species are significantly negative / positive when compared to the distribution of the native species. We do this for all plants.
```{r}
# Get the quantile values of all native plant species.Cutoff values at 0.01 and 0.05 chances (both upper and lower limit)
quantiles <- quantile(speciesDominance$coverMean[!speciesDominance$neophyte], probs = c(0, 0.01, 0.05, 0.95, 0.99, 1))

# Assign each species to a class
speciesDominance$significantCoverMean <- as.factor(cut(speciesDominance$coverMean, breaks = quantiles, labels = c("p=0.01 negative", "p=0.05 negative", "", "p=0.05 positive", "p=0.01 positive"), right = TRUE))
```

Store the results in file for co-authors / publication.
```{r}
write_csv(speciesDominance, "speciesDominance.csv")
```


Finally, visualize the distribution of positive and negative plants for native plants and neophytes:
```{r}
speciesDominance <- read_csv("speciesDominance.csv")

ggplot(speciesDominance, mapping = aes(x = log(coverMean), after_stat(density), colour = neophyte)) + geom_freqpoly(binwidth = 0.1)

nNAT<- sum(!speciesDominance$neophyte)
nNEO<- sum(speciesDominance$neophyte)

p<-ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(coverMean), colour = neophyte)) + 
  geom_violin() + ylab("Mean cover when present (%)") + xlab(NULL)+
  scale_colour_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Legend",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 2.2))+
  scale_x_discrete(labels= c("Native", "Neophyte"))+ theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )+
  labs(subtitle= "Welch t-test   p<0.0001")+ theme(plot.subtitle=element_text(size=12))+ 
  annotate("text", x = 1, y = max(log10(speciesDominance$coverMean[!speciesDominance$neophyte])), label = "b", size = 4, vjust = -0.5, hjust = 0.5, size=4) +
  annotate("text", x = 2, y = max(log10(speciesDominance$coverMean[speciesDominance$neophyte])), label = "a", size = 4, vjust = -0.5, hjust = 0.5, size=4)+
  annotate("text", x=1, y=2.2, label= paste("n=",nNAT), vjust = -0.5, hjust = 0.5, size=4)+
  annotate("text", x=2, y=2.2, label= paste("n=",nNEO), vjust = -0.5, hjust = 0.5, size=4)
p

ggsave("Dominance.jpeg", p )

```


Assess, whether the dominance distribution of native plants and neophytes is significantly different:
```{r}
# Check variances
var.test(log10(coverMean) ~ neophyte, speciesDominance, alternative = "two.sided")
# Not significantly different (but small p values--> Welch probably better e.g. https://doi.org/10.5334/irsp.82)

# Perform Welch t Test
stat.test<-t.test(log10(coverMean) ~ neophyte, speciesDominance)
stat.test

# Kolmogorov-Smirnov test 
ks.test(log10(speciesDominance$coverMean[speciesDominance$neophyte]), log10(speciesDominance$coverMean[!speciesDominance$neophyte]))

# Wilcox test
wilcox.test(log10(coverMean) ~ neophyte, speciesDominance)

# Get the datasets for the species seperately
neophyteDominance <- speciesDominance[speciesDominance$neophyte,]
nativeDominance <- speciesDominance[!speciesDominance$neophyte,]

# Which species have the highest and lowest cover, calculated separately for natives and neophytes
nativeDominance[match(min(nativeDominance$coverMean), nativeDominance$coverMean),]
neophyteDominance[match(min(neophyteDominance$coverMean), neophyteDominance$coverMean),]

nativeDominance[match(max(nativeDominance$coverMean), nativeDominance$coverMean),]
neophyteDominance[match(max(neophyteDominance$coverMean), neophyteDominance$coverMean),]

write.table(neophyteDominance %>%
    arrange(desc(coverMean)) %>%
    slice(1:20), "clipboard" )

```

# Frequency
Analyse whether the frequency distributions of native and neophyte plants are significantly different:
```{r}
# Load in the file containing the number of plots per species
plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope.csv")
# Only keep when there are at least 10 plots
plotsPerSpecies <- plotsPerSpecies[plotsPerSpecies$numberOfPlots > 10,]
# Give neophyte names
plotsPerSpecies$neophyte <- plotsPerSpecies$names %in% neophyteNames


ggplot(plotsPerSpecies, mapping = aes(x = log(numberOfPlots), after_stat(density), colour = neophyte)) + geom_freqpoly(binwidth = .5)

nNEO<- sum(plotsPerSpecies$neophyte)
nNAT<- sum(!plotsPerSpecies$neophyte)

p<- ggplot(plotsPerSpecies, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte)) + 
   geom_violin() +
  ylab("log10(number of occurences)") + xlab(NULL)+
  scale_colour_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Legend",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_x_discrete(labels= c("Native", "Neophyte"))+theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )+
  labs(subtitle= "Welch t-test   p<0.0001")+  theme(plot.subtitle=element_text(size=12))+ 
   annotate("text", x = 1, y = max(log10(plotsPerSpecies$numberOfPlots[!plotsPerSpecies$neophyte])), label = "a", size = 4, vjust = -0.5, hjust = 0.5) +
  annotate("text", x = 2, y = max(log10(plotsPerSpecies$numberOfPlots[plotsPerSpecies$neophyte])), label = "b", size = 4, vjust = -0.5, hjust = 0.5)+
    annotate("text", x=1, y=5.5, label= paste("n=",nNAT), vjust = -0.5, hjust = 0.5, size=4)+
  annotate("text", x=2, y=5.5, label= paste("n=",nNEO), vjust = -0.5, hjust = 0.5, size=4)
p
ggsave("Frequency.jpeg", p )


summary(log10(plotsPerSpecies$numberOfPlots[!plotsPerSpecies$neophyte]))
summary(log10(plotsPerSpecies$numberOfPlots[plotsPerSpecies$neophyte]))

var.test(log10(numberOfPlots) ~ neophyte, plotsPerSpecies, alternative = "two.sided")
t.test(log10(numberOfPlots) ~ neophyte, plotsPerSpecies)
ks.test(log10(plotsPerSpecies$numberOfPlots[plotsPerSpecies$neophyte]), log10(plotsPerSpecies$numberOfPlots[!plotsPerSpecies$neophyte]))
```

