---
title: "Analysis of abundance and frequency"
subtitle: "based on the work of Christian Ledergerber"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---
This script analyses whether neophytes are more dominant and frequent than native plants or not:

# Start
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Subsequently we load the required packages.
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(ggplot2)
library(arsenal)
library(doParallel)
```

This script analyses whether neophytes are more dominant and frequent than native plants or not.
```{r, warnings=F, message=F}
# Eva data for which plots are available and after correcting for incorrect or weird plot observations
eva <- read_csv("fullPlotEva.csv")
```

Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it.
```{r, warnings=F, message=F}
# header like data, filtered and appended with EIVE and DIV
fullPlotData <- read_csv("fullPlotData.csv")

fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```

Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
neophyteDefinitions <- read_csv("../Neophyte Assignments/UniqueTaxaEurope-2023-04-23.csv")
# Assign these names to the eva list
neophyteNames <- neophyteDefinitions$species[neophyteDefinitions$statusEurope == "neo"]
neophyteNames <- neophyteNames[neophyteNames %in% eva$species]
```


# Cover distribution
Compute which species has which cover distribution in eva:
```{r}
# For speed optimization we extract only the necessary information from EVA
reducedEva <- eva |> select(PlotObservationID, species, `Cover %`)

# Get the species from eva
uniqueSpecies <- unique(reducedEva$species)

# Make new dataframe to store in the obtained data. 
speciesDominance <- data.frame(names = character(), coverMedian = numeric(), coverMean = numeric(), var = numeric(), numberOfPlots = integer(), neophyte = logical())

parallel::detectCores()
n.cores <- parallel::detectCores() - 2
my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

begin<- Sys.time()
speciesDominance<-foreach(i= 1:length(uniqueSpecies), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
  # Get dataset for each unique species
  tmp <- reducedEva[reducedEva$species == uniqueSpecies[i],]
  # Make ID a factor
  tmp$PlotObservationID <- as.factor(tmp$PlotObservationID)
  # Group by plot and calculate the mean species cover for that species in all plots where it is present
  tmp <- tmp |> group_by(PlotObservationID) |> summarise(cover = mean(`Cover %`))
  # We are only interested in the dominance of species occuring in at least 50 plots
  if(length(tmp$PlotObservationID) >= 50) {
    # Get the summary (median and mean at place 3 and 4) for when the species is present
    sum <- summary(tmp$cover)
    speciesDominance <- 
      c(names = uniqueSpecies[i], coverMedian = sum[[3]], coverMean = sum[[4]], var = var(tmp$cover), numberOfPlots = length(tmp$PlotObservationID), neophyte = (uniqueSpecies[i] %in% neophyteNames))
  }
}
end<- Sys.time()
round(end-begin)


speciesDominance<- as.data.frame(speciesDominance)
row.names(speciesDominance)<- c(1:length(speciesDominance$names))
parallel::stopCluster(cl = my.cluster)
speciesDominance$neophyte<- as.logical(speciesDominance$neophyte)
speciesDominance[,2]<- as.numeric(speciesDominance[,2])
speciesDominance[,3]<- as.numeric(speciesDominance[,3])
speciesDominance[,4]<- as.numeric(speciesDominance[,4])
speciesDominance[,5]<- as.numeric(speciesDominance[,5])
```

We also analyse which species are significantly negative / positive when compared to the distribution of the native species. We do this for all plants.
```{r}
# Get the quantile values of all native plant species.Cutoff values at 0.01 and 0.05 chances (both upper and lower limit)
quantiles <- quantile(speciesDominance$coverMean[!speciesDominance$neophyte], probs = c(0, 0.01, 0.05, 0.95, 0.99, 1))

# Assign each species to a class
speciesDominance$significantCoverMean <- as.factor(cut(speciesDominance$coverMean, breaks = quantiles, labels = c("p=0.01 negative", "p=0.05 negative", "", "p=0.05 positive", "p=0.01 positive"), right = TRUE))
```

Store the results in file for co-authors / publication.
```{r}
#write_csv(speciesDominance, "speciesDominance.csv")
```


Finally, visualize the distribution of positive and negative plants for native plants and neophytes:
```{r}
# Read full document
speciesDominance <- read_csv("speciesDominance.csv")

# compare with previously made document
speciesDominance2<-read_csv("C:/Users/u0166342/Documents/Boeren/Impact/eva_neophytes/Willem R/speciesDominance.csv")
comparedf(speciesDominance, speciesDominance2)
all.equal(speciesDominance,speciesDominance2)

# plot frequency
ggplot(speciesDominance, mapping = aes(x = log10(coverMean), after_stat(density), colour = neophyte)) + geom_freqpoly(binwidth = 0.1)

# count number of native and alien species
nNAT<- sum(!speciesDominance$neophyte)
nNEO<- sum(speciesDominance$neophyte)

# plot for their alien status the log10 of the mean cover
p<-ggplot(speciesDominance, mapping = aes(x= neophyte, y = log10(coverMean), colour = neophyte)) + 
  # create violin plot with x axis no title and y 
  geom_violin() + ylab("Mean cover when present (%)") + xlab(NULL)+
  # colour coding
  scale_colour_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Legend",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_y_continuous(breaks = c(-1, 0, 1, 2), labels = c(0.1,1,10,100), limits=c(-1, 2.2))+
  scale_x_discrete(labels= c("Native", "Neophyte"))+ theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )+
  labs(subtitle= "Welch t-test   p<0.0001")+ theme(plot.subtitle=element_text(size=12))+ 
  annotate("text", x = 1, y = max(log10(speciesDominance$coverMean[!speciesDominance$neophyte])), label = "b", size = 4, vjust = -0.5, hjust = 0.5, size=4) +
  annotate("text", x = 2, y = max(log10(speciesDominance$coverMean[speciesDominance$neophyte])), label = "a", size = 4, vjust = -0.5, hjust = 0.5, size=4)+
  annotate("text", x=1, y=2.2, label= paste("n=",nNAT), vjust = -0.5, hjust = 0.5, size=4)+
  annotate("text", x=2, y=2.2, label= paste("n=",nNEO), vjust = -0.5, hjust = 0.5, size=4)
p

#ggsave("Dominance.jpeg", p )

```


```{r}
topdominant<- speciesDominance[order(speciesDominance$coverMean, decreasing=T),]
topdominantneophytes<- topdominant[topdominant$neophyte,]
namesTDN<- topdominantneophytes$names
writeClipboard(namesTDN)
# chat GPT cultivation status (quick check, has to be improved and checked)
# Create a data frame with species names and cultivated status
# Species and cultivated status matrix (totally wrong...)
species_matrix <- matrix(c(
    "Oryza sativa", "cultivated",
    "Reynoutria x bohemica", "not cultivated",
    "Zea mays", "cultivated",
    "Ceratochloa carinata", "not cultivated",
    "Bidens aurea", "not cultivated",
    "Reynoutria sachalinensis", "not cultivated",
    "Lycium barbarum", "not cultivated",
    "Arctotheca calendula", "not cultivated",
    "Solanum tuberosum", "cultivated",
    "Sporobolus vaginiflorus", "not cultivated",
    "Rosa rugosa", "not cultivated",
    "Elaeagnus commutata", "not cultivated",
    "Reynoutria japonica", "not cultivated",
    "Carpobrotus acinaciformis", "not cultivated",
    "Pinus contorta", "not cultivated",
    "Citrus x aurantium", "cultivated",
    "Populus x canadensis", "not cultivated",
    "Baccharis halimifolia", "not cultivated",
    "Erigeron karvinskianus", "not cultivated",
    "Vicia faba", "cultivated",
    "Tradescantia fluminensis", "not cultivated",
    "Acacia saligna", "not cultivated",
    "Vaccinium macrocarpon", "not cultivated",
    "Impatiens balfourii", "not cultivated",
    "Paspalum distichum", "not cultivated",
    "Picea sitchensis", "not cultivated",
    "Larix kaempferi", "not cultivated",
    "Lemna minuta", "not cultivated",
    "Helianthus tuberosus", "not cultivated",
    "Amaranthus tamariscinus", "not cultivated",
    "Paspalum vaginatum", "not cultivated",
    "Oxalis latifolia", "not cultivated",
    "Amaranthus crispus", "not cultivated",
    "Lolium x boucheanum", "not cultivated",
    "Elodea canadensis", "not cultivated",
    "Amaranthus muricatus", "not cultivated",
    "Cyperus squarrosus", "not cultivated",
    "Populus balsamifera", "not cultivated",
    "Populus carolinensis", "not cultivated",
    "Cortaderia selloana", "not cultivated",
    "Acorus calamus", "not cultivated",
    "Rudbeckia laciniata", "not cultivated",
    "Elodea nuttallii", "not cultivated",
    "Azolla filiculoides", "not cultivated",
    "Iva xanthiifolia", "not cultivated",
    "Sporobolus indicus", "not cultivated",
    "Asclepias syriaca", "not cultivated",
    "Grindelia squarrosa", "not cultivated",
    "Elaeagnus angustifolia", "not cultivated",
    "Pinus radiata", "not cultivated",
    "Fraxinus americana", "not cultivated",
    "Epilobium brachycarpum", "not cultivated",
    "Setaria faberi", "not cultivated",
    "Mimulus guttatus", "not cultivated",
    "Impatiens glandulifera", "not cultivated",
    "Amorpha fruticosa", "not cultivated",
    "Paspalum dilatatum", "not cultivated",
    "Robinia pseudoacacia", "not cultivated",
    "Panicum dichotomiflorum", "not cultivated",
    "Claytonia perfoliata", "not cultivated",
    "Symphoricarpos albus", "not cultivated",
    "Oxalis pes-caprae", "not cultivated",
    "Panicum capillare", "not cultivated",
    "Lupinus polyphyllus", "not cultivated",
    "Bidens connata", "not cultivated",
    "Ailanthus altissima", "not cultivated",
    "Spartina x townsendii", "not cultivated",
    "Echinochloa colonum", "not cultivated",
    "Carpobrotus edulis", "not cultivated",
    "Platanus hispanica", "not cultivated",
    "Symphyotrichum lanceolatum", "not cultivated",
    "Stenotaphrum secundatum", "not cultivated",
    "Solidago canadensis", "not cultivated",
    "Pseudotsuga menziesii", "not cultivated",
    "Caragana arborescens", "not cultivated",
    "Eleusine tristachya", "not cultivated",
    "Eleusine indica", "not cultivated",
    "Ulmus pumila", "not cultivated",
    "Symphyotrichum novi-belgii aggr.", "not cultivated",
    "Solidago gigantea", "not cultivated",
    "Amaranthus viridis", "not cultivated",
    "Parthenocissus quinquefolia", "not cultivated",
    "Ceratochloa cathartica", "not cultivated",
    "Ambrosia psilostachya", "not cultivated",
    "Helianthus annuus", "not cultivated",
    "Cucurbita pepo", "not cultivated",
    "Lonicera japonica", "not cultivated",
    "Euphorbia maculata", "not cultivated",
    "Echinochloa oryzoides", "not cultivated",
    "Opuntia stricta", "not cultivated",
    "Galinsoga parviflora", "not cultivated",
    "Buddleja davidii", "not cultivated",
    "Parthenocissus inserta", "not cultivated",
    "Cotula coronopifolia", "not cultivated",
    "Cyperus eragrostis", "not cultivated",
    "Artemisia verlotiorum", "not cultivated",
    "Cornus sericea", "not cultivated",
    "Lindernia dubia", "not cultivated",
    "Sorghum halepense", "not cultivated",
    "Digitaria ciliaris", "not cultivated",
    "Amaranthus deflexus", "not cultivated",
    "Ambrosia artemisiifolia", "not cultivated",
    "Paspalum distichum subsp. distichum", "not cultivated",
    "Lepidium didymum", "not cultivated",
    "Xanthium orientale subsp. riparium", "not cultivated",
    "Amaranthus blitoides", "not cultivated",
    "Oplismenus hirtellus subsp. undulatifolius", "not cultivated",
    "Amaranthus hybridus", "not cultivated",
    "Echinocystis lobata", "not cultivated",
    "Crocosmia x crocosmiiflora", "not cultivated",
    "Gleditsia triacanthos", "not cultivated",
    "Mirabilis jalapa", "not cultivated",
    "Galinsoga quadriradiata", "not cultivated",
    "Impatiens parviflora", "not cultivated",
    "Oenothera glazioviana", "not cultivated",
    "Cenchrus longispinus", "not cultivated",
    "Fraxinus pennsylvanica", "not cultivated",
    "Panicum miliaceum", "not cultivated",
    "Prunus armeniaca", "not cultivated",
    "Oxalis", "not cultivated",
    "Dysphania ambrosioides", "not cultivated",
    "Quercus rubra", "not cultivated",
    "Juncus tenuis", "not cultivated",
    "Xanthium orientale subsp. italicum", "not cultivated",
    "Opuntia ficus-indica", "not cultivated",
    "Datura stramonium", "not cultivated",
    "Euphorbia serpens", "not cultivated",
    "Tsuga heterophylla", "not cultivated",
    "Agave americana", "not cultivated",
    "Amelanchier spicata", "not cultivated",
    "Nicotiana glauca", "not cultivated",
    "Acer negundo", "not cultivated",
    "Juglans nigra", "not cultivated",
    "Phacelia tanacetifolia", "not cultivated",
    "Amaranthus powellii", "not cultivated",
    "Pinus strobus", "not cultivated",
    "Oenothera fallax", "not cultivated",
    "Amaranthus retroflexus", "not cultivated",
    "Aronia x prunifolia", "not cultivated",
    "Bidens frondosa", "not cultivated",
    "Xanthium orientale", "not cultivated",
    "Cenchrus spinifex", "not cultivated",
    "Euphorbia prostrata", "not cultivated",
    "Erigeron floribundus", "not cultivated",
    "Symphyotrichum squamatum", "not cultivated",
    "Xanthium spinosum" , "not cultivated",
    "Symphyotrichum novi-belgii", "not cultivated",
    "Elsholtzia ciliata", "not cultivated",
    "Amaranthus albus", "cultivated",
    "Matricaria discoidea", "not cultivated",
    "Veronica peregrina", "not cultivated",
    "Solanum chenopodioides", "not cultivated",
    "Oxalis dillenii", "not cultivated",
    "Lepidium virginicum", "not cultivated",
    "Prunus serotina", "not cultivated",
    "Morus nigra", "not cultivated",
    "Oxalis corniculata", "not cultivated",
    "Senecio inaequidens", "not cultivated",
    "Epilobium brunnescens", "not cultivated",
    "Erigeron bonariensis", "not cultivated",
    "Helianthus lenticularis", "not cultivated",
    "Commelina communis", "not cultivated",
    "Potentilla indica", "not cultivated",
    "Ligustrum lucidum", "not cultivated",
    "Oenothera oakesiana", "not cultivated",
    "Erigeron sumatrensis", "not cultivated",
    "Lycopersicon esculentum", "cultivated",
    "Acalypha virginica", "not cultivated",
    "Amaranthus cruentus", "cultivated",
    "Bidens vulgata", "not cultivated",
    "Physocarpus opulifolius", "not cultivated",
    "Pittosporum tobira", "not cultivated",
    "Phytolacca americana", "not cultivated",
    "Cyperus sharonensis", "not cultivated",
    "Prunus cerasus", "not cultivated",
    "Veronica persica", "not cultivated",
    "Oenothera stucchii", "not cultivated",
    "Erigeron annuus subsp. annuus", "not cultivated",
    "Datura stramonium subsp. stramonium", "not cultivated",
    "Oenothera suaveolens", "not cultivated",
    "Euphorbia polygonifolia", "not cultivated",
    "Xanthium", "not cultivated",
    "Amelanchier lamarckii", "not cultivated",
    "Solanum linnaeanum", "not cultivated",
    "Erigeron canadensis", "not cultivated",
    "Galanthus elwesii", "not cultivated",
    "Erigeron annuus", "not cultivated",
    "Lepidium densiflorum", "not cultivated",
    "Oxalis fontana", "not cultivated",
    "Oenothera biennis", "not cultivated",
    "Morus alba", "not cultivated",
    "Cephalaria syriaca", "not cultivated",
    "Oxalis stricta", "not cultivated",
    "Cuscuta campestris", "not cultivated",
    "Mahonia aquifolium", "not cultivated",
    "Erechtites hieraciifolius", "not cultivated",
    "Oenothera parviflora", "not cultivated",
    "Erigeron annuus subsp. septentrionalis", "not cultivated",
    "Hordeum jubatum", "not cultivated",
    "Oenothera rubricaulis", "not cultivated",
    "Hemerocallis fulva", "not cultivated",
    "Epilobium ciliatum", "not cultivated",
    "Bromus danthoniae", "not cultivated",
    "Oenothera biennis aggr.", "not cultivated",
    "Epilobium ciliatum subsp. adenocaulon", "not cultivated",
    "Symphyotrichum laeve", "not cultivated",
    "Oenothera", "not cultivated",
    "Sisyrinchium septentrionale", "not cultivated"
), ncol = 2, byrow = TRUE)
species_matrix<- as.data.frame(species_matrix)
colnames(species_matrix)<- c("names","status")

# join
topdominantneophytes<- left_join(topdominantneophytes, species_matrix)

# count number of cultivated species
topdominantneophytes[topdominantneophytes$status=="cultivated",]
```


Assess, whether the dominance distribution of native plants and neophytes is significantly different:
```{r}
# Check variances
var.test(log10(coverMean) ~ neophyte, speciesDominance, alternative = "two.sided")
# Not significantly different (but small p values--> Welch probably better e.g. https://doi.org/10.5334/irsp.82)

# Perform Welch t Test --> check whether the values are significantly different from each other
# test assumes normal distribution with unequal variances --> assumed better even when equal
stat.test<-t.test(log10(coverMean) ~ neophyte, speciesDominance)
stat.test

# Kolmogorov-Smirnov test 
# non-parametrical
ks.test(log10(speciesDominance$coverMean[speciesDominance$neophyte]),log10(speciesDominance$coverMean[!speciesDominance$neophyte]))

# Wilcox test
wilcox.test(log10(coverMean) ~ neophyte, speciesDominance)

# Get the datasets for the species seperately
neophyteDominance <- speciesDominance[speciesDominance$neophyte,]
nativeDominance <- speciesDominance[!speciesDominance$neophyte,]

# Which species have the highest and lowest cover, calculated separately for natives and neophytes
nativeDominance[match(min(nativeDominance$coverMean), nativeDominance$coverMean),]
neophyteDominance[match(min(neophyteDominance$coverMean), neophyteDominance$coverMean),]

nativeDominance[match(max(nativeDominance$coverMean), nativeDominance$coverMean),]
neophyteDominance[match(max(neophyteDominance$coverMean), neophyteDominance$coverMean),]
```

# Frequency
Analyse whether the frequency distributions of native and neophyte plants are significantly different:
```{r, message=F}
# Load in the file containing the number of plots per species
plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope.csv")

plotsPerSpecies2 <- read_csv("../Intermediate Data/plotsPerSpeciesInEuropeAfterFiltering.csv")

# Only keep when there are at least 50 plots
plotsPerSpecies <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >= 50,]
# Give neophyte names
plotsPerSpecies$neophyte <- plotsPerSpecies$names %in% neophyteNames

# frequency
ggplot(plotsPerSpecies, mapping = aes(x = log10(numberOfPlots), after_stat(density), colour = neophyte)) + geom_freqpoly(binwidth = .5)

# count number of natives and neophytes
nNEO<- sum(plotsPerSpecies$neophyte)
nNAT<- sum(!plotsPerSpecies$neophyte)

# plot frequency
p<- ggplot(plotsPerSpecies, mapping = aes(x= neophyte, y = log10(numberOfPlots), colour = neophyte)) + 
   geom_violin() +
  ylab("log10(number of occurences)") + xlab(NULL)+
  scale_colour_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Legend",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes")) +
  guides(color="none")+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)+
  theme_classic()+
  scale_x_discrete(labels= c("Native", "Neophyte"))+theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=12) )+
  labs(subtitle= "Welch t-test   p<0.0145")+  theme(plot.subtitle=element_text(size=12))+ 
   annotate("text", x = 1, y = max(log10(plotsPerSpecies$numberOfPlots[!plotsPerSpecies$neophyte])), label = "a", size = 4, vjust = -0.5, hjust = 0.5) +
  annotate("text", x = 2, y = max(log10(plotsPerSpecies$numberOfPlots[plotsPerSpecies$neophyte])), label = "b", size = 4, vjust = -0.5, hjust = 0.5)+
    annotate("text", x=1, y=5.5, label= paste("n=",nNAT), vjust = -0.5, hjust = 0.5, size=4)+
  annotate("text", x=2, y=5.5, label= paste("n=",nNEO), vjust = -0.5, hjust = 0.5, size=4)
p
#ggsave("Frequency.jpeg", p )


summary(log10(plotsPerSpecies$numberOfPlots[!plotsPerSpecies$neophyte]))
summary(log10(plotsPerSpecies$numberOfPlots[plotsPerSpecies$neophyte]))

var.test(log10(numberOfPlots) ~ neophyte, plotsPerSpecies, alternative = "two.sided")
t.test(log10(numberOfPlots) ~ neophyte, plotsPerSpecies)
ks.test(log10(plotsPerSpecies$numberOfPlots[plotsPerSpecies$neophyte]), log10(plotsPerSpecies$numberOfPlots[!plotsPerSpecies$neophyte]))
```

