---
title: "Base model new"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

The purpose of this file is to explore which base model best explains the species richness.

# 1 START
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

Load packages
```{r, warnings=F, message=F}
library(readr)
library(AER)
library(mgcv)
library(modEvA)
library(lubridate)
library(visreg)
library(ggplot2)
library(parallel)
library(doParallel)
library(IDPmisc)
library(ggeffects)
library(dplyr)
```

## 1.1 Load
```{r, message=F}
# Data from header with the EIVE and DIV data
fullPlotData <- read_csv("fullPlotData_ESy.csv")
fullPlotData$Dataset <- as.factor(fullPlotData$Dataset)

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
}
```


## 1.2 Correlations
Checking for correlation among predictor variables:
```{r}
head(fullPlotData)
fullPlotData <- fullPlotData |> relocate(c("ENS0","ENS1","ENS2"), .after= numberOfVascularPlantSpecies)
# take correlation omitting the INF values in the log disturbances
correlation <- cor(NaRV.omit(fullPlotData[,c(10:53)]))

# set all values to 0 which are not highly correlated
correlation[correlation < 0.7 & correlation> -0.7] <- 0
correlation
# see correlation--> correlation --> hard to see some things within dataset
# Interesting (but logical) --> EIVEresL correlated with disturbance frequency --> higher disturbance is species living more in sunlight
# Correlations between same indicator different calculations and weighting methods --> likely
```



The following variables are all highly correlated: DistFrequency and EIVEresL. We therefore decided to only keep the overall disturbance severity and frequency in the model and remove the other correlated variables. DistFrequencyHerblayer and DistSeveretityHerblayer are only estimated for a subset of the habitats and hence should not be used for all habitats as we are doing. This is why we also removed this variable from the predictors.

Check assumptions quasipoisson
```{r}
# ENS0
mean(fullPlotData$ENS0)
var(fullPlotData$ENS0)

# ENS1
mean(fullPlotData$ENS1)
var(fullPlotData$ENS1)

# ENS1
mean(fullPlotData$ENS2)
var(fullPlotData$ENS2)
```

Check distribution
```{r}
hist(fullPlotData$ENS1)
```


# 2 MODEL
## 2.1 Distribution
We chose to include light instead of dist frequency
```{r}
model<- (ENS0 ~ log(Area) + 
                    EIVEresM.sqrt + I(EIVEresM.sqrt^2) + 
                    EIVEresN.sqrt + I(EIVEresN.sqrt^2) + 
                    EIVEresR.sqrt + I(EIVEresR.sqrt^2) + 
                    EIVEresL.sqrt + I(EIVEresL.sqrt^2) + 
                    EIVEresT.sqrt + I(EIVEresT.sqrt^2) +
                    DistSeverity + I(DistSeverity^2) +
                    Soil.Disturbance + I(Soil.Disturbance^2) +
                    Grazing.Pressure + I(Grazing.Pressure^2)+
                    Mowing.Frequency + I(Mowing.Frequency^2))
                   
# This is the simplest example possible, with normal distribution
glm1 <- glm(model, family=gaussian, fullPlotData)
glm2 <- glm(model, family= poisson, fullPlotData)
glm3 <- glm(model, family= quasipoisson, fullPlotData)


# poisson
AIC(glm1, glm2, glm3)

dispersiontest(glm2)
```


## 2.2 Method
```{r}
model<- (ENS0 ~ 
              log(Area) + 
              s(EIVEresM.sqrt, bs = 'tp') +
              s(EIVEresN.sqrt, bs = 'tp') +
              s(EIVEresR.sqrt, bs = 'tp') +
              s(EIVEresL.sqrt, bs = 'tp') +
              s(EIVEresT.sqrt, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp'))
gam0 <- bam(model, family = gaussian, data = fullPlotData,  method = 'fREML',discrete=TRUE, nthreads=4)

summary(gam0)

# GAM has a considerably lower AIC
AIC(glm2, gam0)

# GAM0 has a considerably higher Dsquared
dsquaredGam0 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam0), family = "gaussian")
dsquaredGam0
```



# 3 FORMULA
## 3.1 Database
We add database (random effect) and longitude and latitude as smoothers
```{r}
model<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM.sqrt, bs = 'tp') +
              s(EIVEresN.sqrt, bs = 'tp') +
              s(EIVEresR.sqrt, bs = 'tp') +
              s(EIVEresL.sqrt, bs = 'tp') +
              s(EIVEresT.sqrt, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
model2<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM.sqrt, bs = 'tp') +
              s(EIVEresN.sqrt, bs = 'tp') +
              s(EIVEresR.sqrt, bs = 'tp') +
              s(EIVEresL.sqrt, bs = 'tp') +
              s(EIVEresT.sqrt, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              Dataset
gam3 <- bam(model,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam4 <- bam(model2,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)

summary(gam3)
#summary(gam4)

# GAM3 has a considerably higher Dsquared
dsquaredGam3 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam3), family = "gaussian")
dsquaredGam3

# Incorporating the database considerably increases the Dsquared
# Making it a fixed factor does not increase much of the explained variation
# I do not run it here as it gives a very long output and produces NaNs
```

## 3.2 Weight EIVEs
```{r}
model1<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM.nw, bs = 'tp') +
              s(EIVEresN.nw, bs = 'tp') +
              s(EIVEresR.nw, bs = 'tp') +
              s(EIVEresL.nw, bs = 'tp') +
              s(EIVEresT.nw, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')

model2<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM.cnw, bs = 'tp') +
              s(EIVEresN.cnw, bs = 'tp') +
              s(EIVEresR.cnw, bs = 'tp') +
              s(EIVEresL.cnw, bs = 'tp') +
              s(EIVEresT.cnw, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
model3<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')

gam4 <- bam(model1,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam5 <- bam(model2,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam6 <- bam(model3,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)

# fREML is lowest for GAM with no weight
summary(gam3)
summary(gam4)
summary(gam5)
summary(gam6)

# AIC is lowest for GAM6
AIC(gam3,gam4,gam5,gam6)

# DSQUARED is highest for GAM6
dsquaredGam4 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam4), family = "gaussian")
dsquaredGam4
dsquaredGam5 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam5), family = "gaussian")
dsquaredGam5
dsquaredGam6 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam6), family = "gaussian")
dsquaredGam6
```


## 3.3 Weight DIVs
```{r}
model1<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity, bs = 'tp') +
              s(Soil.Disturbance, bs = 'tp')+
              s(Grazing.Pressure, bs = 'tp')+
              s(Mowing.Frequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
model2<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
model3<- ENS0 ~ 
              log(Area) + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(logDistSeverity, bs = 'tp') +
              s(logSoil.Disturbance, bs = 'tp')+
              s(logGrazing.Pressure, bs = 'tp')+
              s(logMowingFrequency, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
gam7 <- bam(model1,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam8 <- bam(model2,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
#gam9 <- bam(model3,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)

# GAM 8 has a lower AIC
summary(gam7)
summary(gam8)

dsquaredGam7 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam7), family = "gaussian")
dsquaredGam7
dsquaredGam8 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam8), family = "gaussian")
dsquaredGam8

# GAM8 is the overall best model
AIC(gam7,gam8)

# Check gam 6 and 8 against each other
#plot(gam6)
#plot(gam8)
```


## 3.4 Area
```{r}
model<-  ENS0 ~ 
              poly(log(Area),2) + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
model2<-  ENS0 ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')

gam9 <- bam(model,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam10 <- bam(model2,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)

# GAM 10 is the best now
summary(gam9)
summary(gam10)
summary(gam11)

dsquaredGam9 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam9), family = "gaussian")
dsquaredGam9
dsquaredGam10 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam10), family = "gaussian")
dsquaredGam10



AIC(gam11, gam9, gam10)
```

## 3.5 HFP and Climate
```{r}
model2<-  ENS0 ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')
model3<-  ENS0 ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(hfp, bs="tp")+
              s(chelsaP, bs="tp")+
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')

model4<-  ENS0 ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(hfp, bs="tp")+
              s(chelsaP, bs="tp")+
              s(elev, bs="tp")+
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')


gam12 <- bam(model2,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam13 <- bam(model3,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam14 <- bam(model4,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)

summary(gam12)
summary(gam13)
AIC(gam12, gam13, gam14)


dsquaredGam12 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam12), family = "gaussian")
dsquaredGam12
dsquaredGam13 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam13), family = "gaussian")
dsquaredGam13
dsquaredGam14 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam14), family = "gaussian")
dsquaredGam14

```


## 3.6 Final check
```{r}
gam12 <- bam(model4,family = gaussian, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam13 <- bam(model4,family = poisson, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
gam14 <- bam(model4,family = quasipoisson, data = fullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)

dsquaredGam12 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam12), family = "gaussian")
dsquaredGam12
dsquaredGam13 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam13), family = "poisson")
dsquaredGam13
dsquaredGam14 <- Dsquared(obs = fullPlotData$ENS0, pred = fitted(gam14), family = "gaussian")
dsquaredGam14

AIC(gam12, gam13)
```


# 4 PLOT
```{r}
visreg(gam13, xvar = "Area", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "EIVEresM", ylab = "Richness", xlab = "M", gg = TRUE, partial=FALSE, rug = F,  scale="response") 
visreg(gam13, xvar = "EIVEresN", ylab = "Richness", xlab = "N", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "EIVEresR", ylab = "Richness", xlab = "R", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "EIVEresL", ylab = "Richness", xlab = "L", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "EIVEresT", ylab = "Richness", xlab = "T", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "DistSeverity.sqrt", ylab = "Richness", xlab = "Distrubance Severity (SQRT)", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "Soil.Disturbance.sqrt", ylab = "Richness", xlab = "Soil Disturbance (SQRT)", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "Grazing.Pressure.sqrt", ylab = "Richness", xlab = "Grazing Pressure (SQRT)", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "Mowing.Frequency.sqrt", ylab = "Richness", xlab = "Mowing Frequency (SQRT)", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "hfp", ylab = "Richness", xlab = "hfp", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "chelsaP", ylab = "Richness", xlab = "P mean", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(gam13, xvar = "elev", ylab = "Richness", xlab = "Elevation", gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 

plot(ggpredict(gam13, terms=c("EIVEresM")))
```




