---
title: "Impact analysis per class"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

Based on the best model as determined in the baseModelExploration.qmd we assess the impact of individual neophytes in this script. We compute impact of all plants in EVA based on the response ratio between observed and expected species richness and divide this over multiple cover classes.

# 1 START
We first clean the environment.
```{r, echo=F, warning=F, message=FALSE}
rm(list=ls())
gc()
```

We load the required packages
```{r, echo=F, message=F, warning=F}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(tidyr)
library(foreach)
library(doParallel)
library(rstatix)
library(DescTools)
library(asbio)
library(lme4)
library(glmmTMB)
library(ggeffects)
```


## 1.1 Eva + header
We load the eva data with the species per plot
```{r}
eva <- read_csv("fullPlotEva_ESy.csv",show_col_types = FALSE)
```


Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it:
```{r}
fullPlotData <- read_csv("fullPlotData_ESy.csv",show_col_types = FALSE)

fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}

#eva<- read.csv("downsampledEva.csv", header=T)
#colnames(eva)[8]<- "Cover %"
#fullPlotData<- read.csv("downsampledPlotData.csv")
```


## 1.2 Neophyte
Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=F
if(native_intra_analysis){
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
} else{
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
  species_country_status$Neophyte[species_country_status$Neophyte=="native_intra"] <- "native"
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])


```


Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status[, -c(2:4,6:8)], by= c("Region"= "Region", "name"= "name"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
eva <- eva_country_neophyte

# there are still na's present --> due to being named 'plant'
nas<- eva_country_neophyte[is.na(eva_country_neophyte$Neophyte),]
eva_country_neophyte <- eva_country_neophyte[!eva_country_neophyte$name=="Plant",]
eva <- eva[!eva$name=="Plant",]
```



# 2 ANALYSIS
The idea is to compute the gam for every plot, which gives us the expected number of species based on the EIVE and DIV (and other indicators). We then compare this fitted (expected) values with the observed (real) species richness through the natural logarithm of the response ratio, which is defined as observed/ expected (commonly used in for example review papers on alien species).

Since Christians code is rather long, I try to simplify it by cutting it into more chunks.

## 2.1 Prepare
First, we make a list of the number of plots per species in our smaller dataset (if we do the analysis for all plots we can just make the file created earlier).
```{r, message=F}

if(fast) {
  reducedEva <- eva_country_neophyte |> select(PlotObservationID, Region,name, Neophyte,Layer,`Cover %`)

  reducedEva <- reducedEva |> group_by(PlotObservationID, Region,name, Neophyte, Layer) |> summarise(cover= sum(`Cover %`))
  
  reducedEva <- reducedEva|> group_by(PlotObservationID, Region, name, Neophyte) |> 
    summarise(cover = (1-prod((1-ifelse(cover/100>1, 1, cover/100 ))))*100)
  
  uniqueSpecies <- unique(reducedEva[, c(3,4)])
  
  # Empty dataset

  # Prepare parallel
  parallel::detectCores()
  n.cores <- parallel::detectCores() - 2
  my.cluster <- parallel::makeCluster(
    n.cores, 
    type = "PSOCK"
  )
  print(my.cluster)
  doParallel::registerDoParallel(cl = my.cluster)
  foreach::getDoParRegistered()
  foreach::getDoParWorkers()
  
  
  plotsPerSpecies <- data.frame(names = character(), numberOfPlots = integer(), Neophyte= character())
  # Count in how many plots a native species occurs
  begin<- Sys.time()
plotsPerSpecies<-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
    tmp <- reducedEva$PlotObservationID[reducedEva$name == uniqueSpecies$name[i]& 
                                          reducedEva$Neophyte == uniqueSpecies$Neophyte[i]]
    names=uniqueSpecies$name[i]
    numberOfPlots= length(unique(tmp))
    plotsPerSpecies[i,]<- c(names, numberOfPlots,Neophyte=uniqueSpecies$Neophyte[i] )
}

  plotsPerSpecies<- as.data.frame(plotsPerSpecies)
  plotsPerSpecies<-setNames(plotsPerSpecies, c("names", "numberOfPlots","Neophyte"))
  plotsPerSpecies$numberOfPlots <- as.numeric(plotsPerSpecies$numberOfPlots)
  row.names(plotsPerSpecies)<- c(1: nrow(uniqueSpecies))
  
  end<- Sys.time()
  round(end-begin)
  
  parallel::stopCluster(cl = my.cluster)
  #write.csv(plotsPerSpecies,"plotsPerSpeciesInEurope_native_intra.csv", row.names = F )
} else {
  if(!native_intra_analysis){
  plotsPerSpecies <- read_csv("speciesDominance.csv", show_col_types = FALSE)
  plotsPerSpecies <- plotsPerSpecies[,c(1,5:6)]
  colnames(plotsPerSpecies) <- c("names","numberOfPlots","neophyte")
  } else {
  plotsPerSpecies <- read_csv("speciesDominance_native_intra.csv", show_col_types = FALSE)
  plotsPerSpecies <- plotsPerSpecies[,c(1,5:6)]
  colnames(plotsPerSpecies) <- c("names","numberOfPlots","neophyte")
  }
}
```


Get the names of the species that are present in more than the required minimum number of plots
```{r, message=F}
minimumNumberOfPlots<- 30
taxaForCoverClassAnalysis <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >= minimumNumberOfPlots,]

fullPlotData <- fullPlotData %>% 
    select(PlotObservationID, ENS0, Area, 
           EIVEresM, EIVEresN, EIVEresR, EIVEresL, EIVEresT, 
           DistSeverity.sqrt, Soil.Disturbance.sqrt, 
           Grazing.Pressure.sqrt, Mowing.Frequency.sqrt, 
           Latitude, Longitude, Dataset, chelsaP, hfp, elev)


total_cover <- reducedEva |> group_by(PlotObservationID) |> summarise(cover= sum(cover))
fullPlotData <- left_join(fullPlotData, total_cover[, c(1,2)])
colnames(fullPlotData)[19] <- "total"


alien_cover <- reducedEva |> group_by(PlotObservationID, Neophyte) |> summarise(cover= sum(cover), n= n())
alien_cover <- left_join(alien_cover, fullPlotData[,c(1,2,19)])
alien_cover <- alien_cover |> mutate(cover_rel= cover/total, n_rel=n/ENS0, n_weighted= cover/total/n*ENS0)

test <- alien_cover |> pivot_wider(names_from = c(Neophyte),
                                   values_from = c(cover, n, cover_rel, n_rel, n_weighted))

fullPlotData <- left_join(fullPlotData, test)

fullPlotData[is.na(fullPlotData)] <- 0
```



## 2.2 Base model
Here we try the glmm approach
```{r}
# Make Dataset a factor
fullPlotData$Dataset <- as.factor(fullPlotData$Dataset)

# scale or not
scaled=T
if(scaled){
  
  # copy dataset
  fullPlotData2 <-fullPlotData
  
  # scale data 
  fullPlotData[, c(4:12, 16:18, 20:34 )] <- scale(fullPlotData[, c(4:12, 16:18, 20:34)], center=F)
  # what about the cover values? Also scale these?
}

# run model
MDL <- glmmTMB(ENS0 ~ EIVEresN+ hfp+ cover_rel_intra + cover_rel_extra+ EIVEresM+ EIVEresR+ EIVEresL+ EIVEresT+ 
                DistSeverity.sqrt+ Soil.Disturbance.sqrt+ Grazing.Pressure.sqrt+ Mowing.Frequency.sqrt+
                chelsaP+ (1| Dataset), data=fullPlotData, family= poisson(link=log))
summary(MDL)

# predict responses and plot
pr <- predict_response(MDL, "cover_rel_intra [all]")
plot(pr, show_data = TRUE, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "cover_rel_extra [all]")
plot(pr, show_data = TRUE, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "EIVEresN [all]")
plot(pr, show_data = TRUE, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "EIVEresM [all]")
plot(pr, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "EIVEresR [all]")
plot(pr, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "EIVEresT [all]")
plot(pr, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "EIVEresL [all]")
plot(pr, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "hfp [all]")
plot(pr, show_data = TRUE, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "Soil.Disturbance.sqrt [all]")
plot(pr, show_data = TRUE, show_residuals = TRUE, show_residuals_line = TRUE)
pr <- predict_response(MDL, "chelsaP [all]")
plot(pr, show_data = TRUE, show_residuals = TRUE, show_residuals_line = TRUE)


# plot using visreg
visreg::visreg(MDL, gg = TRUE, partial=FALSE, rug = F,  scale="response")
```


Compare with the gamm model
```{r}
# this is the model 
model<-  ENS0 ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(chelsaP, bs='tp')+
              s(hfp, bs='tp')+
              s(elev, bs='tp')+
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')+
              s(cover_rel_intra)+
              s(cover_rel_extra)

# run gamm
baseModel <- bam(model,family = poisson, data = fullPlotData,  
                 method = 'fREML',  discrete=TRUE, nthreads=4)
summary(baseModel)

# plot
plot(ggpredict(baseModel, terms=c("EIVEresN")))
plot(ggpredict(baseModel, terms= c("cover_rel_intra")))
plot(ggpredict(baseModel, terms= c("cover_rel_extra")))
plot(ggpredict(baseModel, terms= c("hfp")))


# use visreg for all
visreg::visreg(baseModel, scale= "response")

# try to get it for only EIVEresN
z <-visreg::visreg(baseModel, xvar = "EIVEresN", gg = TRUE, partial=T, rug = F,  scale="response", plot=F)
plot(z, ylim=c(0,50))      
points(fullPlotData$EIVEresN, fullPlotData$ENS0,  col = rgb(red=0.5, green=0, blue=0.5, 0.1))

# now only cover intra
z <-visreg::visreg(baseModel, xvar = "cover_rel_intra", gg = TRUE, partial=T, rug = F,  scale="response", plot=F)
plot(z, ylim=c(0,50)) 
points(fullPlotData$cover_rel_intra, fullPlotData$ENS0, col= rgb(red=0.5, green=0, blue=0.5, 0.1))
```


Try boosted regression tree
```{r}
library(gbm)
library(dismo)

fullPlotData2 <- as.data.frame(fullPlotData2)
BRTmodel <- gbm.step(data=fullPlotData2,
                             gbm.x = c(4:12, 27:28), 
                             gbm.y = 2, 
                             family = "poisson",
                             tree.complexity = 3,
                             learning.rate = 0.001,
                             bag.fraction = 0.7)

BRTmodel$contributions
gbm.plot(BRTmodel, n.plots=11, write.title = F,
         common.scale = F, y.label="Fitted function",
         show.contrib = TRUE)
```


```{r}
library(xgboost)
train <- fullPlotData2[runif(length(fullPlotData2$PlotObservationID)) > 0.30,]
test <- fullPlotData2[!fullPlotData2$PlotObservationID %in% train$PlotObservationID,]

bst <- xgboost(data = as.matrix(train[, c(4:12)]), label = train$ENS0, max.depth = 4, eta = 1, nthread = 1, nrounds = 20, objective = "count:poisson")

pred <- predict(bst, as.matrix(test[,c(4:12)]))
err <- mean(as.numeric(pred)- test$ENS0)
err
max(pred)
mean(pred)
max(train$ENS0)
mean(train$ENS0)

```

