---
title: "Impact analysis per class"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

Based on the best model as determined in the baseModelExploration.qmd we assess the impact of individual neophytes in this script. We compute impact of all plants in EVA based on the response ratio between observed and expected species richness and divide this over multiple cover classes.

# 1 START
We first clean the environment.
```{r, echo=F, warning=F, message=FALSE}
rm(list=ls())
gc()
```

We load the required packages
```{r, echo=F, message=F, warning=F}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(tidyr)
library(foreach)
library(doParallel)
library(rstatix)
library(DescTools)
library(asbio)
library(lme4)
library(glmmTMB)
#library(gam.hp)
library(ggeffects)
```


For complete data, load "I:/imporance.RData"


## 1.1 Eva + header
We load the eva data with the species per plot
```{r}
eva <- read_csv("fullPlotEva_ESy.csv",show_col_types = FALSE)
```


Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it:
```{r}
fullPlotData <- read_csv("fullPlotData_ESy.csv",show_col_types = FALSE)
fullPlotData$Date <- as.Date(fullPlotData$Date, format = c("%Y-%m-%d"))
fullPlotData <- fullPlotData[fullPlotData$Date > "1980-01-01",]

fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}

#eva<- read.csv("downsampledEva.csv", header=T)
#colnames(eva)[8]<- "Cover %"
#fullPlotData<- read.csv("downsampledPlotData.csv")
```


## 1.2 Neophyte
Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=F
if(native_intra_analysis){
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
} else{
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
  species_country_status$Neophyte[species_country_status$Neophyte=="native_intra"] <- "native"
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
```


Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status[, -c(2:4,6:8)], by= c("Region"= "Region", "name"= "name"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
eva <- eva_country_neophyte

# there are still na's present --> due to being named 'plant'
nas<- eva_country_neophyte[is.na(eva_country_neophyte$Neophyte),]
eva_country_neophyte <- (eva_country_neophyte[!eva_country_neophyte$name=="Plant",])
eva <- eva[!eva$name=="Plant",]
```



# 2 ANALYSIS
The idea is to compute the gam for every plot, which gives us the expected number of species based on the EIVE and DIV (and other indicators). We then compare this fitted (expected) values with the observed (real) species richness through the natural logarithm of the response ratio, which is defined as observed/ expected (commonly used in for example review papers on alien species).

Since Christians code is rather long, I try to simplify it by cutting it into more chunks.

## 2.1 Prepare
First, we make a list of the number of plots per species in our smaller dataset (if we do the analysis for all plots we can just make the file created earlier).
```{r, message=F}

# create smaller dataset
reducedEva <- eva_country_neophyte |> dplyr::select(PlotObservationID, Region,name,
                                             Neophyte,Layer,`Cover %`)

# sum cover species per plot
reducedEva <- reducedEva |> group_by(PlotObservationID, Region,name, Neophyte, Layer) |>
  summarise(cover= sum(`Cover %`))

# Fisher methode  
reducedEva <- reducedEva|> group_by(PlotObservationID, Region, name, Neophyte) |>
  summarise(cover = (1-prod((1-ifelse(cover/100>1, 1, cover/100 ))))*100)

# create vector unique species 
uniqueSpecies <- unique(reducedEva[, c(3,4)])

if(fast) {
  # Prepare parallel
  parallel::detectCores()
  n.cores <- parallel::detectCores() - 2
  my.cluster <- parallel::makeCluster(
    n.cores, 
    type = "PSOCK"
  )
  print(my.cluster)
  doParallel::registerDoParallel(cl = my.cluster)
  foreach::getDoParRegistered()
  foreach::getDoParWorkers()
  
  # empty dataset
  plotsPerSpecies <- data.frame(names = character(), numberOfPlots = integer(), 
                                Neophyte= character())
  # Count in how many plots a native species occurs
  begin<- Sys.time()
  plotsPerSpecies<-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind',
                           .packages=c("dplyr","mgcv")) %dopar% {
      tmp <- reducedEva$PlotObservationID[reducedEva$name == uniqueSpecies$name[i]& 
                                            reducedEva$Neophyte == uniqueSpecies$Neophyte[i]]
      names=uniqueSpecies$name[i]
      numberOfPlots= length(unique(tmp))
      plotsPerSpecies[i,]<- c(names, numberOfPlots,Neophyte=uniqueSpecies$Neophyte[i] )
  }

  plotsPerSpecies<- as.data.frame(plotsPerSpecies)
  plotsPerSpecies<-setNames(plotsPerSpecies, c("names", "numberOfPlots","Neophyte"))
  plotsPerSpecies$numberOfPlots <- as.numeric(plotsPerSpecies$numberOfPlots)
  row.names(plotsPerSpecies)<- c(1: nrow(uniqueSpecies))
  
  end<- Sys.time()
  round(end-begin)
  
  parallel::stopCluster(cl = my.cluster)
  #write.csv(plotsPerSpecies,"plotsPerSpeciesInEurope.csv", row.names = F )
} else {
  if(!native_intra_analysis){
  plotsPerSpecies <- read_csv("speciesDominance_1980.csv", show_col_types = FALSE)
  plotsPerSpecies <- plotsPerSpecies[,c(-1)]
  colnames(plotsPerSpecies) <- c("names","numberOfPlots","neophyte")
  } else {
  plotsPerSpecies <- read_csv("speciesDominance_native_intra.csv", show_col_types = FALSE)
  plotsPerSpecies <- plotsPerSpecies[,c(1,5:6)]
  colnames(plotsPerSpecies) <- c("names","numberOfPlots","neophyte")
  }
}
```


Get the names of the species that are present in more than the required minimum number of plots
```{r, message=F}
# downsize number of species
minimumNumberOfPlots<- 30
taxaForCoverClassAnalysis <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >=
                                               minimumNumberOfPlots,]

# downsize fullPlotData
fullPlotData <- fullPlotData %>% 
    dplyr::select(PlotObservationID, ENS0, Area, 
           EIVEresM, EIVEresN, EIVEresR, EIVEresL, EIVEresT, 
           DistSeverity.sqrt , Soil.Disturbance.sqrt , 
           Grazing.Pressure.sqrt , Mowing.Frequency.sqrt , 
           Latitude, Longitude, Dataset, chelsaP, hfp, elev)
```


Compute some metrics on alien and native species
```{r, message=F}
# compute total cover per size and add to the fullPlotData
total_cover <- reducedEva |> group_by(PlotObservationID) |> summarise(cover= sum(cover))
fullPlotData <- left_join(fullPlotData, total_cover[, c(1,2)])
colnames(fullPlotData)[19] <- "total"


# calculate the sum of cover per plot per class (alien or native) and number of species
alien_cover <- reducedEva |> group_by(PlotObservationID, Neophyte) |> 
  summarise(cover= sum(cover), n= n())

# add fullPlotData plot info and total cover
alien_cover <- left_join(alien_cover, fullPlotData[,c(1,2,19)])
# calculated weighted values
alien_cover <- alien_cover |> mutate(cover_rel= cover/total, n_rel=n/ENS0, 
                                     n_weighted= cover/total/n*ENS0,
                                     DI = (cover/total + n/ENS0)/2)

# make column per value
test <- alien_cover |> pivot_wider(names_from = c(Neophyte),
                                   values_from = c(cover, n, cover_rel, n_rel,n_weighted, DI))

# add to full plot data
fullPlotData <- left_join(fullPlotData, test)
```



## 2.2 Base model
Downsize dataset
```{r}
rel_extra <- sum(!is.na(fullPlotData$cover_rel_extra))/ length(fullPlotData$PlotObservationID)
rel_intra <- sum(!is.na(fullPlotData$cover_rel_intra))/ length(fullPlotData$PlotObservationID) 

downsize =F
if(downsize){ 
  if(rel_intra< 0.10){
    needed_size <- (sum(!is.na(fullPlotData$cover_rel_intra))/0.10)-sum(!is.na(fullPlotData$cover_rel_intra))-
      sum(!is.na(fullPlotData$cover_rel_extra)& is.na(fullPlotData$cover_rel_intra))
    data_to_sample <- fullPlotData[is.na(fullPlotData$cover_rel_intra)& is.na(fullPlotData$cover_rel_extra),]
    sample <- sample(data_to_sample$PlotObservationID, needed_size, replace=F)
    update <- data_to_sample[data_to_sample$PlotObservationID %in% sample,]
    subset_data <- rbind(fullPlotData[!is.na(fullPlotData$cover_rel_intra),],update, 
                         fullPlotData[is.na(fullPlotData$cover_rel_intra)& !is.na(fullPlotData$cover_rel_extra),])
    size<- length(subset_data$PlotObservationID)
  }
  fullPlotData <- subset_data
}

# make all NAs zero
fullPlotData[is.na(fullPlotData)] <- 0

# Make Dataset a factor
fullPlotData$Dataset <- as.factor(fullPlotData$Dataset)
```



Here we try the glmm approach
```{r}
# scale or not
scaled=T
if(scaled){
  
  # copy dataset for original values
  fullPlotData2 <-fullPlotData

  # scale data 
  fullPlotData[, c(4:12,13:14, 16:18, 20:22, 24:37)] <- 
    scale(fullPlotData[, c(4:12,13:14, 16:18,20:22, 24:37)], center=F, scale=T)
}
```




## 2.3 GAMM
Compare with the gamm model
```{r}
# this is the model 
# we take the log of the area already (cannot be in model)
fullPlotData$Area <- log(fullPlotData$Area)

# final model
model<-  n_native ~ 
              s(Area,bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(hfp, bs="tp")+
              s(chelsaP, bs="tp")+
              s(elev, bs='tp')+
              s(DI_intra) + s(DI_extra)

# run gamm
mod <- bam(model, family = nb(), data = fullPlotData,  
                 method = 'fREML',  discrete=T, nthreads=4)

sum <- summary(mod)
sum
```


```{r}
library(DHARMa)

# create simulation data
simulationOutput <- simulateResiduals(fittedModel = mod, plot = F)
# look at simulation 
# QQ plot --> line normal (is it correct distribution and is there no overdispersion)
# residual plot --> if correct line at 0.5
plot(simulationOutput)

# test dispersion
testDispersion(simulationOutput)
plotResiduals(simulationOutput)
```


```{r}
source("C:/Users/u0166342/Documents/Boeren/Extra/Gam.hp/gam.hp/R/gam.hp.R")
source("C:/Users/u0166342/Documents/Boeren/Extra/Gam.hp/gam.hp/R/creatbin.R")
source("C:/Users/u0166342/Documents/Boeren/Extra/Gam.hp/gam.hp/R/genList.R")
source("C:/Users/u0166342/Documents/Boeren/Extra/Gam.hp/gam.hp/R/plot.gamhp.R")


z <- gam.hp(mod, type="dev")
z
plot(z)
```


I made the graph here...
```{r}
var = c("EIVEresT","EIVEresN","EIVEresM","EIVEresR","EIVEresL","DistSeverity.sqrt",
        "Soil.Disturbance.sqrt","Grazing.Pressure.sqrt","Mowing.Frequency.sqrt","hfp",
        "DI_intra","DI_extra")
i=c(1:12)
plotList <- lapply(
  i ,
  FUN= function(i){
    j <- var[i]
    plot_pred <- ggeffects::predict_response(mod, j)
    plot_end <- ggplot(plot_pred, aes(x=x, y= predicted))+
                      geom_line()+
                      geom_ribbon(aes(ymin= conf.low, ymax=conf.high), alpha=0.2)+
                      xlab(j) #+
                      #geom_point(data= fullPlotData, 
                                 #aes(x= as.vector(fullPlotData[, colnames(fullPlotData)==j]), y= n_native))
    assign(paste("plot_", i,sep=""), plot_end)
    plot_end
    }
)

plot <- ggpubr::ggarrange(plotlist= plotList, ncol=3, nrow=4, labels= letters[1:12])
plot
```


## 2.4 GAMM EXTRA
Compare with the gamm model
```{r}
# this is the model 
# we take the log of the area already (cannot be in model)

model2<-  n_extra ~ 
              s(Area,bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(hfp, bs="tp")+
              s(chelsaP, bs="tp")+
              s(elev, bs='tp')

# run gamm
mod2 <- bam(model2, family= nb(), data = fullPlotData,  method = 'fREML',  discrete=T, nthreads=4)

sum2 <- summary(mod2)
sum2
```


```{r}
# create simulation data
simulationOutput <- simulateResiduals(fittedModel = mod2, plot = F)
# look at simulation 
# QQ plot --> line normal (is it correct distribution and is there no overdispersion)
# residual plot --> if correct line at 0.5
plot(simulationOutput)

# test dispersion
testDispersion(simulationOutput)
plotResiduals(simulationOutput)
```



```{r}
library(visreg)
visreg(mod2, xvar = "EIVEresN", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(mod2, xvar = "DistSeverity.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod2, xvar = "Soil.Disturbance.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod2, xvar = "Mowing.Frequency.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod2, xvar = "Grazing.Pressure.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod2, xvar = "hfp", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
```


```{r}
z_extra <- gam.hp(mod2, type="dev")
z_extra
change <- data.frame(old= test,
                     new = c("Area","EIVE M","EIVE N","EIVE R","EIVE L","EIVE T","DIV severity",
                             "DIV soil","DIV grazing","DIV mowing","HFP","MAP","Elevation"))
rownames(z_extra$hierarchical.partitioning) <- change$new
z_extra$variables <- change$new
plot(z_extra)
```


## 2.5 GAMM INTRA
```{r}
# this is the model 
# we take the log of the area already (cannot be in model)

model3<-  n_intra ~ 
              s(Area,bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(hfp, bs="tp")+
              s(chelsaP, bs="tp")+
              s(elev, bs='tp')

# run gamm
mod3 <- bam(model3, family= nb(), data = fullPlotData,  method = 'fREML',  discrete=T, nthreads=4)

sum3 <- summary(mod3)
sum3
```


```{r}
# create simulation data
simulationOutput <- simulateResiduals(fittedModel = mod3, plot = F)
# look at simulation 
# QQ plot --> line normal (is it correct distribution and is there no overdispersion)
# residual plot --> if correct line at 0.5
plot(simulationOutput)

# test dispersion
testDispersion(simulationOutput)
plotResiduals(simulationOutput)
```




```{r}
visreg(mod3, xvar = "EIVEresN", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response") 
visreg(mod3, xvar = "DistSeverity.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod3, xvar = "Soil.Disturbance.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod3, xvar = "Mowing.Frequency.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod3, xvar = "Grazing.Pressure.sqrt", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
visreg(mod3, xvar = "hfp", ylab = "Richness" ,gg = TRUE, partial=FALSE, rug = FALSE,  scale="response")
```



```{r}
z_intra <- gam.hp(mod3, type="dev")
z_intra
change <- data.frame(old= test,
                     new = c("Area","EIVE M","EIVE N","EIVE R","EIVE L","EIVE T","DIV severity",
                             "DIV soil","DIV grazing","DIV mowing","HFP","MAP","Elevation"))
rownames(z_intra$hierarchical.partitioning) <- change$new
z_intra$variables <- change$new

plot(z_intra)
```


# 3 PLOT
```{r}

library(rlist)
library(ggplot2)
library("ggsci")

setwd("I:/")
load(file = "importance.rdata") 
plot(x)

y <- x$hierarchical.partitioning
sum(y[, 4])

data <- data.frame(variable = character(), variance = numeric())
add <- data.frame(variable= "Extra", variance =sum(y[c(15),4]))
data <- add_row(data, add)


data <- data[order(data$variance, decreasing=T),]

ggplot(data, aes(x= reorder(variable, -variance), y= variance))+
  geom_bar(stat = "identity")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  NULL+
  theme_bw()+
  xlab("Variable type")+
  ylab("Explained variance (%)")

rownames(y) <- c("Area", "EIVE M", "EIVE N", "EIVE R","EIVE L", "EIVE T", "DIV severity","DIV soil","DIV grazing","DIV mowing","HFP","MAP","Elevation","Intra","Extra")

y <- as.data.frame(y)
y$name <- c("Area", "EIVE M", "EIVE N", "EIVE R","EIVE L", "EIVE T", "DIV severity","DIV soil","DIV grazing","DIV mowing","HFP","MAP","Elevation","Intra","Extra")

data <- y
data$type <- c(1,2,2,2,2,2,3,3,3,3,3,2,2,4,5)
data$type <- as.factor(data$type)
colnames(data)[4] <- c("variance")


data <- data %>% mutate(type = factor(type, 
                          levels = c("1", "2","3","4","5"),
                          labels = c("Plot", 
                                     "Environment", 
                                     "Disturbance",
                                     "Intra-European alien",
                                     "Extra-European alien")))

plot <- ggplot(data, aes(x= reorder(name, -variance), y= variance, fill=type))+
  geom_bar(stat = "identity")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  NULL+
  theme_bw()+
  xlab("Variable type")+
  ylab("Explained variance (%)")+
  scale_fill_npg(name= "Variable type")+ 
  theme(legend.position = "bottom",
          legend.box = "vertical")
plot

setwd("C:/Users/u0166342/Documents/Boeren/Impact/eva_neophytes/Neophyte-Impact")
#ggsave("variance.png", width=7.5, height=5)

```


# 4 MAP
```{r}
# get regions back to FullPlotData
fullPlotData$Region <- fullPlot2$Region[match(fullPlotData$PlotObservationID, fullPlot2$PlotObservationID)]
colnames(fullPlotData)
map <- fullPlotData[, c(1,20, 25,26,27)]
colnames(map)
#map[is.na(map)] <- 0

density <- map |> group_by(Region) |> summarise(rel_invaded = sum(!is.na(n_extra) | !is.na(n_intra))/length(PlotObservationID))
```


```{r}
library(sf)
library(sp)
library(terra)
# We load the Med regions and give correct CRS
medRegions <- read_sf("../Europe-regions-shapefiles-2023", "Emed_regions")
medRegions <- st_transform(medRegions, CRS("+proj=longlat +datum=WGS84"))
density <- left_join(medRegions, density, by= c("Region"="Region"))

rel_invaded<- ggplot()+
  geom_sf(data= density, aes(fill= rel_invaded))+
  scale_fill_viridis_c(option = "magma",begin = 0.1)+
  labs(title = "Relative number of invaded plots") +
  theme_bw()

#ggsave(rel_invaded, file="rel_invaded.png", bg="white")
```

