---
title: "Impact analysis per class"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

Based on the best model as determined in the baseModelExploration.qmd we assess the impact of individual neophytes in this script. We compute impact of all plants in EVA based on the response ratio between observed and expected species richness and divide this over multiple cover classes.

# 1 START
We first clean the environment.
```{r, echo=F, warning=F, message=FALSE}
rm(list=ls())
gc()
```

We load the required packages
```{r, echo=F, message=F, warning=F}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(tidyr)
library(foreach)
library(doParallel)
library(rstatix)
library(DescTools)
library(asbio)
library(lme4)
library(glmmTMB)
library(gam.hp)
library(ggeffects)
```


## 1.1 Eva + header
We load the eva data with the species per plot
```{r}
eva <- read_csv("fullPlotEva_ESy.csv",show_col_types = FALSE)
```


Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it:
```{r}
fullPlotData <- read_csv("fullPlotData_ESy.csv",show_col_types = FALSE)

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.999,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}

#eva<- read.csv("downsampledEva.csv", header=T)
#colnames(eva)[8]<- "Cover %"
#fullPlotData<- read.csv("downsampledPlotData.csv")
```


## 1.2 Neophyte
Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=F
if(native_intra_analysis){
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
} else{
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
  species_country_status$Neophyte[species_country_status$Neophyte=="native_intra"] <- "native"
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])


```


Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status[, -c(2:4,6:8)], by= c("Region"= "Region", "name"= "name"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
eva <- eva_country_neophyte

# there are still na's present --> due to being named 'plant'
nas<- eva_country_neophyte[is.na(eva_country_neophyte$Neophyte),]
eva_country_neophyte <- (eva_country_neophyte[!eva_country_neophyte$name=="Plant",])
eva <- eva[!eva$name=="Plant",]
```



# 2 ANALYSIS
The idea is to compute the gam for every plot, which gives us the expected number of species based on the EIVE and DIV (and other indicators). We then compare this fitted (expected) values with the observed (real) species richness through the natural logarithm of the response ratio, which is defined as observed/ expected (commonly used in for example review papers on alien species).

Since Christians code is rather long, I try to simplify it by cutting it into more chunks.

## 2.1 Prepare
First, we make a list of the number of plots per species in our smaller dataset (if we do the analysis for all plots we can just make the file created earlier).
```{r, message=F}

# create smaller dataset
reducedEva <- eva_country_neophyte |> dplyr::select(PlotObservationID, Region,name,
                                             Neophyte,Layer,`Cover %`)

# sum cover species per plot
reducedEva <- reducedEva |> group_by(PlotObservationID, Region,name, Neophyte, Layer) |>
  summarise(cover= sum(`Cover %`))

# Fisher methode  
reducedEva <- reducedEva|> group_by(PlotObservationID, Region, name, Neophyte) |>
  summarise(cover = (1-prod((1-ifelse(cover/100>1, 1, cover/100 ))))*100)

# create vector unique species 
uniqueSpecies <- unique(reducedEva[, c(3,4)])

if(fast) {
  # Prepare parallel
  parallel::detectCores()
  n.cores <- parallel::detectCores() - 2
  my.cluster <- parallel::makeCluster(
    n.cores, 
    type = "PSOCK"
  )
  print(my.cluster)
  doParallel::registerDoParallel(cl = my.cluster)
  foreach::getDoParRegistered()
  foreach::getDoParWorkers()
  
  # empty dataset
  plotsPerSpecies <- data.frame(names = character(), numberOfPlots = integer(), 
                                Neophyte= character())
  # Count in how many plots a native species occurs
  begin<- Sys.time()
  plotsPerSpecies<-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind',
                           .packages=c("dplyr","mgcv")) %dopar% {
      tmp <- reducedEva$PlotObservationID[reducedEva$name == uniqueSpecies$name[i]& 
                                            reducedEva$Neophyte == uniqueSpecies$Neophyte[i]]
      names=uniqueSpecies$name[i]
      numberOfPlots= length(unique(tmp))
      plotsPerSpecies[i,]<- c(names, numberOfPlots,Neophyte=uniqueSpecies$Neophyte[i] )
  }

  plotsPerSpecies<- as.data.frame(plotsPerSpecies)
  plotsPerSpecies<-setNames(plotsPerSpecies, c("names", "numberOfPlots","Neophyte"))
  plotsPerSpecies$numberOfPlots <- as.numeric(plotsPerSpecies$numberOfPlots)
  row.names(plotsPerSpecies)<- c(1: nrow(uniqueSpecies))
  
  end<- Sys.time()
  round(end-begin)
  
  parallel::stopCluster(cl = my.cluster)
  #write.csv(plotsPerSpecies,"plotsPerSpeciesInEurope.csv", row.names = F )
} else {
  if(!native_intra_analysis){
  plotsPerSpecies <- read_csv("speciesDominance.csv", show_col_types = FALSE)
  plotsPerSpecies <- plotsPerSpecies[,c(-1)]
  colnames(plotsPerSpecies) <- c("names","numberOfPlots","neophyte")
  } else {
  plotsPerSpecies <- read_csv("speciesDominance_native_intra.csv", show_col_types = FALSE)
  plotsPerSpecies <- plotsPerSpecies[,c(1,5:6)]
  colnames(plotsPerSpecies) <- c("names","numberOfPlots","neophyte")
  }
}
```


Get the names of the species that are present in more than the required minimum number of plots
```{r, message=F}
# downsize number of species
minimumNumberOfPlots<- 30
taxaForCoverClassAnalysis <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >=
                                               minimumNumberOfPlots,]

# downsize fullPlotData
fullPlotData <- fullPlotData %>% 
    dplyr::select(PlotObservationID, ENS0, Area, 
           EIVEresM, EIVEresN, EIVEresR, EIVEresL, EIVEresT, 
           DistSeverity.sqrt , Soil.Disturbance.sqrt , 
           Grazing.Pressure.sqrt , Mowing.Frequency.sqrt , 
           Latitude, Longitude, Dataset, chelsaP, hfp, elev)
```


Compute some metrics on alien and native species
```{r, message=F}
# compute total cover per size and add to the fullPlotData
total_cover <- reducedEva |> group_by(PlotObservationID) |> summarise(cover= sum(cover))
fullPlotData <- left_join(fullPlotData, total_cover[, c(1,2)])
colnames(fullPlotData)[19] <- "total"


# calculate the sum of cover per plot per class (alien or native) and number of species
alien_cover <- reducedEva |> group_by(PlotObservationID, Neophyte) |> 
  summarise(cover= sum(cover), n= n())

# add fullPlotData plot info and total cover
alien_cover <- left_join(alien_cover, fullPlotData[,c(1,2,19)])
# calculated weighted values
alien_cover <- alien_cover |> mutate(cover_rel= cover/total, n_rel=n/ENS0, 
                                     n_weighted= cover/total/n*ENS0,
                                     DI = (cover/total + n/ENS0)/2)

# make column per value
test <- alien_cover |> pivot_wider(names_from = c(Neophyte),
                                   values_from = c(cover, n, cover_rel, n_rel,n_weighted, DI))

# add to full plot data
fullPlotData <- left_join(fullPlotData, test)
```



## 2.2 Base model
Downsize dataset
```{r}
rel_extra <- sum(!is.na(fullPlotData$cover_rel_extra))/ length(fullPlotData$PlotObservationID)
rel_intra <- sum(!is.na(fullPlotData$cover_rel_intra))/ length(fullPlotData$PlotObservationID) 

downsize =F
if(downsize){ 
  if(rel_intra< 0.10){
    needed_size <- (sum(!is.na(fullPlotData$cover_rel_intra))/0.10)-sum(!is.na(fullPlotData$cover_rel_intra))-
      sum(!is.na(fullPlotData$cover_rel_extra)& is.na(fullPlotData$cover_rel_intra))
    data_to_sample <- fullPlotData[is.na(fullPlotData$cover_rel_intra)& is.na(fullPlotData$cover_rel_extra),]
    sample <- sample(data_to_sample$PlotObservationID, needed_size, replace=F)
    update <- data_to_sample[data_to_sample$PlotObservationID %in% sample,]
    subset_data <- rbind(fullPlotData[!is.na(fullPlotData$cover_rel_intra),],update, 
                         fullPlotData[is.na(fullPlotData$cover_rel_intra)& !is.na(fullPlotData$cover_rel_extra),])
    size<- length(subset_data$PlotObservationID)
  }
  fullPlotData <- subset_data
}

# make all NAs zero
fullPlotData[is.na(fullPlotData)] <- 0

# Make Dataset a factor
fullPlotData$Dataset <- as.factor(fullPlotData$Dataset)
```



Here we try the glmm approach
```{r}
# scale or not
scaled=T
if(scaled){
  
  # copy dataset for original values
  fullPlotData2 <-fullPlotData

  # scale data 
  fullPlotData[, c(4:12,13:14, 16:18, 20:22, 24:37)] <- 
    scale(fullPlotData[, c(4:12,13:14, 16:18,20:22, 24:37)], center=F, scale=T)
}
```


## 2.3 GLMM
```{r}
# run model

sqrt =F
if(sqrt){
  MDL <- glmmTMB(ENS0 ~ EIVEresN+ hfp+ cover_rel_intra + cover_rel_extra+ EIVEresM+ EIVEresR+
               EIVEresL+ EIVEresT+ DistSeverity + I(DistSeverity ^2)+ log(Area)+
               Soil.Disturbance + I(Soil.Disturbance ^2)+ 
               Grazing.Pressure +I(Grazing.Pressure ^2)+ Mowing.Frequency +
               I(Mowing.Frequency ^2)+ chelsaP+Longitude+Latitude+ (1| Dataset), 
               data=fullPlotData, family= poisson(link=log))
} else {
  MDL <- glmmTMB(n_native ~ EIVEresN+ hfp+ n_intra+ n_extra + EIVEresM+ EIVEresR+
               EIVEresL+ EIVEresT+ DistSeverity.sqrt + I(DistSeverity.sqrt ^2)+ log(Area)+
               Soil.Disturbance.sqrt + I(Soil.Disturbance.sqrt ^2)+ 
               Grazing.Pressure.sqrt +I(Grazing.Pressure.sqrt ^2)+ Mowing.Frequency.sqrt +
               I(Mowing.Frequency.sqrt ^2)+ chelsaP+Longitude+Latitude+ (1| Dataset), 
               data=fullPlotData, family= poisson(link=log))
}

 
# the sqrt model works better
sum<- summary(MDL)
sum
piecewiseSEM::rsquared(MDL)
```



```{r}
# plot using visreg
z<- visreg::visreg(MDL, xvar="EIVEresN", gg = TRUE, partial=FALSE, rug = F,  scale="response", plot=F)
eive_n_coef <- sum$coefficients$cond[2,1]/ z$fit$visregFit[1]

z2 <- visreg::visreg(MDL, xvar="n_extra", gg = TRUE, partial=FALSE, rug = F,  scale="response", plot=F)
extra_coef <- sum$coefficients$cond[5,1]/ z2$fit$visregFit[1]

z3 <-visreg::visreg(MDL, xvar="n_intra", gg = TRUE, partial=FALSE, rug = F,  scale="response", plot=F)
intra_coef <- sum$coefficients$cond[4,1]/ z3$fit$visregFit[1]
```



Check
```{r}
# check data
library(DHARMa)

# create simulation data
simulationOutput <- simulateResiduals(fittedModel = MDL, plot = F)
# look at simulation 
# QQ plot --> line normal (is it correct distribution and is there no overdispersion)
# residual plot --> if correct line at 0.5
plot(simulationOutput)

# test dispersion
testDispersion(simulationOutput)
plotResiduals(simulationOutput)

# are there more outliers than expected by accident
testOutliers(simulationOutput)

# compare location with expected
testQuantiles(simulationOutput)

# more zeroes than expected
testZeroInflation(simulationOutput)

# plot per random effect group
simulationOutput2 <- recalculateResiduals(simulationOutput, group = fullPlotData$Dataset)
plot(simulationOutput2, quantreg = FALSE) 
```


## 2.4 GAMM
Compare with the gamm model
```{r}
# this is the model 
model<-  n_native ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(hfp, bs="tp")+
              s(chelsaP, bs="tp")+
              s(elev, bs='tp')+
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')+
              s(DI_intra) + s(DI_extra)


model<-  n_native ~ 
              s(EIVEresN, bs = 'tp') + # factor of interest
              s(DistSeverity.sqrt, bs = 'tp') + # factor of interest
              s(Soil.Disturbance.sqrt, bs = 'tp')+ # factor of interest
              s(Grazing.Pressure.sqrt, bs = 'tp')+ # factor of interest
              s(Mowing.Frequency.sqrt, bs = 'tp') + # factor of interest
              s(hfp, bs="tp")+ # factor of interest
              s(DI_intra) + s(DI_extra) # factors of interest

# run gamm
baseModel <- bam(model, family = poisson, data = fullPlotData,  
                 method = 'fREML',  discrete=T, nthreads=4)

sum <- summary(baseModel)
sum
baseModel

gam.hp=F
if(gam.hp){
  baseModel <- gam(model, family = poisson, data = fullPlotData,  
                 method = 'REML')
  sum <- summary(baseModel)
  sum
  z <- gam.hp(baseModel, type="dev")
  z
  plot(z)
}


# use visreg for all
visreg::visreg(baseModel, scale= "response")

# try to get it for only EIVEresN
z <-visreg::visreg(baseModel, xvar = "EIVEresN", gg = TRUE, partial=T, rug = F,  scale="response", plot=F)
plot(z, ylim=c(0,50))      
points(fullPlotData$EIVEresN, fullPlotData$ENS0,  col = rgb(red=0.5, green=0, blue=0.5, 0.1))

# now only cover intra
z <-visreg::visreg(baseModel, xvar = "DI_extra", gg = TRUE, partial=T, rug = F,  scale="response", plot=F)
plot(z, ylim=c(0,50)) 
points(fullPlotData$cover_rel_intra, fullPlotData$ENS0, col= rgb(red=0.5, green=0, blue=0.5, 0.1))
```


```{r}
# this is the model I want to fit
model<-  n_native ~ 
              s(log(Area),bs='tp') + # variable between 0 and 100
              # Ecological indicator values
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') + # factor of interest
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              # Disturbance indicator values
              s(DistSeverity.sqrt, bs = 'tp') + # factor of interest
              s(Soil.Disturbance.sqrt, bs = 'tp')+ # factor of interest
              s(Grazing.Pressure.sqrt, bs = 'tp')+ # factor of interest
              s(Mowing.Frequency.sqrt, bs = 'tp') + # factor of interest
              s(hfp, bs="tp")+ # factor of interest (human footprint value Venter et al)
              s(chelsaP, bs="tp")+ # MAP
              s(elev, bs='tp')+ 
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')+
              s(DI_intra) + s(DI_extra) # factors of interest, percentage 

# all explanatory variables are scaled, except for Longitude, Latitude and Dataset

# run gamm using the BAM function (fast)
baseModel <- bam(model, family = quasipoisson, data = fullPlotData,  
                 method = 'fREML',  discrete=T, nthreads=4)

# summarise the output
summary(baseModel)

# perform gam.hp
# only works with gam, not bam
baseModel <- gam(model, family = quasipoisson, data = fullPlotData,  
                 method = 'REML')
# gam.hp
z <- gam.hp(baseModel, type="dev")
z
plot(z)
```


## 2.5 BRTs
Try boosted regression tree
```{r}
library(gbm)
library(dismo)

fullPlotData2 <- as.data.frame(fullPlotData)
BRTmodel1 <- gbm.step(data=fullPlotData2,
                             gbm.x = c(4:12, 27:28), 
                             gbm.y = 2, 
                             family = "poisson",
                             tree.complexity = 4,
                             learning.rate = 0.001,
                             bag.fraction = 0.7)

BRTmodel1$contributions
gbm.plot(BRTmodel1, n.plots=11, write.title = F,
         common.scale = F, y.label="Fitted function",
         show.contrib = TRUE)

BRTmodel1$contributions
gbm.plot(BRTmodel1, n.plots=11, write.title = F,
         common.scale = F, y.label="Fitted function",
         show.contrib = TRUE)
```


```{r}
library(xgboost)
train <- fullPlotData2[runif(length(fullPlotData2$PlotObservationID)) > 0.30,]
test <- fullPlotData2[!fullPlotData2$PlotObservationID %in% train$PlotObservationID,]

bst <- xgboost(data = as.matrix(train[, c(4:12)]), label = train$ENS0, max.depth = 4, eta = 1, nthread = 1, nrounds = 20, objective = "count:poisson")

pred <- predict(bst, as.matrix(test[,c(4:12)]))
err <- mean(as.numeric(pred)- test$ENS0)
err
max(pred)
mean(pred)
max(train$ENS0)
mean(train$ENS0)

```

