---
title: "Impact Analysis based on the residuals classified per cover class"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
params: 
  start: "NA"
  end: "NA"
---
Based on the best model as determined in the baseModelExploration.qmd we assess the impact of individual neophytes in this script.

We compute impact of all plants in EVA based on cover classes.

# Start
We first clean the environment.
```{r, echo=T}
rm(list=ls())
```

We load the required packages
```{r}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(tidyr)
```

We load the eva data with the species per plot
```{r}
eva <- read_csv("fullPlotEva.csv")
```
Load filtered Data containing the richness and indicator values for each plot. In case we want to save computing time: down sample it:
```{r}
fullPlotData <- read_csv("fullPlotData.csv")

fast <- F
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```
Load the neophyte definitions and species frequency information:
```{r}
neophyteDefinitions <- read_csv("../Neophyte Assignments/UniqueTaxaEurope-2023-04-23.csv")
neophyteNames <- neophyteDefinitions$species[neophyteDefinitions$statusEurope == "neo"]
neophyteNames <- neophyteNames[neophyteNames %in% eva$species]
```

# Impact
The idea is to compute the gam for the individual neophytes and native plants to get an estimate on whether the neophyte coincides with higher or lower plant diversity. We therefore compare the expected with the observed number of species. 

Since Christians code is rather long, I try to simplify it by cutting it into more chunks.

## Load data and analysis
First, we read the list of species which are present in at least 10 sites at the highest and lowest cover classes. This set was created in [rmarkdown link](4_Data_Preparation.Rmd).

We then make a function to compute the mean if the species is present and test this quickly.
```{r}
taxaForCoverClassAnalysis = read_csv("taxaForCoverClassAnalysis.csv")$taxa
length(taxaForCoverClassAnalysis)

# Return average cover if plant is present multiple times
computeSpeciesCover <- function(name, species, cover) {
  result <- mean(cover[species==name])
  if(is.na(result)) {
    0
  } else {
    result
  }
}

# Tests of the above function
computeSpeciesCover("nonExistentPlant", c("plant1", "plant2"), c(1,2))
computeSpeciesCover("plant1", c("plant1", "plant2"), c(1,2))
computeSpeciesCover("plant2", c("plant1", "plant2"), c(1,2))
computeSpeciesCover("plant2", c("plant1", "plant2", "plant2"), c(1,2, 3))
```


## Base model
We compute our base model with the best solution (see [rmarkdown link](6_Base_Model.Rmd)), which includes smoothers for all EIVE without weighting, smoothers for the transformed DIV, the effect of the log(area), a smoother for spatial autocorrelation and the dataset as a random effect. We use a poisson distribution with a log link function. 
```{r}
# Make Dataset a factor
fullPlotData$Dataset <- as.factor(fullPlotData$Dataset)
# GAM
baseModel <- bam(numberOfVascularPlantSpecies ~ 
              log(Area) + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(transformedDisturbanceSeverity, bs = 'tp') +
              s(transformedDisturbanceFrequency, bs = 'tp') +
              s(Dataset, bs = 're'), 
              family = poisson(link = log), 
              data = fullPlotData, 
              method = 'fREML',
              discrete=TRUE,
              nthreads=4)
```


## Residuals
```{r}
# Look at the relative residuals in the dataset
fullPlotData$relativRes <- (fullPlotData$numberOfVascularPlantSpecies - fitted(baseModel))/fitted(baseModel)
fullPlotData$logRelativRes <- (log(fullPlotData$numberOfVascularPlantSpecies) - log(fitted(baseModel)))/log(fitted(baseModel))
fullPlotData$RR<- (log(fullPlotData$numberOfVascularPlantSpecies/fitted(baseModel)))
summary(fullPlotData$relativRes)
summary(fullPlotData$logRelativRes)
summary(fullPlotData$RR)
```

## Analysis
Function calculating impact using a number of ways
```{r}
# Get Eva with plot, species and cover
reducedEva <- eva[,c("PlotObservationID", "species", "Cover %")]

# Get a vector with the name of the plant to analyse
plantName <- taxaForCoverClassAnalysis[1]

# Function to calculate the impact for each species (stored in plantName)
computePlantImpact <- function(plantName) {
  
  #Create a temporary copy of fullPlotData
  tmpFullPlotData <- fullPlotData
  
  # Get the index of all observations of the species
  indexOfPlant <- eva$species == plantName
  
  # The majority of the time is spent in computing the cover classes. 
  # We select only the rows containing the species first because this is much faster than doing it for all plots.
  # We group these observations by plot observation ID
  # We make a new variable calles coverClass for which we compute the mean cover per plot observation ID for the species under assessment.
  # We label them using breaks
  coverClassForPresent <- reducedEva[indexOfPlant,] |>
    group_by(PlotObservationID) |>
    summarise(coverClass = cut(computeSpeciesCover(plantName, species, `Cover %`), breaks = c(-1, 0, 1, 10, 50, 101), labels = 
                                 c("0%", "0%-1%", "1%-10%", "10%-50%", "50%-100%"), right = TRUE) )
  
  # Because the cover class was only computed for plots containing the species we need to assign 0% class to the remaining plots.
  tmpFullPlotData <- left_join(tmpFullPlotData, coverClassForPresent, by = "PlotObservationID")
  tmpFullPlotData$coverClass[is.na(tmpFullPlotData$coverClass)] <- "0%"
  
  # Get the mean RR per cover class
  # We subtract the value for when the species is absent (as a baseline)
  result <- tmpFullPlotData |> group_by(coverClass) |> summarise(meanRelativRes = mean(relativRes), meanLogRelativRes = mean(logRelativRes), meanRR= mean(RR))
  result$meanRelativRes <- result$meanRelativRes - result$meanRelativRes[1]
  result$meanLogRelativRes <- result$meanLogRelativRes - result$meanLogRelativRes[1]
  result$meanRR<- result$meanRR-result$meanRR[1]
  
  # Make empty dataframe to store the results. We want to have the species name, mean residuals per cover, total number of plots. The log defines the method applied
  res <- data.frame(taxa = character(), log = character(),  
                    absent = numeric(), present = numeric(), low = numeric(), med = numeric(), dominant = numeric(), 
                    numberOfPlots = integer())

  # Add the findings to the dataframe
  res <- add_row(res, taxa = plantName, log = "1",
                 absent = result$meanRelativRes[1], present = result$meanRelativRes[2], low = result$meanRelativRes[3], med = result$meanRelativRes[4], dominant = result$meanRelativRes[5],
                 numberOfPlots = sum(!(tmpFullPlotData$coverClass == "0%")))
  res <- add_row(res, taxa = plantName, log ="2",
                 absent = result$meanLogRelativRes[1], present = result$meanLogRelativRes[2], low = result$meanLogRelativRes[3], med = result$meanLogRelativRes[4], dominant = result$meanLogRelativRes[5],
                 numberOfPlots = sum(!(tmpFullPlotData$coverClass == "0%")))
  res <- add_row(res, taxa = plantName, log ="3",
                 absent = result$meanRR[1], present = result$meanRR[2], low = result$meanRR[3], med = result$meanRR[4], dominant = result$meanRR[5],
                 numberOfPlots = sum(!(tmpFullPlotData$coverClass == "0%")))
  res
}
```

Get the impact for all species
```{r}
# Get the impact
# Make new dataframe
coverClassImpactForCandidates <- data.frame(taxa = character(), log = logical(),
                                            absent = numeric(), present = numeric(), low = numeric(), med = numeric(), dominant = numeric(), 
                                            numberOfPlots = integer())

# All species which are deemed good to analyse based on criteria
start <- 1
end <- length(taxaForCoverClassAnalysis)
#if(!is.na(params$start) && params$start != "NA") { start <- params$start}
#if(!is.na(params$end) && params$end != "NA") { end <- params$end}

# Run for all species
for(i in start:end) {
  # Take previously made function and store temporarily in res
  res <- computePlantImpact(taxaForCoverClassAnalysis[i])
  # Enlarge the final dataframe
  coverClassImpactForCandidates <- rbind(coverClassImpactForCandidates, res)
  
  # If we have run for 100 species (the remainder of division by 100 is 0), we store the file already
  if(i %% 100 == 0) {
    fileName <- paste("coverClassImpactForCandidates", start, "_", i, ".csv", sep = "")
    #write_csv(coverClassImpactForCandidates, fileName)
  }
}

fileName <- paste("coverClassImpactForCandidates", start, "_", end, ".csv", sep = "")
#write_csv(coverClassImpactForCandidates, fileName)
```
## Synthesis
Create a connected box plot for all neophytes and native species to show their average impact by cover class:
```{r}
# Load the file diversity impact by cover class (equals coverClassImpactforCandidates)
fileName <- "diversityImpactByCoverClass.csv"
coverClassImpactForCandidates <- read_csv(fileName)

coverClassImpactForCandidates$neophyte<- coverClassImpactForCandidates$taxa %in% neophyteNames

# Select all columns from the dataset except number of plots and pivot this data so that we have two new columns with the cover class name as a factor and their values
dataToPlot <- coverClassImpactForCandidates |> 
  select(-numberOfPlots) |> 
  pivot_longer(cols = c(`absent`, `present`, `low`, `med`, `dominant`), names_to = "cover_class", values_to = "diversity_impact") |>
  mutate(cover_class = factor(cover_class, levels = c("absent", "present", "low", "med", "dominant")))

# Check whether the species are neophyte (gives T or F vector)
dataToPlot$neophyte <- dataToPlot$taxa %in% neophyteNames

# Summarize the data per classification (Neophyte or not), cover class and method and calculate the mean and standard deviation
summarizedData <- dataToPlot |> 
  group_by(neophyte, cover_class, log) |> 
  summarise(n=n(), mean = mean(diversity_impact), sd = sd(diversity_impact), ci= qt(p=1-0.05/2, df=n-1)*sd/sqrt(n))
```

## Plot per neophyte
```{r}
# Plot the data per neophyte with the cover class on the x axis and the mean impact on the y axis
# For the pure residuals
p<- ggplot(summarizedData[summarizedData$log=="1",], aes(x=cover_class, y=mean, group=neophyte, color=neophyte)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(0.05))
p

# Log residuals
p<- ggplot(summarizedData[summarizedData$log=="2",], aes(x=cover_class, y=mean, group=neophyte, color=neophyte)) + 
  geom_line(linewidth=1) +
  geom_point()+
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci), width=.2, position=position_dodge(0.05), linewidth=1) +
  ylab("Impact of Plant Species on Plant Diversity") + 
  xlab("Cover Class") +
  scale_x_discrete(labels = c('0% (absent)','[0%-1%]','[1% - 10%]', '[10%-50%]', '[50%-100%]')) +scale_y_continuous(limits = c(-0.06, 0.10))+
  scale_colour_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Legend",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes"))+theme_classic()+geom_hline(yintercept = 0, linetype="longdash")+ 
  theme(legend.position = "bottom")

p

# RR as percentage (but gives error on estimation... but see Mollot 2017 who also did it)
p<- ggplot(summarizedData[summarizedData$log=="3",], aes(x=cover_class, y=(exp(mean)-1)*100, group=neophyte, color=neophyte)) + 
  geom_line() +
  geom_point()
p

# RR as average effect size
p<- ggplot(summarizedData[summarizedData$log=="3",], aes(x=cover_class, y=mean, group=neophyte, color=neophyte)) + 
  geom_line(linewidth=1) +
  geom_point()+
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci), width=.2, position=position_dodge(0.05),  linewidth=1) +
  ylab("Impact of Plant Species on Plant Diversity") + 
  xlab("Cover Class") +
  scale_x_discrete(labels = c('0%\n(absent)','[0%-1%]\n(very low)','[1% - 10%]\n(low) ', '[10%-50%]\n(middle)', '[50%-100%]\n(high)')) +
  scale_colour_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Status in Europe",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes"))+ 
  theme_classic()+geom_hline(yintercept = 0, linetype="longdash")+ 
  theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=14) ,axis.title.x=element_text(size=14))+
  theme(legend.text= element_text(size=12), legend.title =element_text(size=14))+
  annotate("text", x=2, y= max(summarizedData$mean[summarizedData$log=="3" & summarizedData$cover_class=="present"] + summarizedData$ci[summarizedData$log=="3" & summarizedData$cover_class=="present"]*2), label=paste("p=",round(vl$p.value,4)))+
  annotate("text", x=3, y= max(summarizedData$mean[summarizedData$log=="3" & summarizedData$cover_class=="low"] + summarizedData$ci[summarizedData$log=="3" & summarizedData$cover_class=="present"]*2), label=paste("p=",round(l$p.value,4)))+
  annotate("text", x=4, y= max(summarizedData$mean[summarizedData$log=="3" & summarizedData$cover_class=="med"] + summarizedData$ci[summarizedData$log=="3" & summarizedData$cover_class=="present"]*2), label=paste("p=",round(m$p.value,4)))+
  annotate("text", x=5, y= max(summarizedData$mean[summarizedData$log=="3" & summarizedData$cover_class=="dominant"] + summarizedData$ci[summarizedData$log=="3" & summarizedData$cover_class=="present"]*2), label=paste("p=",round(d$p.value,4)))+
  theme(legend.position = "bottom")
p

ggsave("RR_Cover_Class.jpeg", p)

# R but as barplot
p<- ggplot(summarizedData[summarizedData$log=="3",], aes(x=cover_class, y=mean,   fill=neophyte)) + 
  geom_bar(stat="identity",position = position_dodge(0.7), width = 0.7) +
  geom_point(position=position_dodge(0.7))+
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci), width=.2, position=position_dodge(0.7)) +
  ylab("Impact of Plant Species on Plant Diversity") + 
  xlab("Cover Class") +
  scale_x_discrete(labels = c('0%\n(absent)','[0%-1%]\n(very low)','[1% - 10%]\n(low) ', '[10%-50%]\n(middle)', '[50%-100%]\n(high)')) +
  scale_fill_manual(values=c("#27cf6a", "#cf274e"), 
                    name="Status in Europe",
                    breaks=c(FALSE, TRUE),
                    labels=c("native species", "neophytes"))+ 
  theme_classic()+geom_hline(yintercept = 0, linetype="longdash") +
  theme(axis.text.x = element_text(size=12), axis.title.y=element_text(size=14) ,axis.title.x=element_text(size=14))+
  theme(legend.text= element_text(size=12), legend.title =element_text(size=14), legend.position = "bottom")
p

ggsave("RR_Cover_Class_Barplot.jpeg", p)
```

## Compare intra-class
Compare the different classes using Whelch t-test (mean), Wilcoxon test (mean) and F-test (variance):
```{r}
# Count how many neophyte observations we have summarized
sum(summarizedData$neophyte)
# Split dataset between neophyte and native species
neophyteData <- dataToPlot[dataToPlot$neophyte & dataToPlot$log=="3",]
nativeData <- dataToPlot[!dataToPlot$neophyte & dataToPlot$log=="3",]

# Whelch test
vl<-t.test(neophyteData$diversity_impact[neophyteData$cover_class == "present"], nativeData$diversity_impact[nativeData$cover_class == "present"], alternative= c("two.sided"), var.equal = T)
t.test(neophyteData$diversity_impact[neophyteData$cover_class == "present"], nativeData$diversity_impact[nativeData$cover_class == "present"], alternative= c("two.sided"), var.equal = F)
l<-t.test(neophyteData$diversity_impact[neophyteData$cover_class == "low"], nativeData$diversity_impact[nativeData$cover_class == "low"])
m<-t.test(neophyteData$diversity_impact[neophyteData$cover_class == "med"], nativeData$diversity_impact[nativeData$cover_class == "med"])
d<-t.test(neophyteData$diversity_impact[neophyteData$cover_class == "dominant"], nativeData$diversity_impact[nativeData$cover_class == "dominant"])

round(vl$p.value, 4)

# Wilcoxon
ks.test(neophyteData$diversity_impact[neophyteData$cover_class == "present"], nativeData$diversity_impact[nativeData$cover_class == "present"])
ks.test(neophyteData$diversity_impact[neophyteData$cover_class == "low"], nativeData$diversity_impact[nativeData$cover_class == "low"])
ks.test(neophyteData$diversity_impact[neophyteData$cover_class == "med"], nativeData$diversity_impact[nativeData$cover_class == "med"])
ks.test(neophyteData$diversity_impact[neophyteData$cover_class == "dominant"], nativeData$diversity_impact[nativeData$cover_class == "dominant"])

# F-test
var.test(neophyteData$diversity_impact[neophyteData$cover_class == "present"], nativeData$diversity_impact[nativeData$cover_class == "present"])
var.test(neophyteData$diversity_impact[neophyteData$cover_class == "low"], nativeData$diversity_impact[nativeData$cover_class == "low"])
var.test(neophyteData$diversity_impact[neophyteData$cover_class == "med"], nativeData$diversity_impact[nativeData$cover_class == "med"])
var.test(neophyteData$diversity_impact[neophyteData$cover_class == "dominant"], nativeData$diversity_impact[nativeData$cover_class == "dominant"])
```

Compute for how many plant species the measure was computed and how many are neophytes:
```{r}
#sum(coverClassImpactForCandidates$log=="1")
#sum(coverClassImpactForCandidates$log=="2" & coverClassImpactForCandidates$neophyte)
#sum(coverClassImpactForCandidates$log=="3" & !coverClassImpactForCandidates$neophyte)
```

## Compare inter-class
Compute the two factor ANOVA for neophyte and cover_class to assess whether neophytes are significantly worse for plant species richness than native plants.
```{r}
# Residuals
model <- aov(diversity_impact ~ neophyte + cover_class, dataToPlot[dataToPlot$log=="1",])
hist(model$residuals)
summary(model)
summary.lm(model)

# Log Residuals
model <- aov(diversity_impact ~ neophyte + cover_class, dataToPlot[dataToPlot$log=="2",])
model
hist(model$residuals)
summary(model)
summary.lm(model)

# Response ratio
model <- aov(diversity_impact ~ neophyte + cover_class, dataToPlot[dataToPlot$log=="3",])
model
hist(model$residuals, breaks = c(50))
plot(model)
summary(model)
summary.lm(model)
```

According to Francois Gillet not valid since assumptions linear model (normal distribution, homoscedasticity)
```{r}
# Interaction included
summary(aov(rank(diversity_impact) ~ neophyte * cover_class, dataToPlot[dataToPlot$log=="3",]))

# Interaction not significant, 
result<-(aov(rank(diversity_impact) ~ neophyte + cover_class, dataToPlot[dataToPlot$log=="3",]))
summary(result)
TukeyHSD(result, which= "cover_class")
```

# Dominant
```{r}
var.test(dominant ~ neophyte, coverClassImpactForCandidates[coverClassImpactForCandidates$log=="3",], alternative = "two.sided")
t.test(dominant ~ neophyte, coverClassImpactForCandidates[coverClassImpactForCandidates$log=="3",])

nativeCover <- coverClassImpactForCandidates[!coverClassImpactForCandidates$neophyte & coverClassImpactForCandidates$log=="3",]
neophyteCover <- coverClassImpactForCandidates[coverClassImpactForCandidates$neophyte & coverClassImpactForCandidates$log=="3",]
ks.test(nativeCover$dominant, neophyteCover$dominant)

nativeCover[match(min(nativeCover$dominant), nativeCover$dominant),]
neophyteCover[match(min(neophyteCover$dominant), neophyteCover$dominant),]

nativeCover[match(max(nativeCover$dominant), nativeCover$dominant),]
neophyteCover[match(max(neophyteCover$dominant), neophyteCover$dominant),]
```



## probability dominant
```{r}
quantiles <- quantile(coverClassImpactForCandidates$dominant[!coverClassImpactForCandidates$neophyte], probs = c(0, 0.01, 0.05, 0.95, 0.99, 1))

coverClassImpactForCandidates$significanceInDominantClass <- cut(coverClassImpactForCandidates$dominant, breaks = quantiles, labels = c("p=0.01 negative", "p=0.05 negative", "", "p=0.05 positive", "p=0.01 positive"), right = TRUE)

write_csv(coverClassImpactForCandidates, "diversityImpactByCoverClass.csv")
```


