---
title: "Impact with base model for new classification"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

# 1 START

## 1.1 General
```{r, echo=FALSE}
rm(list=ls())
gc()
```


Load packages
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(doParallel)
library(foreach)
library(tidyverse)
library(geometry)
library(factoextra)
library(ggpubr)
library(rstatix)
library(rcompanion)
library(ARTool)
library(brms)
library(bayesplot)
library(ggeffects)
library(emmeans)
```


Here we set some parameters that are useful in the script
```{r}
correctArea<- F
```


## 1.2 Data
Load eva and plot data
```{r, warnings=F, message=F}
eva <- read_csv("fullPlotEva_ESy.csv")
fullPlotData <- read_csv("fullPlotData_ESy.csv")
```


Downsample by a factor of 100 if fast is selected
```{r}
fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.70,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}

#eva<- read.csv("downsampledEva.csv", header=T)
#colnames(eva)[8]<- "Cover %"
#fullPlotData<- read.csv("downsampledPlotData.csv")
```


## 1.3 Neophyte
Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=F
if(native_intra_analysis){
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
} else{
  species_country_status<- read_csv("country_species_ESy.csv", show_col_types = FALSE)
  species_country_status$Neophyte[species_country_status$Neophyte=="native_intra"] <- "native"
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
```



Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status[, -c(2:4,6:8)], by= c("Region"= "Region", "name"= "name"))
# Look at how much from every type are present
table(eva_country_neophyte$Neophyte)
```




## 1.4 Number of plots
Reduce EVA to a smaller dataset to accelerate computations.    
Count the number of plots in which the species in the dataset exist.  
```{r, message=F}
eva <- eva_country_neophyte

if(fast) {
  reducedEva <- eva |> select(PlotObservationID, name, `Cover %`, 
                              Neophyte, Layer)

  uniqueSpecies <- unique(reducedEva[, c(2,4)])
  
  # Empty dataset

  # Prepare parallel
  parallel::detectCores()
  n.cores <- parallel::detectCores() - 2
  my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK")
  print(my.cluster)
  doParallel::registerDoParallel(cl = my.cluster)
  foreach::getDoParRegistered()
  foreach::getDoParWorkers()

  plotsPerSpecies <- data.frame(names = character(), 
                                numberOfPlots = integer(), 
                                Neophyte= character())
  
  # Count in how many plots a species occurs
  begin<- Sys.time()
  plotsPerSpecies<-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind',
                         .packages=c("dplyr","mgcv")) %dopar% {
    tmp <- reducedEva$PlotObservationID[reducedEva$name == uniqueSpecies$name[i] &
                                          reducedEva$Neophyte == uniqueSpecies$Neophyte[i]]
    names=uniqueSpecies$name[i]
    numberOfPlots= length(unique(tmp))
    plotsPerSpecies[i,]<- c(names, numberOfPlots,
                            Neophyte=uniqueSpecies$Neophyte[i] )
  }
  
  plotsPerSpecies<- as.data.frame(plotsPerSpecies)
  plotsPerSpecies<-setNames(plotsPerSpecies, c("names", "numberOfPlots","Neophyte"))
  plotsPerSpecies$numberOfPlots <- as.numeric(plotsPerSpecies$numberOfPlots)
  row.names(plotsPerSpecies)<- c(1: nrow(uniqueSpecies))
  
  end<- Sys.time()
  round(end-begin)
  
  parallel::stopCluster(cl = my.cluster)
  write.csv(plotsPerSpecies, "plotsPerSpeciesInEurope.csv")
} else {
  if(!native_intra_analysis){
  plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope.csv", show_col_types = FALSE)
  plotsPerSpecies<- plotsPerSpecies[, -c(1)]
  } else {
  plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope_native_intra.csv", show_col_types = FALSE)
  }
}
```


## 1.5 Select
We want all species that are present in at least 50 plots, we order this set from highest to lowest and subsequently make a function to compute the total cover of the species across all sites.
```{r}
# If you want to have a look at the distribution of the species richness per plot: hist(plotsPerSpecies$numberOfPlots[plotsPerSpecies$numberOfPlots>1000])

# determine which plant species should be computed
minimumNumberOfPlots<- 30
allPlants <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >= minimumNumberOfPlots,]
allPlants <- allPlants[order(-allPlants$numberOfPlots),]
rownames(allPlants)<- c(1: nrow(allPlants))
```



## 1.6 Total cover
Compute total cover for each plot.
```{r}
# create
reducedEva <- eva |> select(PlotObservationID, name, `Cover %`, Neophyte, Layer)

reducedEva <- reducedEva |> group_by(PlotObservationID, name, Neophyte, Layer) |> summarise(cover= sum(`Cover %`))


reducedEva_official <- reducedEva|> group_by(PlotObservationID, name, Neophyte) |> 
  summarise(cover = (1-prod((1-ifelse(cover/100>1, 1, cover/100 ))))*100)


fullPlotData <- 
  reducedEva %>% 
  group_by(PlotObservationID) %>% 
  summarise(totalCover = sum(cover)) %>% 
  left_join(fullPlotData, by = "PlotObservationID")
```


## 1.7 PCA
```{r}
head(fullPlotData)
# Perform PCA with scaled (i.e. unit variance) variables on all EIVE and DIV variables in base model + chelsa P and hfp
PCA<- prcomp(fullPlotData[, c(26:31, 33:35, 51,53)  ], scale=T)
  

# get amount of variance explained by top 4 axes
eig.val <- get_eigenvalue(PCA)
eig.val<- eig.val[4,3]
  
# get coordinates of first 4 axes and bind to the dataset
PCA<- PCA$x[,1:4]
dim(PCA)
fullPlotData <- cbind(fullPlotData, PCA)
```



# 2 ANALYSIS
## 2.1 Function 1
```{r, meassage=F}
# The name of the first species (to test the function)
plantName = allPlants$names[1]
plantStatus= allPlants$Neophyte[1]

# Function to compute the impact for each species, so we only need the plant name as input
computePlantImpact1 <- function(plantName, plantStatus) {
  
  # Copy dataset temporarily to be able to make changes
  tmpFullPlotData <- fullPlotData
  
  reducedEva <- reducedEva_official
  # Select only explanatory variables
  tmpFullPlotData <- tmpFullPlotData %>% 
    select(numberOfVascularPlantSpecies,PlotObservationID, Area, 
           EIVEresM, EIVEresN, EIVEresR, EIVEresL, EIVEresT, 
           DistSeverity.sqrt, Soil.Disturbance.sqrt, 
           Grazing.Pressure.sqrt, Mowing.Frequency.sqrt, 
           Latitude, Longitude, Dataset, totalCover, chelsaP, hfp, PC1,PC2,PC3,PC4)
  
  # Copy the full dataset without aberrant totalCover
  tmpFullPlotData <- tmpFullPlotData %>% filter(totalCover <= 900)
  reducedEva <- reducedEva[reducedEva$PlotObservationID %in% tmpFullPlotData$PlotObservationID,]

  # Get the indices of the plant in the eva dataset
  indexOfPlant <- reducedEva$name == plantName & reducedEva$Neophyte== plantStatus
  
  # We select only the rows containing the species first because this is much faster than doing it for all plots.
  # We make the dataset smaller (only those plots where the species is present), group the data per plot and compute the total cover     (all species) and cover of the species we are highlighting now
  plotsWherePlantOccurs <- reducedEva[indexOfPlant,c(1,4)]
  colnames(plotsWherePlantOccurs)[2]<- "speciesCover"
  
  # Join the computed cover plots with the full data
  tmpFullPlotData <- left_join(tmpFullPlotData, plotsWherePlantOccurs, 
                             by = "PlotObservationID")
  
  # Identify all plots where the species investigated occurs
  tmpFullPlotData$plantOccurs = tmpFullPlotData$PlotObservationID %in% reducedEva$PlotObservationID[indexOfPlant]
  
  
  # reduce the dataset to obtain only those plots in which the species is present
  species_plots <- tmpFullPlotData[tmpFullPlotData$plantOccurs, c(19:22)  ]
  species_plot_number <- nrow(species_plots)
  
  # make matrix and reduce duplicated columns
  species_plots<- as.matrix(species_plots)
  duplicated<- duplicated(species_plots)
  species_plots<- species_plots[!duplicated, ]
  
  # take ConvexHull around the PCA coordinates in which the species is present
  convex_hull <- convhulln((species_plots))
  # Extract all points from the dataset that are additionally within this ConvexHull and do not contain the species
  is_within_hull <- inhulln(convex_hull, as.matrix(tmpFullPlotData[, c(19:22)]))
  subset_data <- tmpFullPlotData[is_within_hull, ]
  
  
  # Take size of subset data
  size<- length(subset_data$plantOccurs)
  # change name tmpFullPlotData
  tmpFullPlotData <- subset_data
  
  sum(tmpFullPlotData$plantOccurs)
  
  # If species is not present:
  # Make 1 total cover and 0 species cover
  tmpFullPlotData$totalCover[is.na(tmpFullPlotData$totalCover)] <- 100.0
  tmpFullPlotData$speciesCover[is.na(tmpFullPlotData$speciesCover)] <- 0.0
  
  # Exclude all plots which have 0 total cover
  tmpFullPlotData <- tmpFullPlotData[tmpFullPlotData$totalCover >0,] 
  
  # Do SR - 1 for all plots in which the species is present
  tmpFullPlotData$numberOfVascularPlantSpecies <- tmpFullPlotData$numberOfVascularPlantSpecies - tmpFullPlotData$plantOccurs
  
  correctArea= F
  # Correct area: if true we calculate the area that is reserved for the other plant species. 
  if(correctArea) {
    tmpFullPlotData$Area <- tmpFullPlotData$Area*(1.0 - tmpFullPlotData$speciesCover/tmpFullPlotData$totalCover)
  }

  # There is a numerical problem in the bam function if we provide empty plots...
  tmpFullPlotData <- tmpFullPlotData |> filter(numberOfVascularPlantSpecies > 0, Area > 0)
  
  # Make dataset a factor to add it as a random variable
  tmpFullPlotData$Dataset <- as.factor(tmpFullPlotData$Dataset)
  
  # Add relative species cover
  tmpFullPlotData <- tmpFullPlotData %>% 
    mutate(relSpeciesCover = 100 * speciesCover / totalCover)
  
  
  # Species occurrence as factor (not logical)
  tmpFullPlotData$plantOccursF <- factor(tmpFullPlotData$plantOccurs)
  
  # Base model
  model<-  numberOfVascularPlantSpecies ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(hfp, bs='tp')+
              s(chelsaP, bs='tp')+
              s(Dataset, bs = 're')+
              plantOccursF

  base <- bam(model,family = poisson, 
              data = tmpFullPlotData,  method = 'fREML',  discrete=TRUE, 
              nthreads=4)
  
  # Get summary from model
  baseModel <- summary(base)
  baseModel
  
  # Make a dataframe to add the residuals to
  res <- data.frame(taxa = character(), Estimate = numeric(), 
                    StdErr = numeric(), `zValue` = numeric(), 
                    pr = numeric(), numberOfPlots = integer(), 
                    Neophyte= character(),
                    RelDiff= numeric(),
                    size= numeric(),
                    species_plots= numeric(),
                    rel= numeric())
  
  res <- add_row(res, 
                 taxa = plantName, Estimate = baseModel$p.table[2,1], 
                 StdErr = baseModel$p.table[2,2], 
                 zValue = baseModel$p.table[2,3], 
                 pr = baseModel$p.table[2,4], 
                 numberOfPlots = length(plotsWherePlantOccurs$PlotObservationID), 
                 Neophyte= plantStatus,
                 RelDiff= 100* 
                   (mean(fitted(base)[tmpFullPlotData$plantOccursF == TRUE]) -
                    mean(fitted(base)[tmpFullPlotData$plantOccursF == FALSE])) /
                    mean(fitted(base)[tmpFullPlotData$plantOccursF == FALSE]),
                 size= size, 
                 species_plots = species_plot_number,
                 rel =species_plot_number/size)
  
  res
}
```



## 2.2 Function 2
```{r, meassage=F}
# The name of the first species (to test the function)
plantName = allPlants$names[661]
plantStatus= allPlants$Neophyte[661]

which(allPlants$names=="Viola alba")

# Function to compute the impact for each species, so we only need the plant name as input
computePlantImpact2 <- function(plantName, plantStatus) {
  
  
  # Copy the full dataset without aberrant totalCover
  tmpFullPlotData <- fullPlotData %>% filter(totalCover <=900)
  reducedEva <- reducedEva_official
  reducedEva <- reducedEva[reducedEva$PlotObservationID %in% tmpFullPlotData$PlotObservationID,]
  
  # Select only explanatory variables
  tmpFullPlotData <- tmpFullPlotData %>% 
    select(PlotObservationID, numberOfVascularPlantSpecies, Area, 
           EIVEresM, EIVEresN, EIVEresR, EIVEresL, EIVEresT, 
           DistSeverity.sqrt, Soil.Disturbance.sqrt, 
           Grazing.Pressure.sqrt, Mowing.Frequency.sqrt, 
           Latitude, Longitude, Dataset, chelsaP, hfp, totalCover, PC1, PC2, PC3,PC4)
  
  # Get the indices of the plant in the eva dataset
  indexOfPlant <- reducedEva$name == plantName & reducedEva$Neophyte== plantStatus
  any(indexOfPlant)
  # We select only the rows containing the species first because this is much faster than doing it for all plots.
  # We make the dataset smaller (only those plots where the species is present), group the data per plot and 
  # compute the cover of the species we are highlighting now
  plotsWherePlantOccurs <- reducedEva[indexOfPlant,c(1,4)]
  colnames(plotsWherePlantOccurs)[2]<- "speciesCover"
  
  
  # Join the computed cover plots with the full data
  tmpFullPlotData <- left_join(tmpFullPlotData, plotsWherePlantOccurs, by = "PlotObservationID")
  # Identify all plots where the species investigated occurs
  tmpFullPlotData$plantOccurs = tmpFullPlotData$PlotObservationID %in% reducedEva$PlotObservationID[indexOfPlant]

  # reduce the dataset to obtain only those plots in which the species is present
  species_plots <- tmpFullPlotData[tmpFullPlotData$plantOccurs, c(19:22)]
  # make matrix and reduce duplicated columns
  species_plots<- as.matrix(species_plots)
  duplicated<- duplicated(species_plots)
  species_plots<- species_plots[!duplicated,]
  
  # take ConvexHull around the PCA coordinates in which the species is present
  convex_hull <- convhulln((species_plots))

  # Extract all points from the dataset that are additionally within this ConvexHull and do not contain the species
  is_within_hull <- inhulln(convex_hull, as.matrix(tmpFullPlotData[, c(19:22)]))
  subset_data <- tmpFullPlotData[is_within_hull, ]
  
  # Take size of subset data
  size<- length(subset_data$plantOccurs)
  species_plot_number <- sum(subset_data$plantOccurs)

  
  # check relative size difference datasets
  rel = species_plot_number/size
  
  if(rel< 0.10){
    needed_size <- (species_plot_number/0.10)-species_plot_number
    data_to_sample <- subset_data[!subset_data$plantOccurs,]
    sample <- sample(data_to_sample$PlotObservationID, needed_size, replace=F)
    update <- data_to_sample[data_to_sample$PlotObservationID %in% sample,]
    subset_data <- rbind(subset_data[subset_data$plantOccurs,], update)
    size<- length(subset_data$plantOccurs)
  }
   
  if(rel> 0.30){
    needed_size <- round((species_plot_number/0.30)-size)
    data_to_sample <- tmpFullPlotData[!tmpFullPlotData$PlotObservationID %in% subset_data$PlotObservationID,]
    sample <- sample(data_to_sample$PlotObservationID, needed_size, replace=F)
    update <- data_to_sample[data_to_sample$PlotObservationID %in% sample,]
    subset_data <- rbind(subset_data, update)
    size<- length(subset_data$plantOccurs)
  }  
  
  if(size < 100){
    needed_size <- 100-size
    data_to_sample <- tmpFullPlotData[!tmpFullPlotData$PlotObservationID %in% subset_data$PlotObservationID,]
    sample <- sample(data_to_sample$PlotObservationID, needed_size, replace=F)
    update <- data_to_sample[data_to_sample$PlotObservationID %in% sample,]
    subset_data <- rbind(subset_data, update)
    size<- length(subset_data$plantOccurs)
  }
  
  
     
  # change data 
  tmpFullPlotData <- subset_data

  # If species is not present:
  tmpFullPlotData$speciesCover[is.na(tmpFullPlotData$speciesCover)] <- 0.0
  tmpFullPlotData$totalCover[is.na(tmpFullPlotData$speciesCover)] <- 100.0

  # Exclude all plots which have 0 total cover
  tmpFullPlotData <- tmpFullPlotData[tmpFullPlotData$totalCover >= 1,] 
  
  # Do SR - 1 for all plots in which the species is present
  tmpFullPlotData$numberOfVascularPlantSpecies <- tmpFullPlotData$numberOfVascularPlantSpecies - tmpFullPlotData$plantOccurs
  
  # Correct area: if true we calculate the area that is reserved for the other plant species. 
  if(correctArea) {
    tmpFullPlotData$Area <- tmpFullPlotData$Area*(1.0 - tmpFullPlotData$speciesCover/tmpFullPlotData$totalCover)
  }
  
  # There is a numerical problem in the bam function if we provide empty plots...
  tmpFullPlotData <- tmpFullPlotData %>% filter(numberOfVascularPlantSpecies > 0, Area > 0)
  
  # Make dataset a factor to add it as a random variable
  tmpFullPlotData$Dataset <- as.factor(tmpFullPlotData$Dataset)

  # Add relative species cover
  tmpFullPlotData <- tmpFullPlotData %>% 
    mutate(relSpeciesCover = 100 * speciesCover / totalCover) |>
    mutate(coverClass = cut(relSpeciesCover,breaks = c(-1, 0, 1, 10, 30, 50, 70, 101), 
                            labels = c("0%", "0%-1%", "1%-10%", "10%-30%", "30%-50%","50%-70%","70%-100%")))
  
  tmpFullPlotData$coverClass<- as.character( tmpFullPlotData$coverClass)
  tmpFullPlotData$coverClass[(tmpFullPlotData$speciesCover==0.0)] <- c("0%")
  
  # Species occurrence as factor (not logical)
  tmpFullPlotData$plantOccursF <- factor(tmpFullPlotData$plantOccurs)
  tmpFullPlotData$coverClass <- factor(tmpFullPlotData$coverClass)
  
  # Base model
  model<-  (numberOfVascularPlantSpecies) ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(chelsaP, bs='tp')+
              s(hfp, bs='tp')+
              s(Dataset, bs="re")+
              coverClass
  
  
  base <- bam(model,family = poisson(link=log), data = tmpFullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
  baseModel<- summary(base)
  baseModel
  
  # Make a dataframe to add the residuals to
  res <- data.frame(taxa = character(), Estimate = numeric(),
                    StdErr = numeric(),
                    `zValue` = numeric(),
                    pr = numeric(),
                    Intercept= numeric(), 
                    numberOfPlots = integer(), 
                    Neophyte= character(),
                    RelDiff= numeric(),
                    class= character(), n=numeric(),
                    size= numeric(),
                    rel= numeric())
  


    # Helper function to add rows if the index is within bounds
  add_row_if_exists <- function(res, baseModel, idx, class, coverClass, plantName, plantStatus, tmpFullPlotData, plotsWherePlantOccurs) {
      res <- add_row(res, 
                     taxa = plantName, 
                     Estimate = baseModel$p.table[idx, 1],
                     StdErr = baseModel$p.table[idx, 2],
                     zValue = baseModel$p.table[idx, 3], 
                     pr = baseModel$p.table[idx, 4],
                     Intercept = baseModel$p.table[1, 1],  
                     numberOfPlots = length(plotsWherePlantOccurs$PlotObservationID), 
                     Neophyte = plantStatus,
                     RelDiff = 
                       (mean(fitted(base)[tmpFullPlotData$coverClass == coverClass]) -
                        mean(fitted(base)[tmpFullPlotData$coverClass == "0%"])) /
                        mean(fitted(base)[tmpFullPlotData$coverClass == "0%"]),
                     class = class, n = sum(tmpFullPlotData$coverClass == coverClass),
                     size= size, rel= length(plotsWherePlantOccurs$PlotObservationID)/ size)
    return(res)
  }
  
  uniqueFactors  <- sort(unique(tmpFullPlotData$coverClass))
  uniqueFactors
  
  for(i in 2:  nrow(baseModel$p.table)){
      index <- uniqueFactors[i]
      res <- add_row_if_exists(res, baseModel, i, index, index, plantName, plantStatus, tmpFullPlotData,
                             plotsWherePlantOccurs)
  }
  res

}

```



## 2.3 Impact
Use foreach to speed up computations
```{r}
if(fast){

# prepare parallel
n.cores<- parallel::detectCores()-2.
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK" )
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

i=58 
# run for all and time
begin<-Sys.time()
x<- foreach(i= 1:length(allPlants$names), 
            .combine = "rbind", 
            .packages = c("dplyr","mgcv","geometry","factoextra")) %dopar% {
  res <- computePlantImpact2(allPlants$names[i], allPlants$Neophyte[i])
            }

end<-Sys.time()
round(end-begin, 2)

parallel::stopCluster(cl = my.cluster)
#write_csv(x, 'Impact2_new70%_ENS0_reldiff_PCA.csv')
}


```





Run this if all data, then foreach does not work
```{r}
if(!fast){

allPlantsImpact <-  data.frame(taxa = character(), Estimate = numeric(), StdErr = numeric(), `zValue` = numeric(), pr = numeric(), numberOfPlots = integer(), Neophyte= character())

i=1
begin<- Sys.time()
for(i in 1:length(allPlants$names)) {
   res <- computePlantImpact2(allPlants$names[i], allPlants$Neophyte[i])
   allPlantsImpact <- rbind(allPlantsImpact, res)
}
end<- Sys.time()
round(end-begin)
x<- allPlantsImpact

}

#write.csv(x, 'Impact2_ENS0.csv')

```

Analyse the cluster size
```{r}
rel <- x$rel[!duplicated(x$taxa)]

hist(rel)
summary(rel)
```



## 2.4 Significance
```{r}
# Probably (but currently only 10% run) no normality for especially the intra class and also very little data compared to the rest, so ranked anova performed
method=2

if(method==1){
result<-(aov(rank(RelDiff) ~ Neophyte, x))
summary(result)
TukeyHSD(result)
} else {
  ggplot(x, aes(x=RelDiff, group= Neophyte, color=Neophyte, fill= Neophyte))+
    geom_histogram(aes(y=..density..), position="identity", alpha=0.5)
  
  ggplot(x, aes(x=RelDiff, group= class, color=class, fill= class))+
    geom_histogram(aes(y=stat(density)),position="identity", alpha=0.5)
  
  # anova --> not possible, sample variances are not equal and sample sizes also differ
  result<-(aov((RelDiff) ~ Neophyte*class, x))
  hist(result$residuals,breaks=100 )
  summary(result)
  z<- TukeyHSD(result)
  z <- z[["Neophyte:class"]]
  
  result <- x |> kruskal_test(RelDiff ~ Neophyte)
  
  library(ARTool)
  x$Neophyte<- as.factor(x$Neophyte)
  x$class <- as.factor(x$class)
  model= art(RelDiff~ Neophyte*class, data=x)
  model
  anova(model)
  marginal= art.con(model, "Neophyte:class", adjust="bonferroni", interaction=F)
  marginal
  plot(marginal)
  

  Sum = groupwiseMedian(RelDiff ~ Neophyte*class,
                      data=x,
                      bca=FALSE, percentile=TRUE)

  Sum
  
  pd = position_dodge(0.4)
  
  ggplot(Sum,
       aes(x     = class,
           y     = Median,
           color = Neophyte)) +

    geom_point(shape  = 16,
               size   = 2,
             position = pd) +

    geom_errorbar(aes(ymin  =  Percentile.lower,
                      ymax  =  Percentile.upper),
                      width =  0.2,
                      size  =  0.7,
                      position = pd) +
    theme_bw() +
    theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0))
}



```


# 3 PLOT
## 3.1 Estimate
```{r}
x <- x %>% mutate(Species = factor(Neophyte, 
                          levels = c("native", "intra","extra"),
                          labels = c("native in the country", 
                                     "intra-European neophyte", 
                                     "extra-European neophyte")))

very_low <- x[x$class =="0%-1%",]
low <- x[x$class =="1%-10%",]
medium <- x[x$class =="10%-30%",]
high <- x[x$class =="30%-50%",]
very_high <- x[x$class =="50%-70%",]
dominant <-  x[x$class =="70%-100%",]
```


### Very low
```{r}
result<-(aov(rank(RelDiff) ~ Neophyte , very_low))
summary(result)
TukeyHSD(result)

my_comparisons <- list( c("extra", "intra"), c("intra", "native"), 
                        c("extra", "native") )

p<-ggplot(very_low, aes(x=Neophyte, y= RelDiff, group = Species,
             color = Species, fill = Species))+
  geom_violin(alpha=0.5, scale="width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", 
                     label.y = c(4.25,4.50, 4.90))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
  scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
  theme(legend.position = "none")+theme(legend.text= element_text(size=12), 
                                          legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(very_low$RelDiff[very_low$Neophyte=="extra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(very_low$RelDiff[very_low$Neophyte=="intra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(very_low$RelDiff[very_low$Neophyte=="native"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=4.00, 
           label= paste("n=",length(very_low$Estimate[very_low$Neophyte=="extra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=4.00, 
           label= paste("n=",length(very_low$Estimate[very_low$Neophyte=="intra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=4.00, 
           label= paste("n=",length(very_low$Estimate[very_low$Neophyte=="native"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.05"))))+
  ylab("Relative difference")+
  ggtitle("Very low cover [0%-1%]")+
  xlab(NULL)+
  geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
very_low_plot <-p
p
```



### Low
```{r}
result<-(aov(rank(RelDiff) ~ Neophyte , low))
summary(result)
TukeyHSD(result)

my_comparisons <- list( c("extra", "intra"), c("intra", "native"), 
                        c("extra", "native") )

p<-ggplot(low, aes(x=Neophyte, y= RelDiff, group = Species,
             color = Species, fill = Species))+
  geom_violin(alpha=0.5, scale="width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", 
                     label.y = c(2.20,2.50, 2.90))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
  scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
  theme(legend.position = "none")+theme(legend.text= element_text(size=12), 
                                          legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(low$RelDiff[low$Neophyte=="extra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(low$RelDiff[low$Neophyte=="intra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(low$RelDiff[low$Neophyte=="native"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=2.00, 
           label= paste("n=",length(low$Estimate[low$Neophyte=="extra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.00, 
           label= paste("n=",length(low$Estimate[low$Neophyte=="intra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.00, 
           label= paste("n=",length(low$Estimate[low$Neophyte=="native"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.05"))))+
  ylab("Relative difference")+
  xlab(NULL)+
  ggtitle("Low cover [1%-10%]")+
  geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
low_plot <-p
p
```


### Medium
```{r}
result<-(aov(rank(RelDiff) ~ Neophyte , medium))
summary(result)
TukeyHSD(result)

my_comparisons <- list( c("extra", "intra"), c("intra", "native"), 
                        c("extra", "native") )

p<-ggplot(medium, aes(x=Neophyte, y= RelDiff, group = Species,
             color = Species, fill = Species))+
  geom_violin(alpha=0.5, scale="width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", 
                     label.y = c(3.20,3.50, 3.90))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
  scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
  theme(legend.position = "none")+theme(legend.text= element_text(size=12), 
                                          legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(medium$RelDiff[medium$Neophyte=="extra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(medium$RelDiff[medium$Neophyte=="intra"],
                                   na.rm=T), 
           label = "ab", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(medium$RelDiff[medium$Neophyte=="native"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=3.00, 
           label= paste("n=",length(medium$Estimate[medium$Neophyte=="extra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=3.00, 
           label= paste("n=",length(medium$Estimate[medium$Neophyte=="intra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=3.00, 
           label= paste("n=",length(medium$Estimate[medium$Neophyte=="native"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P < 0.05"))))+
  ylab("Relative difference")+
  ggtitle("Medium cover [10%-30%]")+
  xlab(NULL)+
  geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
medium_plot <-p
p
```



### High
```{r}
result<-(aov(rank(RelDiff) ~ Neophyte , high))
summary(result)
TukeyHSD(result)

my_comparisons <- list( c("extra", "intra"), c("intra", "native"), 
                        c("extra", "native") )

p<-ggplot(high, aes(x=Neophyte, y= RelDiff, group = Species,
             color = Species, fill = Species))+
  geom_violin(alpha=0.5, scale="width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", 
                     label.y = c(2.40,2.60, 2.80))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
  scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
  theme(legend.position = "none")+theme(legend.text= element_text(size=12), 
                                          legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(high$RelDiff[high$Neophyte=="extra"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(high$RelDiff[high$Neophyte=="intra"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(high$RelDiff[high$Neophyte=="native"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=2.00, 
           label= paste("n=",length(high$Estimate[high$Neophyte=="extra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=2.00, 
           label= paste("n=",length(high$Estimate[high$Neophyte=="intra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=2.00, 
           label= paste("n=",length(high$Estimate[high$Neophyte=="native"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P > 0.05"))))+
  ylab("Relative difference")+
  ggtitle("High cover [30%-50%]")+
  xlab(NULL)+
  geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
high_plot <-p
p
```


### Very high
```{r}
result<-(aov(rank(RelDiff) ~ Neophyte , very_high))
summary(result)
TukeyHSD(result)

my_comparisons <- list( c("extra", "intra"), c("intra", "native"), 
                        c("extra", "native") )

p<-ggplot(very_high, aes(x=Neophyte, y= RelDiff, group = Species,
             color = Species, fill = Species))+
  geom_violin(alpha=0.5, scale="width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", 
                     label.y = c(2.00,2.20, 2.40))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
  scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
  theme(legend.position = "none")+theme(legend.text= element_text(size=12), 
                                          legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(very_high$RelDiff[very_high$Neophyte=="extra"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(very_high$RelDiff[very_high$Neophyte=="intra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(very_high$RelDiff[very_high$Neophyte=="native"],
                                   na.rm=T), 
           label = "ab", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=1.50, 
           label= paste("n=",length(very_high$Estimate[very_high$Neophyte=="extra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=1.50, 
           label= paste("n=",length(very_high$Estimate[very_high$Neophyte=="intra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=1.50, 
           label= paste("n=",length(very_high$Estimate[very_high$Neophyte=="native"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P > 0.05"))))+
  ylab("Relative difference")+
  ggtitle("Very high cover [50%-70%]")+
  xlab(NULL)+
  geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
very_high_plot <-p
p
```



### Dominant
```{r}
result<-(aov(rank(RelDiff) ~ Neophyte , very_high))
summary(result)
TukeyHSD(result)

my_comparisons <- list( c("extra", "intra"), c("intra", "native"), 
                        c("extra", "native") )

p<-ggplot(dominant, aes(x=Neophyte, y= RelDiff, group = Species,
             color = Species, fill = Species))+
  geom_violin(alpha=0.5, scale="width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_pubr()+
  stat_compare_means(comparisons=my_comparisons, label= "p.signif", 
                     label.y = c(2.00,2.20, 2.40))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
  scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
  theme(legend.position = "none")+theme(legend.text= element_text(size=12), 
                                          legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(very_high$RelDiff[very_high$Neophyte=="extra"],
                                   na.rm=T), 
           label = "b", size = 4, vjust = -1, hjust = 0.5, alpha=0.8) +
  annotate("text", x = 2, y = mean(very_high$RelDiff[very_high$Neophyte=="intra"],
                                   na.rm=T), 
           label = "ab", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x = 3, y = mean(very_high$RelDiff[very_high$Neophyte=="native"],
                                   na.rm=T), 
           label = "a", size = 4, vjust = -1, hjust = 0.5, alpha=0.8)+
  annotate("text", x=1, y=1.50, 
           label= paste("n=",length(very_high$Estimate[very_high$Neophyte=="extra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=1.50, 
           label= paste("n=",length(very_high$Estimate[very_high$Neophyte=="intra"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=1.50, 
           label= paste("n=",length(very_high$Estimate[very_high$Neophyte=="native"])), 
           vjust = -0.5, hjust = 0.5, size=3.5)+
  labs(subtitle = substitute(paste("Kruskal-Wallis test ", italic("P > 0.05"))))+
  ylab("Relative difference")+
  ggtitle("Very high cover [50%-70%]")+
  xlab(NULL)+
  geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
dominant_plot <-p
p
```


## 3.2 Combine
```{r}
plot <- ggarrange(very_low_plot, low_plot, medium_plot, high_plot,
                  very_high_plot, dominant_plot, nrow=3, ncol=2, labels= c("a","b", "c","d","e","f"),
                  font.label = list(size = 12))
plot

#ggsave("Impact.jpg", plot= plot, width = 10, height = 10)

```


## 3.3 New plot
```{r}

# take cover classes
cover_class <- unique(x$class)

# if not for loop, choose i value
i=1
  

# can be done with for loop, but than it is hard to change some parameters so better not to
#for(i in 1:6){
  
  # create a vector for the title
  index <- case_when(
    i == 1  ~ "Very low cover",  
    i == 2~ "Low cover"  ,
    i == 3 ~ "Medium cover"  ,
    i == 4 ~ "High cover"  ,
    i == 5~ "Very high cover" ,
    i == 6~ "Dominant cover")
  
  # downsize dataset 
  Dataset <- x[x$class==cover_class[i],]
  
  # create title
  title <- paste(index," ]",cover_class[i], "]", sep="")
  
  # test differences and significance
  fancy=F  
  if(!fancy){
    result<-welch_anova_test(Dataset, (RelDiff) ~ Neophyte )
    result
    games_howell_test(Dataset, RelDiff ~ Neophyte)
    ggdensity(Dataset$RelDiff[Dataset$Neophyte=="extra"])
    ggqqplot(Dataset$RelDiff[Dataset$Neophyte=="extra"])
    shapiro.test(Dataset$RelDiff[Dataset$Neophyte=="extra"])
  }
  
  
  # perform fancy test (has to be updated)
  if(fancy){  
    # bayesian model test with both mean and sigma depending on the group
    fit1 <- brm(bf(RelDiff ~ Neophyte, sigma ~ Neophyte), data = Dataset, 
                family = gaussian())
    summary(fit1)
    
    # calculate the differences between the neophytes
    ggemmeans(fit1,"Neophyte",x.as.factor = TRUE)
  
    # test hypothesis (does not work)
    hypothesis(fit1, "Neophyteintra - Intercept=0")
    
    # also way to compare the species
    fit2.emm.a <- emmeans(fit1, "Neophyte", data=Dataset)
    pairs(fit2.emm.a, adjust="tukey")
    
    # make 0 models (both with and without sigma)
    fit2 <- brm(RelDiff~1,data=Dataset)
    fit3 <- brm(bf(RelDiff~1, sigma ~Neophyte),data=Dataset)
    
    # calculate Bayes factor
    bayes_factor(fit1,fit3)
  }  
    

  # plot
  p<-ggplot(Dataset, aes(x=Neophyte, y= RelDiff, group = Species,
                 color = Species, fill = Species))+
      geom_violin(alpha=0.5, scale="width")+
      geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
      theme_pubr()+
      stat_summary(fun= "mean",
                   geom = "point", aes(group= Neophyte), size=3)+
      scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"))+
      scale_fill_manual(values = c("#1E88E5", "#FFC107", "#004D40")) +
      theme(legend.position = "none")+
      theme(legend.text= element_text(size=12),legend.title =element_text(size=14))+
      annotate("text", x = 1, y = max(Dataset$RelDiff) + max(Dataset$RelDiff)/10, 
               label = "b", size = 4, vjust = 0.5, hjust = 0.5, alpha=1) +
      annotate("text", x = 2, y = max(Dataset$RelDiff) + max(Dataset$RelDiff)/10, 
               label = "b", size = 4, vjust = 0.5, hjust = 0.5, alpha=1)+
      annotate("text", x = 3, y = max(Dataset$RelDiff) + max(Dataset$RelDiff)/10, 
               label = "a", size = 4, vjust = 0.5, hjust = 0.5, alpha=1)+
      annotate("text", x=1, y=max(Dataset$RelDiff) + max(Dataset$RelDiff)/5, 
               label= paste("n=",length(Dataset$Estimate[Dataset$Neophyte=="extra"])), 
               vjust = 0, hjust = 0.5, size=4)+
      annotate("text", x=2, y=max(Dataset$RelDiff) + max(Dataset$RelDiff)/5, 
               label= paste("n=",length(Dataset$Estimate[Dataset$Neophyte=="intra"])), 
               vjust = 0, hjust = 0.5, size=4)+
      annotate("text", x=3, y=max(Dataset$RelDiff) + max(Dataset$RelDiff)/5, 
               label= paste("n=",length(Dataset$Estimate[Dataset$Neophyte=="native"])), 
               vjust = 0, hjust = 0.5, size=4)+
      labs(subtitle = substitute(paste("Welch's ANOVA ", italic("P = 0.87"))))+
      ylab("Relative difference")+
      ggtitle(title)+
      xlab(NULL)+
      geom_hline(yintercept=0, linetype='dotted', alpha=0.5)
    assign(gsub(" ","_",index), p)
    p
#}

# arrange all plots and save
plot <- ggarrange(Very_low_cover, Low_cover, Medium_cover, High_cover,
                  Very_high_cover, Dominant_cover, nrow=3, ncol=2, labels= c("a","b", "c","d","e","f"),
                  font.label = list(size = 12))
plot
ggsave("Impact2.jpg", plot= plot, width = 10, height = 10)
```






