---
title: "Impact with base model for new classification"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---

# 1 START

## 1.1 General
```{r, echo=FALSE}
rm(list=ls())
```


Load packages
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(doParallel)
library(foreach)
```


Here we set some parameters that are useful in the script
```{r}
correctArea<- F
```


## 1.2 Data
Load eva and plot data
```{r, warnings=F, message=F}
eva <- read_csv("fullPlotEva.csv")
fullPlotData <- read_csv("fullPlotData.csv")
```


Downsample by a factor of 100 if fast is selected
```{r}
fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.99,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}

#eva<- read.csv("downsampledEva.csv", header=T)
#colnames(eva)[8]<- "Cover %"
#fullPlotData<- read.csv("downsampledPlotData.csv")
```


## 1.3 Neophyte
Load the neophyte definitions:
```{r, message=F}
# Data on which species are neophytes
native_intra_analysis=T
if(native_intra_analysis){
  species_country_status <- read_csv("eva2_country_status_new.csv")
} else{
species_country_status<- read_csv("species_country_status_new.csv")
# or if we want to do it with intra seperately
}

# Assign these names to the eva list
extra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="extra"])
intra_EU <- unique(species_country_status$species[species_country_status$Neophyte=="intra"])
native_intra <- unique(species_country_status$species[species_country_status$Neophyte=="native_intra"])
```


Remove species that are not defined or to be excluded
```{r, warnings=F}
remove<- read_csv("not_defined.csv",show_col_types = FALSE)
remove<- as.vector(unlist(remove))
eva<- eva[!(eva$species %in% remove),]
```


Assign classification to species
```{r}
# Only observation ID and Region
fullPlot2<- fullPlotData[,c("PlotObservationID","Region")]
# Right join to keep only species present in fullplot (otherwise a lot of NAs)
eva<- right_join(eva, fullPlot2, by = c("PlotObservationID"="PlotObservationID"))
rm(list=c("fullPlot2"))
# Join eva and classification
eva_country_neophyte<- left_join(eva, species_country_status, by= c("Region"= "Region", "species"= "species"))
eva<- eva_country_neophyte
# Look at how much from every type are present
x<- as.data.frame(table(eva_country_neophyte$Neophyte))
```


## 1.4 Number of plots
Reduce EVA to a smaller dataset to accelerate computations.    
Count the number of plots in which the species in the dataset exist.  
```{r, message=F}
if(fast) {
  reducedEva <- eva |> select(PlotObservationID, species, `Cover %`, Neophyte)

  uniqueSpecies <- unique(reducedEva[, c(2,4)])
  
  # Empty dataset

  # Prepare parallel
  parallel::detectCores()
  n.cores <- parallel::detectCores() - 2
  my.cluster <- parallel::makeCluster(
    n.cores, 
    type = "PSOCK"
  )
  print(my.cluster)
  doParallel::registerDoParallel(cl = my.cluster)
  foreach::getDoParRegistered()
  foreach::getDoParWorkers()

  plotsPerSpecies <- data.frame(names = character(), numberOfPlots = integer(), Neophyte= character())
  uniqueSpecies <- unique(reducedEva[, c(2,4)])
  # Count in how many plots a native species occurs
  begin<- Sys.time()
plotsPerSpecies<-foreach(i= 1:nrow(uniqueSpecies), .combine='rbind', .packages=c("dplyr","mgcv")) %dopar% {
    tmp <- reducedEva$PlotObservationID[reducedEva$species == uniqueSpecies$species[i]& reducedEva$Neophyte == uniqueSpecies$Neophyte[i]]
    names=uniqueSpecies$species[i]
    numberOfPlots= length(unique(tmp))
    plotsPerSpecies[i,]<- c(names, numberOfPlots,Neophyte=uniqueSpecies$Neophyte[i] )
}
  plotsPerSpecies<- as.data.frame(plotsPerSpecies)
  plotsPerSpecies<-setNames(plotsPerSpecies, c("names", "numberOfPlots","Neophyte"))
  plotsPerSpecies$numberOfPlots <- as.numeric(plotsPerSpecies$numberOfPlots)
  row.names(plotsPerSpecies)<- c(1: nrow(uniqueSpecies))
  
  end<- Sys.time()
  round(end-begin)
  
  parallel::stopCluster(cl = my.cluster)
} else {
  if(!native_intra_analysis){
  plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope.csv", show_col_types = FALSE)
  plotsPerSpecies<- plotsPerSpecies[, -c(1)]
  } else {
  plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope_native_intra.csv", show_col_types = FALSE)
  }
}
```


## 1.5 Select
We want all species that are present in at least 50 plots, we order this set from highest to lowest and subsequently make a function to compute the total cover of the species across all sites.
```{r}
# If you want to have a look at the distribution of the species richness per plot: hist(plotsPerSpecies$numberOfPlots[plotsPerSpecies$numberOfPlots>1000])

# determine which plant species should be computed
minimumNumberOfPlots<- 50
allPlants <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >= minimumNumberOfPlots,]
allPlants <- allPlants[order(-allPlants$numberOfPlots),]

# function to give the species cover per species in a plot
computeSpeciesCover <- function(name, species, cover) {
  result <- sum(cover[species==name]) # we use the sum here, since otherwise the total cover / species cover cannot reach one.
  if(is.na(result)) {
    0
  } else {
    result
  }
}
```


Test the function
```{r, warnings=F, message=F, results='hide'}
# Tests of the above function
computeSpeciesCover("nonExistentPlant", c("plant1", "plant2"), c(1,2)) # If the plant is not present 0 is returned.
computeSpeciesCover("plant1", c("plant1", "plant2"), c(1,2)) # If the is present the right value is returned
computeSpeciesCover("plant2", c("plant1", "plant2"), c(1,2))
computeSpeciesCover("plant2", c("plant1", "plant2", "plant2"), c(1,2,3)) # If the plant is present multiple times the sum is returned

reducedEva <- eva[,c("PlotObservationID", "species", "Cover %", "Neophyte")]
```


# 2 ANALYSIS
## 2.1 Function
```{r, meassage=F}
# The name of the first species (to test the function)
plantName = allPlants$names[1]
plantStatus= allPlants$Neophyte[1]


# Function to compute the impact for each species, so we only need the plant name as input
computePlantImpact <- function(plantName, plantStatus) {
  
  # Copy dataset temporarily to be able to make changes
  tmpFullPlotData <- fullPlotData
  
  # Get the indices of the plant in the eva dataset
  indexOfPlant <- eva$species == plantName & eva$Neophyte== plantStatus
  
  # We select only the rows containing the species first because this is much faster than doing it for all plots.
  # We make the dataset smaller (only those plots where the species is present), group the data per plot and compute the total cover (all     species) and cover of the species we are highlighting now
  plotsWherePlantOccurs <- reducedEva[reducedEva$PlotObservationID %in% reducedEva$PlotObservationID[indexOfPlant],] |>
    group_by(PlotObservationID) |>
    summarise(totalCover = sum(`Cover %`), 
              speciesCover = computeSpeciesCover(plantName, species, `Cover %`))
  
  
  # Because the cover class was only computed for plots containing the species we need to assign 0% class to the remaining plots.
  # Join the computed cover plots with the full data
  tmpFullPlotData <- left_join(tmpFullPlotData, plotsWherePlantOccurs, by = "PlotObservationID")
  # Identify all plots where the species investigated occurs
  tmpFullPlotData$plantOccurs = tmpFullPlotData$PlotObservationID %in% reducedEva$PlotObservationID[indexOfPlant]
  
  # If species is not present:
  # Make 1 total cover and 0 species cover
  tmpFullPlotData$totalCover[is.na(tmpFullPlotData$totalCover)] <- 1.0
  tmpFullPlotData$speciesCover[is.na(tmpFullPlotData$speciesCover)] <- 0.0
  
  # Exclude all plots which have 0 total cover
  tmpFullPlotData <- tmpFullPlotData[tmpFullPlotData$totalCover > 0,] 
  
  # Do SR - 1 for all plots in which the species is present
  tmpFullPlotData$numberOfVascularPlantSpecies <- tmpFullPlotData$numberOfVascularPlantSpecies - tmpFullPlotData$plantOccurs
  
  # Correct area: if true we calculate the area that is reserved for the other plant species. 
  if(correctArea) {
    tmpFullPlotData$Area <- tmpFullPlotData$Area*(1.0 - tmpFullPlotData$speciesCover/tmpFullPlotData$totalCover)
  }
  
  # There is a numerical problem in the bam function if we provide empty plots...
  tmpFullPlotData <- tmpFullPlotData[tmpFullPlotData$numberOfVascularPlantSpecies > 0 & tmpFullPlotData$Area > 0,]
  
  # Make dataset a factor to add it as a random variable
  tmpFullPlotData$Dataset <- as.factor(tmpFullPlotData$Dataset)
  
  # Count number of plots in which the species is present
  sum(tmpFullPlotData$plantOccurs)
  
  # Base model
  model<-  numberOfVascularPlantSpecies ~ 
              s(log(Area),bs='tp') + 
              s(EIVEresM, bs = 'tp') +
              s(EIVEresN, bs = 'tp') +
              s(EIVEresR, bs = 'tp') +
              s(EIVEresL, bs = 'tp') +
              s(EIVEresT, bs = 'tp') +
              s(DistSeverity.sqrt, bs = 'tp') +
              s(Soil.Disturbance.sqrt, bs = 'tp')+
              s(Grazing.Pressure.sqrt, bs = 'tp')+
              s(Mowing.Frequency.sqrt, bs = 'tp') +
              s(Latitude, Longitude, bs = 'tp') +
              s(Dataset, bs = 're')+
              plantOccurs

base <- bam(model,family = quasipoisson, data = tmpFullPlotData,  method = 'fREML',  discrete=TRUE, nthreads=4)
  
  
  # Get summary from model
  baseModel <- summary(base)

  # Make a dataframe to add the residuals to
  res <- data.frame(taxa = character(), Estimate = numeric(), StdErr = numeric(), `zValue` = numeric(), pr = numeric(), numberOfPlots = integer(), Neophyte= character())
  
  res <- add_row(res, taxa = plantName, Estimate = baseModel$p.table[2,1], StdErr = baseModel$p.table[2,2], zValue = baseModel$p.table[2,3], pr = baseModel$p.table[2,4], numberOfPlots = length(plotsWherePlantOccurs$PlotObservationID), Neophyte= plantStatus)
  res
}
```



## 2.2 Impact
Use foreach to speed up computations
```{r}
if(fast){
# prepare parallel
n.cores<- parallel::detectCores()-6
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK" )
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

# run for all and time
begin<-Sys.time()
x<- foreach(i= 1:length(allPlants$names), .combine = "rbind", .packages = c("dplyr","mgcv")) %dopar% {
  res <- computePlantImpact(allPlants$names[i], allPlants$Neophyte[i])
}
end<-Sys.time()
round(end-begin, 2)

parallel::stopCluster(cl = my.cluster)
#write.csv(x, 'Impact2_new10%.csv')
}
```

```{r}
if(!fast){

allPlantsImpact <-  data.frame(taxa = character(), Estimate = numeric(), StdErr = numeric(), `zValue` = numeric(), pr = numeric(), numberOfPlots = integer(), Neophyte= character())

i=1
begin<- Sys.time()
for(i in 1:length(allPlants$names)) {
   res <- computePlantImpact(allPlants$names[i], allPlants$Neophyte[i])
   allPlantsImpact <- rbind(allPlantsImpact, res)
}
end<- Sys.time()
round(end-begin)
}

```


## 2.3 Significance
```{r}
# Probably (but currently only 10% run) no normality for especially the intra class and also very little data compared to the rest, so ranked anova performed
result<-(aov(rank(Estimate) ~ Neophyte, x))
summary(result)
TukeyHSD(result)
```



# 3 PLOT
```{r}
if(!native_intra_analysis){
p<-ggplot(x, aes(x=Neophyte, y= Estimate, color=Neophyte))+geom_violin()+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
  scale_colour_manual(values=c("#1E88E5", "#FFC107", "#004D40"), 
                    name="Legend",
                    breaks=c("native", "intra","extra"),
                    labels=c("native species", "intra European neophyte", "extra European neophyte"))+
  theme_classic()+ 
  theme(legend.position = "bottom")+theme(legend.text= element_text(size=12), legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(x$Estimate[x$Neophyte=="extra"], na.rm=T), label = "a", size = 3.5, vjust = -1, hjust = 0.5) +
  annotate("text", x = 2, y = mean(x$Estimate[x$Neophyte=="intra"], na.rm=T), label = "a", size = 3.5, vjust = -1, hjust = 0.5)+
  annotate("text", x = 3, y = mean(x$Estimate[x$Neophyte=="native"], na.rm=T), label = "a", size = 3.5, vjust = -1, hjust = 0.5)+
  annotate("text", x=1, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="extra"])), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="intra"])), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="native"])), vjust = -0.5, hjust = 0.5, size=3.5)
p

#ggsave("Impact_base_model_new.jpeg", p, width = 25, height = 15, units = "cm")
} else {
p<-ggplot(x, aes(x=Neophyte, y= Estimate, color=Neophyte))+geom_violin()+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Neophyte), size=3)+
   scale_colour_manual(values=c("#1E88E5","#B71C1C", "#FFC107", "#388E3C"), 
                    name="Legend",
                    breaks=c("native","native_intra", "intra","extra"),
                    labels=c("native species","native species alien elsewhere", "intra European neophyte", "extra European neophyte"))+
  theme_classic()+ 
  theme(legend.position = "right")+theme(legend.text= element_text(size=12), legend.title =element_text(size=14))+ 
  annotate("text", x = 1, y = mean(x$Estimate[x$Neophyte=="extra"], na.rm=T), label = "a", size = 3.5, vjust = -1, hjust = 0.5) +
  annotate("text", x = 2, y = mean(x$Estimate[x$Neophyte=="intra"], na.rm=T), label = "a", size = 3.5, vjust = -1, hjust = 0.5)+
  annotate("text", x = 3, y = mean(x$Estimate[x$Neophyte=="native"], na.rm=T), label = "a", size = 3.5, vjust = -1, hjust = 0.5)+
  annotate("text", x = 4, y = mean(x$Estimate[x$Neophyte=="native_intra"], na.rm=T), label = "b", size = 3.5, vjust = -1, hjust = 0.5)+
  annotate("text", x=1, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="extra"])), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=2, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="intra"])), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=3, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="native"])), vjust = -0.5, hjust = 0.5, size=3.5)+
  annotate("text", x=4, y=0.8, label= paste("n=",length(x$Estimate[x$Neophyte=="native_intra"])), vjust = -0.5, hjust = 0.5, size=3.5)
p

#ggsave("Impact_base_model_new.jpeg", p, width = 25, height = 15, units = "cm") 
}
```

