---
title: "Impact analysis with base model"
author: "Willem Goossens"
date: "`r Sys.Date()`"
output: html_document
---


```{r, echo=FALSE}
rm(list=ls())
```


Load packages
```{r, warnings=F, message=F}
library(readr)
library(dplyr)
library(mgcv)
library(modEvA)
library(ggplot2)
library(doParallel)
library(foreach)
```


# Start
Here we set some parameters that are useful in the script
```{r}
correctArea<- F
```


Load eva and plot data
```{r, warnings=F, message=F}
eva <- read_csv("fullPlotEva.csv")
fullPlotData <- read_csv("fullPlotData.csv")
```


Downsample by a factor of 100 if fast is selected
```{r}
fast <- T
if(fast) {
  fullPlotData <- fullPlotData[runif(length(fullPlotData$PlotObservationID)) > 0.90,]
  eva <- eva[eva$PlotObservationID %in% fullPlotData$PlotObservationID,]
}
```


Load neophyte definitions and check which of the data exists in EVA
```{r, warnings=F, message=F}
neophyteDefinitions <- read_csv("../Neophyte Assignments/UniqueTaxaEurope-2023-04-23.csv")
neophyteNames <- neophyteDefinitions$species[neophyteDefinitions$statusEurope == "neo"]
neophyteNames <- neophyteNames[neophyteNames %in% eva$species]
```


Reduce EVA to a smaller dataset to accelerate computations.  
Count the number of plots in which the species in the dataset exist.
```{r}
if(fast) {
  reducedEva <- eva |> select(PlotObservationID, species, `Cover %`)
  
  uniqueSpecies <- unique(reducedEva$species)
  
  plotsPerSpecies <- data.frame(names = character(), numberOfPlots = integer())
  
  # Count number of plots
  for(i in 1:length(uniqueSpecies)) {
    tmp <- reducedEva$PlotObservationID[reducedEva$species == uniqueSpecies[i]]
    
    plotsPerSpecies <- 
      add_row(plotsPerSpecies, names = uniqueSpecies[i], numberOfPlots = length(unique(tmp)))
  }
} else {
  plotsPerSpecies <- read_csv("plotsPerSpeciesInEurope.csv")
}
```


# Data preparation
We want all species that are present in at least 50 plots, we order this set from highest to lowest and subsequently make a function to compute the total cover of the species across all sites.
```{r}
# determine which plant species should be computed
minimumNumberOfPlots<- 50
allPlants <- plotsPerSpecies[plotsPerSpecies$numberOfPlots >= minimumNumberOfPlots,]
allPlants <- allPlants[order(-allPlants$numberOfPlots),]

computeSpeciesCover <- function(name, species, cover) {
  result <- sum(cover[species==name]) # we use the sum here, since otherwise the total cover / species cover cannot reach one.
  if(is.na(result)) {
    0
  } else {
    result
  }
}
```


Test the function
```{r, warnings=F, message=F}
# Tests of the above function
computeSpeciesCover("nonExistentPlant", c("plant1", "plant2"), c(1,2)) # If the plant is not present 0 is returned.
computeSpeciesCover("plant1", c("plant1", "plant2"), c(1,2)) # If the is present the right value is returned
computeSpeciesCover("plant2", c("plant1", "plant2"), c(1,2))
computeSpeciesCover("plant2", c("plant1", "plant2", "plant2"), c(1,2,3)) # If the plant is present multiple times the sum is returned

reducedEva <- eva[,c("PlotObservationID", "species", "Cover %")]
```
# Function
```{r}
# The name of the first species
plantName = allPlants$names[1]

# Function to compute the impact for each species, so we only need the plant name as input
computePlantImpact <- function(plantName) {
  
  # Copy dataset temporarily to be able to make changes
  tmpFullPlotData <- fullPlotData
  
  # Get the indices of the plant in the eva dataset
  indexOfPlant <- eva$species == plantName
  
  # The majority of the time is spent in computing the cover classes. 
  # We select only the rows containing the species first because this is much faster than doing it for all plots.
  # We make the dataset smaller (only those plots where the species is present), group the data per plot and compute the total cover (all species) and cover of the species we are highlighting now
  plotsWherePlantOccurs <- reducedEva[reducedEva$PlotObservationID %in% reducedEva$PlotObservationID[indexOfPlant],] |>
    group_by(PlotObservationID) |>
    summarise(totalCover = sum(`Cover %`), 
              speciesCover = computeSpeciesCover(plantName, species, `Cover %`))
  
  
  # Because the cover class was only computed for plots containing the species we need to assign 0% class to the remaining plots.
  # Join the computed cover plots with the full data
  tmpFullPlotData <- left_join(tmpFullPlotData, plotsWherePlantOccurs, by = "PlotObservationID")
  # Identify all plots where the species investigated occurs
  tmpFullPlotData$plantOccurs = tmpFullPlotData$PlotObservationID %in% reducedEva$PlotObservationID[indexOfPlant]
  
  # If species is not present:
  # Make 1 total cover and 0 species cover
  tmpFullPlotData$totalCover[is.na(tmpFullPlotData$totalCover)] <- 1.0
  tmpFullPlotData$speciesCover[is.na(tmpFullPlotData$speciesCover)] <- 0.0
  
  # Exclude all plots which have 0 total cover
  tmpFullPlotData <- tmpFullPlotData[tmpFullPlotData$totalCover > 0,] 
  
  # Do SR - 1 for all plots in which the species is present
  tmpFullPlotData$numberOfVascularPlantSpecies <- tmpFullPlotData$numberOfVascularPlantSpecies - tmpFullPlotData$plantOccurs
  
  # Correct area: if true we calculate the area that is reserved for the other plant species. 
  if(correctArea) {
    tmpFullPlotData$Area <- tmpFullPlotData$Area*(1.0 - tmpFullPlotData$speciesCover/tmpFullPlotData$totalCover)
  }
  
  # There is a numerical problem in the bam function if we provide empty plots...
  # This should normally not be possible (as we did data screening) but I will check again
  tmpFullPlotData <- tmpFullPlotData[tmpFullPlotData$numberOfVascularPlantSpecies > 0 & tmpFullPlotData$Area > 0,]
  
  # Make dataset a factor to add it as a random variable
  tmpFullPlotData$Dataset <- as.factor(tmpFullPlotData$Dataset)
  
  # Count number of plots in which the species is present
  sum(tmpFullPlotData$plantOccurs)
  
  # Base model
  gam11 <- bam(numberOfVascularPlantSpecies ~ 
                 log(Area) + 
                 s(EIVEresM, bs = 'tp') +
                 s(EIVEresN, bs = 'tp') +
                 s(EIVEresR, bs = 'tp') +
                 s(EIVEresL, bs = 'tp') +
                 s(EIVEresT, bs = 'tp') +
                 s(transformedDisturbanceSeverity, bs = 'tp') +
                 s(transformedDisturbanceFrequency, bs = 'tp') +
                 s(Latitude, Longitude, bs = 'tp') +
                 s(Dataset, bs = 're') +
                 plantOccurs, 
               family = poisson(link = log), 
               data = tmpFullPlotData, 
               method = 'fREML',
               discrete=TRUE,
               nthreads=1)
  
  # Get summary from model
  gam11.sum <- summary(gam11)
  
  # Make a dataframe to add the residuals to
  res <- data.frame(taxa = character(), Estimate = numeric(), StdErr = numeric(), `zValue` = numeric(), pr = numeric(), numberOfPlots = integer())
  
  res <- add_row(res, taxa = plantName, Estimate = gam11.sum$p.table[3,1], StdErr = gam11.sum$p.table[3,2], zValue = gam11.sum$p.table[3,3], pr = gam11.sum$p.table[3,4], numberOfPlots = length(plotsWherePlantOccurs$PlotObservationID))
  res
}
```


# Impact
Run whole function for smaller dataset
```{r}
allPlantsImpact <- data.frame(taxa = character(), Estimate = numeric(), StdErr = numeric(), `zValue` = numeric(), pr = numeric(), numberOfPlots = integer())

#testSpecies<- uniqueSpecies[1:100]

#for(j in 1: round(length(uniqueSpecies)/100)){
 # start<- 100*(j-1)+1
 # end<- 100*j
#  for(i in start:end) {
 # res <- computePlantImpact(allPlants$names[i])
  #allPlantsImpact <- rbind(allPlantsImpact, res)
 # }
  #name<-paste("allPlantImpact",j, sep = "")
  #assign(name,allPlantsImpact)
#}

#allPlantImpact94$neophyte<- allPlantImpact94$taxa %in% neophyteNames

#ggplot(allPlantImpact94, aes(x=neophyte, y= Estimate))+geom_violin()+ stat_summary(fun= "mean",  geom = "point", aes(group= neophyte), size=3)

#stat.test<-t.test((Estimate) ~ neophyte, allPlantImpact94)
```


Use foreach to speed up computations
```{r}
n.cores<- parallel::detectCores()-2 
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK" )
print(my.cluster)
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

allPlantsImpact <- data.frame(taxa = character(), Estimate = numeric(), StdErr = numeric(), `zValue` = numeric(), pr = numeric(), numberOfPlots = integer())
length(allPlants$names)

begin<-Sys.time()
x<- foreach(i= 1:length(allPlants$names), .combine = "rbind", .packages = c("dplyr","mgcv")) %dopar% {
  res <- computePlantImpact(allPlants$names[i])
}
end<-Sys.time()
round(end-begin, 2)

x$neophyte<- x$taxa %in% neophyteNames

length(unique(x$taxa[x$neophyte]))

ggplot(x, aes(x=neophyte, y= Estimate))+geom_violin()+
  stat_summary(fun= "mean",
               geom = "point", aes(group= neophyte), size=3)

mean(x$Estimate[x$neophyte])
mean(x$Estimate[!x$neophyte])
t.test(Estimate~neophyte, x)
parallel::stopCluster(cl = my.cluster)
```

